### 2021-10-11 chapters/acknowledgements.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}
\chapter{Acknowledgements}

The most important person for me to acknowledge is my wife Andrea. 
It was only with her forbearance and encouragement that I was able to return to University, obtain my Honours in Advanced Science (Research) and continue on towards my doctorate over a total of a seven year period.
I would also like to thank my daughters: Lauren, Sophie and Phoebe, who have all literally grown up during this period.
I look forward to spending more time with you after this submission.
And of course, I am pleased to acknowledge my supervisors, Paul and Andy, who in their infinite patience encouraged me on this journey, particularly when covid-19 tried to stop the world in its tracks.
My peers in PhD room 125 both present and past, have made life interesting and fun over the last three years.
I'm sure you are all heading towards brilliant futures.
This research was supported by the Faculty of Science Dean's Postgraduate Research Scholarship without which I would not have had the opportunity to attempt this.

\end{document}

### 2021-10-11 chapters/conclusion.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}

\chapter{Conclusion}\label{chapter:Conclusion}
    The direct observation of gravitational waves from the binary neutron star inspiral of GW170817 and coincident detections from GRB 170817A and AT2017gfo launched the nascent field of multi-messenger gravitational-wave astronomy.
    Gravitational waves from the late inspiral and post-merger remnant of GW170817 were not detected because the gravitational-wave detectors lacked sensitivity in high frequencies.
    High frequency gravitational-wave detectors have been proposed  which may allow gravitational-wave observations of the binary neutron star post-merger remnant and late inspiral in the near-future. \par

    Observations of gravitational waves from the late inspiral will inform the equation of state for cold neutron stars, whereas gravitational wave detections from surviving post-merger remnants will inform the extremely dense, hot equation of state.
    Surviving post-merger remnants are binary neutron star remnants that have not yet collapsed to a black hole.
    We need models of the post-merger gravitational-wave strain to enable observations of gravitational waves from post-merger remnants.
    Numerical-relativity simulations are the best basis for these post-merger models. \par

    We should not expect that future observable post-merger gravitational waves will precisely match numerical-relativity simulations, as the simulations are extremely complex.
    Compromises must be made in spatial-resolution, incorporated physics, and convergence testing, for simulations to complete in practical time-frames.
    The absolute accuracy of the phase evolution and the collapse time of the post-merger remnant are uncertain due to these compromises.\par

    With these computational limitations in mind, we need simple models that can be used for gravitational-wave detection and parameter estimation.
    The main purpose of this thesis was to focus on this task.
    We achieve this goal by developing a hierarchical model that can infer the progenitor neutron star system parameters and generate gravitational-wave spectra.
    We develop a simple analytical frequency-drift model which successfully matches the numerical-relativity gravitational-wave strain. 
    We modify the frequency-drift model to allow for collapsing remnants so that we can measure when the post-merger remnant collapses to a black hole.

    In Chapter~\ref{chapter:ComputingFastReliable}, we train our hierarchical model on 35 numerical-relativity simulations and measure leave-one-out cross-validated fitting-factors with a mean of 0.95. 
    We show that we can generate gravitational-wave spectra in a fraction of a second, given only the neutron star mass, $M$, and the tidal coupling constant, $\kappa_2^T$.
    We perform parameter estimation on $\kappa_2^T$ by using the fitting factor as a proxy for the likelihood.
    The hierarchical model is particularly suited to modelling numerical-relativity simulations as it is insensitive to phase uncertainties that can occur between simulations with adjacent spatial resolutions.

    In Chapter~\ref{chapter:DetectionPE},  we implement a third-order exponentially-damped sinusoid with a frequency-drift term to successfully capture the complex gravitational-wave strain from simulations of post-merger remnants.
    We find that we can achieve noise-weighted fitting-factors of $\gtrsim 0.92$, with most fitting factors of $\sim 0.95 - 0.97$.
    We use a detection threshold with a Bayes factor of $\sim 3000$ and find that a post-merger signal-to-noise ratio of $\gtrsim 7$ is needed to detect the post-merger remnant.
    We also find that the sensitivity to the uncertainty in the coalescence time is negligible for any detectable post-merger remnant.
    We calculate the dominant post-merger oscillation frequency, $\fpeak$, from the model and constrain this at 95\% credible intervals to $\sim 1.5\%$ for a post-merger signal-to-noise ratio of 15, and $\sim 0.3\%$ for a post-merger signal-to-noise ratio of 50. 
    We then use the hierarchical model trained in Chapter~\ref{chapter:ComputingFastReliable} to infer the equation of state parameters $\kappa_2^T$ and $C$.
    At 95\% confidence intervals, we constrain $\kappa_2^T$ to $\sim 12\%$ for a post-merger signal-to-noise ratio of 15, and $\sim 5\%$ for a post-merger signal-to-noise ratio of 50.
    Similarly, we constrain $C$ to $\sim 5\%$ for a signal-to-noise ratio of 15, and $\sim 2\%$ for a post-merger signal-to-noise ratio of 50.
    % However, the numerical-relativity simulation under test uses a different numerical code to the simulations in which we train the model.
    % This results in a slight bias between the true $\kappa_2^T$ value and the recovered posteriors and a larger bias in the posteriors for $C$ that can be accounted for in future work by propagating uncertainties in the power-law model.

    In Chapter~\ref{chapter:CollapseTime},   we add a collapse-time function to our frequency-drift model and measure at what distance the collapse time can be measured.
    We find that we need a combination of Einstein Telescope and Cosmic Explorer interferometers to detect a collapse time of $10\,\ms$ at a GW170817-like distance of $\sim 40\,\Mpc$.
    For a network of 2 A+ detectors, a distance of $\sim\,10\Mpc$ is required to detect a $10\,\ms$ collapse time.
    If the high-frequency NEMO detector is added to the 2 A+ network then this detection distance is increased to $\sim 25\,\Mpc$. \par

    These three models address the requirements for future post-merger gravitational-wave detections.
    The post-merger models consistently achieve $\gtrsim 0.90$ fitting factors, making them suitable for matched-filter detection of gravitational-wave signals.
    However, a number of enhancements can be anticipated for future work. \par

    An obvious model enhancement would allow for unequal-mass binaries.
    In this case, the number of available numerical-relativity simulations are significantly reduced.
    This does not exclude training a modified hierarchical model on the unequal-mass simulations, though careful attention should be  paid to the error propagation to count for the reduced number of simulations in this part of the training set.
    The frequency-drift model, with or without the collapse time extension, has yet to be tested with unequal-mass mergers.
    This could be tested and the model updated if required. \par

    Measurements of larger collapse times are limited by waveform systematics in the late post-merger gravitational-wave strain.
    The same waveform systematics are also present in the model used in Chapter~\ref{chapter:DetectionPE} for post-merger signals of length $\gtrsim 15\,\ms$.
    A couple of assumptions can be reassessed for future work on the frequency-drift model to help reduce these systematics. 
    The first is that the linear frequency-drift term may be too restrictive for longer signals and other analytic or semi-analytic methods could be considered.
    Secondly, the assumption of exponential damping could be reassessed, considering, for example, a power-law decay in the amplitude instead.
    \par

    The lifetimes of post-merger remnants are highly uncertain with gravitational-wave timescales ranging from $0.15-12\,\s$ for a range of ellipticities and equations of state (Eq.~\ref{eq:Intro:GWAngularMomentumLoss}).
    If the above waveform systematics are successfully addressed then detection and parameter estimation studies could be performed with much longer collapse times. 
    A successful injection study for collapse times up to $\sim 2\,\s$ may allow the measurement of collapse times for future GW170817-like events.
    This would have significantly aided model selection for the gamma-ray physics associated with GRB 170817A.\par

    For example, measuring any collapse of the post-merger remnant would rule out proposals that rely on surviving neutron star remnants~\cite[e.g.,][]{Yu2018}.
    Measuring a collapse time  of $\lesssim 0.5\,\s$ would reduce support for scenarios in Refs.~\cite{Gill2019,Murguia-Berthier2021} where collapse
    times of at least $\sim 1\,\s$ are required so that jets are launched after black hole formation. 
    % \par
    % The delay time between the short gamma-ray burst after the inferred merger time for GW17081 is, $\sim 1.7\,\s$.
    % The similarity between this delay time and the observed burst duration for GRB 170817A, $\sim~2\,\s$ informed a model of GRB 170817A where the majority of the delay time was caused by the propagation of the GRB~\cite{Zhang2018}.
    % This model is indifferent to the form of the central engine other than the requirement that the jet is launched and breakout occurs $\lesssim 0.5\,\s$. 
    A small collapse time of $\lesssim 0.05\,\s$ will rule out models that rely on the temporary survival of the post-merger remnant to account for the colour of kilonova emission~\cite{Metzger2018}.
    This highlights a few examples of how the collapse times may influence the physics of short gamma-ray bursts.\par

    The models developed in this thesis could be used to develop detection pipelines for post-merger remnants.
    Pipelines already exist that are capable of searching for post-merger remnants: \texttt{cWB}, \texttt{BayesWave}, \texttt{Viterbi}, \texttt{STAMP}, \texttt{ATrHough}, and  \texttt{FreqHough}~\cite{Thrane2011, Cornish2015,Littenberg2015,Viterbi1967,Klimenko2016, Suvorova2016,GW170817Postmerger1,Chatziioannou2017,Miller2018, Torres-Rivas2019, Oliver2019,GW170817Postmerger2}.
    A number of these tools searched for a long-lived post-merger remnant from the GW170817 merger, with searches lasting $\lesssim 500\,\s$ (\texttt{STAMP}, \texttt{cWB})~\cite{GW170817Postmerger1} and searches ranging from $\sim 2 - 24\,$hours (\texttt{Viterbi}, \texttt{STAMP}, \texttt{ATrHough}, and  \texttt{FreqHough})~\cite{GW170817Postmerger2}.
    Searches were also performed over $\lesssim 1\,\s$ (\texttt{cWB})~\cite{GW170817Postmerger1} and \texttt{BayesWave} has characterised the post-merger remnant from binary neutron star mergers~\cite{Clark2016postmerger,Chatziioannou2017,Torres-Rivas2019,Easter2020}. \par

    Pipelines developed from the analytical models in Chapters~\ref{chapter:DetectionPE}-\ref{chapter:CollapseTime} could be designed to work with signals produced from numerical-relativity simulations with lengths of $\lesssim 100\,\ms$, or long signals of $\lesssim 2\,\s$.
    Pipelines that searched for the long-lived post-merger remnant should be used for signals longer than this.
    We propose two pipelines from this work: a short pipeline with a length of $\lesssim 100\,\ms$ and a long pipeline with length  $\lesssim 2\,\s$. \par

    Pipelines that are of direct interest for comparison with the analytical model are \texttt{BayesWave} and  \texttt{cWB}.
    We find in Chapter~\ref{chapter:DetectionPE} that \texttt{BayesWave} is more sensitive when measuring the dominant post-merger frequency, $\fpeak$.
    \texttt{BayesWave} and the analytical model constrained $\fpeak$ to post-merger signal-to-noise ratios of $\gtrsim 9$, and $\gtrsim 15$, respectively. 
    The \texttt{cWB} pipeline is expected to be the least sensitive as it detects unmodelled coherent excess power, assuming waveform systematics have been addressed in the analytical model. \par

    The proposed pipeline would use Bayesian inference for detection and parameter estimation which will find posteriors directly related to the gravitational-wave strain of the post-merger remnant.
    This is an advantage over both \texttt{BayesWave} and \texttt{cWB}. 
    The Bayes factor would serve as a robust detection statistic in this situation due to the loud inspiral signal and the tight constraint expected on the coalescence time.
    It may be necessary to perform computational optimisation to allow the long pipeline to execute in reasonable time frames, though it may be possible to use parallel Bilby~\cite{Smith2020} to speed up this pipeline. \par

    The hierarchical model can only be used on the short pipeline because it is trained on numerical-relativity simulations.
    It is likely that the model will need to be extended to account for both spins and unequal-mass neutron stars, this will need to be assessed.
    The hierarchical model can be trained on as many numerical-relativity simulations as possible with careful consideration taken in selecting the spatial resolution.
    The hierarchical model would then be available to infer equation of state parameters from the posterior waveforms generated from the analytical pipeline.
\par
    Although we are not expecting binary neutron star post-merger detections in the immediate future, they may be possible if proposed high-frequency gravitational-wave detectors like NEMO are constructed, or failing that, when Einstein telescope and Cosmic Explorer detectors are operational.
    With this in mind, we must ensure that we have the tools available to fully utilise any future binary neutron star post-merger detections so that we can probe the extreme physics of these remnants.

    \par

\end{document}

### 2021-10-11 chapters/copyright.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}

\noindent This thesis must be used only under the normal conditions of “fair dealing” under the Copyright Act. It should not be copied or closely paraphrased in whole or in part without the written consent of the author. Proper written acknowledgement should be made for any assistance obtained from this thesis. I certify that I have made all reasonable efforts to secure copyright permissions for third-party content included in this thesis and have not knowingly added copyright content to my work without the owner’s permission.
\\[1cm]
\copyright\ \  Paul Easter

\end{document}

### 2021-10-11 chapters/declaration.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}
\chapter{Declaration of Authorship}
I hereby declare that this thesis contains no material which has been accepted for the award of any other degree or diploma at any university or equivalent institution and that, to the best of my knowledge and belief, this thesis contains no material previously published or written by another person, except where due reference is made in the text of the thesis.

This thesis includes two original papers published in peer reviewed journals and one submitted publications. The core theme of this thesis is probing the physics of binary neutron star post-merger remnants with gravitational waves. The ideas, development and writing up of all the papers in the thesis were the principal responsibility of myself, the student, working within the School of Physics and Astronomy under the supervision of Assoc. Prof. Paul Lasky and Dr. Andrew Casey.

The inclusion of co-authors reflects the fact that the work came from active collaboration between researchers and acknowledges input into team-based research.

I have renumbered sections of submitted or published papers in order to generate a consistent presentation within the thesis.

In the case of Chapters 2, 3,and 4 my contribution to the work is shown on the following page. \\[0.75cm]
%
\tabto{0cm}\textbf{Student name:}  Paul John Easter \\[0.75cm]
\tabto{0cm}\textbf{Student signature:} \tabto{10cm} Date: 8 August 2021 \\[0.75cm]
%
%
I hereby certify that the above declaration correctly reflects the nature and extent of the student’s and co-authors’ contributions to this work. In instances where I am not the responsible author I have consulted with the responsible author to agree on the respective contributions of the authors. \\[0.75cm] 
%
\tabto{0cm}\textbf{Main Supervisor name:} Paul D Lasky \\[0.75cm]
\tabto{0cm}\textbf{Main Supervisor signature:}	\tabto{10cm} Date: 8 August 2021

\pagebreak

\begin{turn}{-90}

% \begin{table}
\begin{tabular}{|p{0.12\linewidth}|p{0.40\linewidth}|p{0.15\linewidth}|p{0.3\linewidth}|p{0.5\linewidth}|p{0.15\linewidth}|}
\hline
Thesis Chapter	& Publication Title	& Status	 
& Student contribution & Co-author contribution	& Monash student co-author\\ \hline \hline

2	
& Computing Fast and Reliable Gravitational Waveforms of Binary Neutron Star Merger Remnants	
& Published (PRD)	
& 75\% \newline Concept, method, code development, analysis, writing	 
& P. Lasky/A. Casey \newline 15\%-Concept, method, analysis, writing \newline L. Rezzolla/K. Takami \newline 10\% - analysis, writing 
& No\\
\hline
3	
& Detection and parameter estimation of binary neutron star merger remnants	
& Published (PRD)	
& 70\% \newline Concept, method, code development, analysis, writing	 
& S. Ghonge \newline 10\%-Concept, method, code development, analysis, writing \newline 
F. Hernandez Vivanco \newline 10\% - Concept, method, code development, analysis, writing \newline 
P. Lasky/A. Casey \newline 5\%-Concept, method, analysis, writing \newline 
J. Clarke/K. Chatziioannou \newline 5\% - Concept, analysis, writing
& Yes \newline Hernandez Vivanco\\
\hline

4	& Can we measure the collapse time of a post-merger remnant for a future GW170817-like event? 
& Submitted (PRD)
& 75\%  Concept, method, code development, analysis, writing	 
& P. Lasky/A. Casey \newline 25\%-Concept, method, analysis, writing 
& No\\
\hline

\hline				
\end{tabular}
% \end{table}

\end{turn}

\end{document}

### 2021-10-11 chapters/introduction.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

\begin{document}

\begingroup
\clearpage% Manually insert \clearpage
\let\clearpage\relax% Remove \clearpage functionality
\vspace*{-2cm}% Insert needed vertical retraction

\chapter*{Abstract}
\endgroup
\addcontentsline{toc}{chapter}{Abstract} 
% Paper 1
% Here we introduce a hierarchical model trained on numerical-relativity simulations, which can generate reliable post-merger spectra in a fraction of a second. Our spectra have mean fitting factors of 0.95, which compares to a fitting factor of 0.93 between different numerical-relativity codes that simulate the same physical system. This method is the first step towards generating large template banks of spectra for use in post-merger detection and parameter estimation.
% Paper 2
% Here we develop a fast, simple model that can generate gravitational waveforms, and show it can be used for both detection and parameter estimation of post-merger remnants.
% The model consists of three exponentially-damped sinusoids with a linear frequency-drift term.
% We test the model against nine equal-mass numerical-relativity simulations selected for emission of gravitational waves for $\gtrsim25$\,ms.
% The median fitting factors between the model waveforms and numerical-relativity simulations exceed 0.90.
% We detect remnants at a post-merger signal-to-noise ratio of $\ge 7$ using a Bayes-factor detection statistic with a threshold of 3000.
% We can constrain the primary post-merger frequency to $\pm_{1.2}^{1.4}\%$ at post-merger signal-to-noise ratios of 15 with an increase in precision to $\pm_{0.2}^{0.3}\%$ for post-merger signal-to-noise ratios of 50.
% The tidal coupling constant can be constrained to $\pm^{9}_{12}\%$ at post-merger signal-to-noise ratios of 15, and $\pm 5\%$ at post-merger signal-to-noise ratios of 50 using a hierarchical inference model.
% Paper 3
% We develop and test a method to measure the collapse time of post-merger remnants.
% We show that for a GW170817-like event at $\sim\!40$\,Mpc, a network of Einstein Telescope with Cosmic Explorer is required to detect collapse times of $\sim\!10$\,ms.
% For a two-detector network at A+ design sensitivity, post-merger remnants with collapse times of $\sim\!10\,\mathrm{ms}$ must be $\lesssim 10$\,Mpc to be measureable.
% This increases to $\sim\!18\Hyphdash*26$\,Mpc if we include the proposed Neutron star Extreme Matter Observatory (NEMO), increasing the effective volume by a factor of $\sim\!30$.
    % Surviving neutron star remnants from binary neutron star mergers are some of the most extreme forms of matter known in the Universe.
    Surviving remnants from binary neutron star mergers, remnants that have not yet collapsed to a black hole,  are some of the most extreme forms of matter known in the Universe.    
    The densities of surviving merger remnants are $\sim 10^{14}-10^{15}\,\mathrm{g\, cm^{-3}}$ which is up to several times nuclear density.
    The temperature of the post-merger remnant can reach values from tens to hundreds of $\mathrm{MeV}$.
    This combination of extreme temperatures and densities result in a state of matter that is observable only in proto-neutron stars and post-merger remnants.
    Although binary neutron star post-merger remnants have yet to be directly observed, they are a target for gravitational-wave detection in the near future. \par

    A surviving post-merger remnant could potentially be detected by applying a matched filter between an incoming gravitational-wave signal and the gravitational-wave strain generated from numerical-relativity simulations.
    However, these simulations are performed in full general relativity and are extremely expensive.
    This prohibits their direct use in the detection and parameter estimation of post-merger remnant properties.
    Other methods are required to generate post-merger gravitational waveforms that are computationally affordable. \par

    Here we develop two methods that can generate gravitational waveforms of post-merger remnants.
    In the first method, we train a hierarchical model to find a relationship between the progenitor neutron star properties and gravitational-wave spectra produced by numerical-relativity simulations.
    After training, we generate gravitational-wave spectra in a fraction of a second and measure cross-validated fitting-factors with a mean of 0.95 where 1.0 is a perfect fit.\par

    For the second method, we introduce an analytical model that can successfully capture the richness of post-merger gravitational waveforms generated by numerical-relativity simulations.
    We measure median fitting factors between the gravitational-wave strain produced by numerical-relativity simulations and our inferred waveforms as $> 0.9$.
    With this method, we can detect surviving post-merger remnants with signal-to-noise ratios of $\geqslant 7$.
    We find that, at signal-to-noise ratios of 15, the dominant post-merger oscillation frequency can be constrained to $\pm_{1.2}^{1.4}\%$, and the tidal coupling constant can be constrained to $\pm^{9}_{12}\%$. \par

    We  then measure when the post-merger remnant will collapse into a black hole.
    We show that we need a gravitational-wave detector network of Einstein Telescope with Cosmic explorer to measure a collapse time of $\sim\!10$\,ms for a GW170817-like event at $\sim\!40$\,Mpc.
    These three methods introduce additional tools that allow for the detection and parameter estimation of post-merger remnants when gravitational-wave instruments achieve sufficient sensitivity.

\begingroup
\clearpage% Manually insert \clearpage
\let\clearpage\relax% Remove \clearpage functionality
\vspace*{-2cm}% Insert needed vertical retraction

\chapter{Introduction} \label{chapter:Introduction}
\endgroup

% ***************************************************************************************************************************************************************
    \section{Gravitational waves} 
    In 2017, the LIGO and Virgo collaborations observed the first direct gravitational waves from a binary neutron star merger, GW170817~\cite{GW170817Detection}.
    This began the era of multi-messenger gravitational-wave observations with coincident detections in many parts of the electromagnetic spectrum~\cite{Goldstein2017,Coulter2017,Troja2017,Nicholl2017,Chornock2017,Margutti2017,Alexander2017,GW170817multi}. \par

    Gravitational waves are a phenomenon predicted by general relativity which was formulated more than 100 years ago by Albert Einstein with the famous formula
    \begin{align}
        G_{\mu\nu} & = R_{\mu\nu} - \frac{1}{2}R  g_{\mu\nu}  =  \frac{8\pi G}{c^4} T_{\mu\nu}, \label{eq:Intro:GeqT}
    \end{align}
    where $G_{\mu\nu}$ is the space-time curvature tensor and $T_{\mu\nu}$ is the stress energy tensor. The local metric tensor is $g_{\mu\nu}$, $R_{\mu\nu}$ is the Ricci tensor and $R$ is the Ricci scalar. 
    The gravitational constant is $G$ and the speed of light in vacuum is $c$, though for remainder of this work we will use geometric units of $G=c=1$ unless otherwise stated. 
    Equation~\ref{eq:Intro:GeqT} demonstrates how the curvature of the space-time, $G_{\mu\nu}$, alters the dynamics of matter, $T_{\mu\nu}$, and vice-versa. \par
    Movement of matter in space can generate gravitational waves.
    For simulation purposes, the movement of matter can be modelled by the Einstein field equations~Eq.~\ref{eq:Intro:GeqT} and perturbations of black holes~\cite[e.g.,][]{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}.
    This methodology allows the calculation of multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$.
    These multipoles can be evaluated by numerical-relativity simulations which allow the extraction of gravitational waves.
% \redtext{Movement of matter in space generates mass multipoles that can be related to perturbations of black holes.
%     These perturbations can be described in terms of the non-rotating metric for black holes, the Schwarzschild metric, and the Einstein field equations,~Eq.~\ref{eq:Intro:GeqT}.
%     Non-spherical gauge-invariant perturbations, outlined in Ref.~\cite{Baiotti2008}, can generate master multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$, that can be estimated my numerical-relativity simulations (see also Refs.~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}).}
% \redtext{
%     Gravitational waves can be extracted for a virtual observer on the 2-sphere of a Schwarzschild metric with an isotropic radius of $r$ from the source.
%     The Schwarzschild radius for this metric is $r_S=r(1-M/2r)^2$, where $M$ is the enclosed mass~\cite[e.g.,][]{Baiotti2008}.
%     Gravitational waves are extracted through non-spherical gauge-invariant perturbations of the Schwarzschild metric, resulting in multipoles  where $(l,m)$ are the angular indices~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}.\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}\par  % Matthews1962
%     % Check out: Abrahams, Rezzolla, Rupright, Anderson, Anninos, Baumgarte, et al., Phys. Rev. Lett. 80, 1812 (1998) 
%     % Rupright, Abrahams, Rezzolla, Phys. Rev. D 58, 044005 (1998).
%     % Rezzolla, Abrahams, Matzner, Rupright, Shapiro, Phys. Rev. D 59, 064001 (1999).
%     }
    The plus, $h_{+}$, and cross, $h_{\times}$, polarisations of the gravitational waves are related to the multipoles as follows~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}:\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}
     \begin{align}
        h_{+} - i h_{\times} & = \frac{1}{r\sqrt{2}}\sum_{l,m} {}_{-2}Y^{lm}\left(\mathcal{Q}^{+}_{lm} -i \int_{-\infty}^t\mathcal{Q}^\times_{lm}(t')dt'\right). \label{eq:Intro:MultipoleExpansion}
    \end{align}   
    Here, $r$ is the distance to the gravitational-wave source and ${}_{-2}Y^{lm}$ are the spin-weighted spherical harmonics with spin $s=-2$ and angular indices of $(l,m)$.
    For most sources, gravitational waves are predominantly emitted in the $l=m=2$ mode. 
    In this case Eq.~\ref{eq:Intro:MultipoleExpansion} evaluates to
     \begin{align}
        h_{+} + i h_{\times} & = \frac{1}{2 r} \sqrt{\frac{5}{2\pi }} e^{2 i \phi } \cos ^4\left(\frac{\theta }{2}\right)\left(\mathcal{Q}^{+}_{22} -i \int_{-\infty}^t\mathcal{Q}^\times_{22}(t')dt'\right), \label{eq:Intro:MultipoleExpansionlm22}
    \end{align}
    where  $\theta$ is the polar angle and $\phi$ is the azimuthal angle usually defined relative to orbital angular momentum vector. \par

    The energy loss due to gravitational waves can be calculated from Eq.~\ref{eq:Intro:MultipoleExpansion} as~\cite{Nagar2005}
     \begin{align}
        \frac{dE}{dt} & = \frac{1}{32 \pi}\sum_{l,m} \left(\left|\frac{\mathrm{d} \mathcal{Q}^{+}_{lm}}{dt}\right|^2 + \left|\mathcal{Q}^\times_{lm}\right|^2\right). \label{eq:Intro:EnergyLoss}
    \end{align}
    For binary systems, this results in an incremental reduction in the orbital separation which was famously measured in the Hulse-Taylor binary neutron star, B1913+16~\cite{Hulse1975,Taylor1982,Weisberg2016}. \par

    If the separation between binary compact objects is small enough, then this loss of energy can lead to a merger.
    This was the case for the first direct detection of a binary black hole merger, GW150914, and first neutron star merger, GW170817~\cite{GW150914Detection,GW170817Detection}. 
    The decrease in the orbital separation and increase in gravitational-wave amplitude, results in a characteristic `chirp' signal for the inspiral of binary compact objects.

    Assuming a circular orbit of two point masses of mass $M_1$ and $M_2$, the Newtonian quadrupole formula can be used to calculate the evolution of the orbital separation, $a$, due to gravitational wave emission:~\cite[e.g.,][]{Cutler1994}
    \begin{align}
        \frac{da}{dt} & = -\frac{64}{5}\frac{\mu M^2}{a^3}, \label{eq:Intro:OrbitalEvolution}        
    \end{align}
    where  $M \equiv M_1 + M_2$ is the total mass, and $\mu \equiv M_1 M_2 / M$ is the reduced mass.
    The time evolution of gravitational-wave strain frequency can then be calculated as
    \begin{align}
        \frac{df}{dt} & = \frac{96}{5}\pi^{8/3}\mathcal{M}^{5/3}f^{11/3}, \label{eq:Intro:FreqEvolution}        
    \end{align}   
    where $\mathcal{M} \equiv \mu^{3/5} M^{2/5}$ is the chirp mass of the binary system. 
    Using Eqs.~\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}, an expression for the frequency and the gravitational-wave strain magnitude can be calculated~~\cite[e.g.,][]{Cutler1994}: 
      \begin{align}
      f &  \propto \mathcal{M}^{- 5/8}(t_c-t)^{- 3/8},\label{eq:Intro:chirp_f}\\
        % |h(t)| & \propto  \frac{\pi^{2/3}}{D}\left(\frac{576\mu^3 M^2}{5}\right)^{1/4}(t_c-t)^{-1/4}, \label{eq:Intro:chirp_ht_old} ,\\
        |h(t)| & \propto  \frac{\mathcal{M}^{5/4}}{D}(t_c-t)^{-1/4}, \label{eq:Intro:chirp_ht} 
        % Calculated by Cutler1994 (2.12) + (2.11) + Chirp mass defn
    \end{align}
    where $t$ is the time, $t_c$ is the time until coalescence, and $D$ is the luminosity distance to the source. 
    The combination of Eq.~\ref{eq:Intro:chirp_f}~and~\ref{eq:Intro:chirp_ht} result in the chirp waveform detected in the inspiral of compact object mergers.
    \par
    The first direct gravitational-wave detection from the inspiral of a binary neutron star merger, GW170817, was observed by a detector network consisting of the LIGO detectors at Hanford, Washington and Livingston, Louisiana in the USA, and the Virgo detector in Cascina, Italy~\cite{AdvancedLIGO2015,AdvancedVirgo2015,GW170817Detection}. 
    All three detectors consist of `L' shaped laser interferometers with free-falling test-masses.
    The gravitational-wave strain is found by measuring the optical path difference between the perpendicular beam tubes. \par

    Another binary neutron star (possibly neutron star - black hole) merger, GW190425, was detected in 2019~\cite{GW190425Detection,GWTC2,PopGWTC2}.
    %, along with another binary neutron star black hole binary with a high false alarm rate, GW190426\_152155.
    However, the luminosity distance of GW190425, $159^{+69}_{-72}\,$Mpc, and the broad sky localisation of the source, $8284\,\mathrm{deg}^2$, precluded any coincident electromagnetic observations~\cite{GW190425Detection}. In comparison, the sky localisation of GW170817 was $28\,\mathrm{deg}^2$ at a distance of $\sim 40\,\Mpc$.
    Therefore, to date, GW170817 is the only gravitational-wave signal that was observed with coincident electromagnetic radiation and will be discussed further in Section~\ref{sec:Intro:NeutronStarMergers}.
    %, allowing the host galaxy to be located~\cite{GW170817multi}.

    Although the LIGO and Virgo collaborations were able to detect part of the inspiral for GW170817, the sensitivity of the existing network was not sufficient to detect the gravitational-wave strain from either the late inspiral or the post-merger remnant.
    This is because the gravitational-wave strain emitted from these phases are at high frequencies, where the existing networks lack sensitivity.
    Figure~\ref{fig:Intro:ASDs} shows the amplitude spectral sensitivity at design sensitivity for LIGO Hanford in blue and Virgo in dashed orange.
    Funding has been secured to upgrade the LIGO network to A+ sensitivity, which will result in an improvement of $~\sim 2-3$~\cite{Miller2015, NEMO2020} and is shown with the green dotted curve.
    \par
\begin{figure}[H]
    \centering
    \includegraphics[width=0.95\textwidth]{{Thesis_ASDs}.pdf}
    \caption{Amplitude spectral sensitivity curves in $\Hz^{-1/2}$ for Virgo (orange, dashed) and Hanford (blue) at advanced design sensitivity, A+ (green, dotted), NEMO (red, dashed), Einstein Telescope (purple, dash-dotted) and Cosmic Explorer (black). The NEMO detector is $\sim 5$ times more sensitive than A+ at $\sim 2\,\kHz$ frequency.} 
        \label{fig:Intro:ASDs}
\end{figure}    
    A high-frequency detector has been proposed that will focus specifically on gravitational waves from the late inspiral and the post-merger remnant~\cite{NEMO2020}.
    This detector will be known as the Neutron star Extreme Matter Observatory (NEMO) and will serve two major goals.
    The first, as mentioned above, is gravitational-wave measurements of post-merger remnants and late time mergers.
    The second goal is to develop technology that will allow the third-generation observatories, Cosmic Explorer and Einstein Telescope, to be built \cite{PSD:CE,Punturo2010}.
    Figure~\ref{fig:Intro:ASDs} also shows the amplitude spectral density for NEMO (red, dashed), Einstein Telescope (Purple, dash-dot), and Cosmic Explorer (black).
    NEMO is $\sim 5$ times more sensitive than A+ at a frequency of $\sim 2\,\kHz$, resulting in a $\sim 10-15$ times improvement when compared to Advanced LIGO.

% ***************************************************************************************************************************************************************
    \section{Neutron stars}
    \label{sec:Intro:NeutronStars}
    Neutron stars are compact stellar remnants that are formed following supernovae of stars with masses of $\gtrsim 8\,\mathrm{M}_\odot$~\cite[e.g.,][]{Lattimer2004}.
    They can also be formed from the dynamics of binary star systems~\cite[e.g.,][]{Ruiter2019,LiuX2020,LiuD2020}.
    Neutron stars are created when the gravitational collapse of the inner stellar core overcomes the electron degeneracy pressure.
    The proto-neutron star matter is then supported against further collapse by neutron (or quark) degeneracy pressure. \par

    Proto-neutron stars cool below the Fermi energy at nuclear density, $\mathcal{E}_\mathrm{F}\sim 50\,$MeV, in the timescale of seconds ~\cite[e.g.,][]{Pons1999,Yakovlev2001,Paschalidis2012,Perego2019}.
    Such relatively low temperatures allow regions of superconductivity and superfluidity to form within the neutron star interior.
    Neutron stars are extremely dense objects with a central density $\sim 5\Hyphdash*10$ times nuclear density with magnetic fields of $\sim 10^8\Hyphdash*10^{15}\,$G~\cite[e.g.,][]{Reisenegger2001,Lattimer2004,Manchester2004}. % Core density Lattimer2004
    They consist of the most dense bulk matter  in the Universe.
     \par
    General-relativistic corrections to Newtonian physics are necessary to determine the relationship between the neutron star mass and radius.
    For non-rotating stars, this relationship calculated by~\cite{Tolman1939,Oppenheimer1939}
   \begin{align}
        \frac{dm}{dr} & = 4\pi \rho r^2,\label{eq:Intro:TOV:mass_gradient}\\
        \frac{dp}{dr} & = -\frac{p + \rho}{r(r- 2  m)}\left(4\pi p r^3 + m \right), \label{eq:Intro:TOV:pressure_gradient} 
    \end{align}
    where $r$ is the radial distance, $m$ is the mass enclosed up to $r$, $p$ is the pressure at $r$, and $\rho$ is the density at $r$. 
    Equations~\ref{eq:Intro:TOV:mass_gradient}\Hyphdash*\ref{eq:Intro:TOV:pressure_gradient}, in combination with the equation of state, allow the the mass and radius of neutron stars to be calculated, as well as the maximum non-rotating neutron star mass, $\mtov$, for the given equation of state.
    \begin{figure}[H]
        \centering
        \includegraphics[scale=0.4]{{NSMassRadiusDiagramHu2020Cropped}.png}
        \caption
        {Mass-radius diagram for several different equations of state (Figure from Ref.~\cite{Hu2020}). The horizontal axis shows the radius in km and the vertical axis shows the neutron star mass in $\msun$. The two most massive known neutron stars, J0348+0432~\cite{Antoniadis2013} and J0740+6620~\cite{Cromartie2020}, are shown as horizontal shaded regions.} 
        \label{fig:Intro:mass-radius}
        % Unless provided in the caption above, the following copyright applies to the content of this slide: © The Author(s) 2020. Published by Oxford University Press on behalf of The Royal Astronomical Society.This is an Open Access article distributed under the terms of the Creative Commons Attribution License (http://creativecommons.org/licenses/by/4.0/), which permits unrestricted reuse, distribution, and reproduction in any medium, provided the original work is properly cited.

\end{figure}
    Figure~\ref{fig:Intro:mass-radius} shows the mass-radius relationship for different prospective neutron star equations of state~\cite{Hu2020}. 
    The $\mtov$ for a given equation of state is indicated by the mass value corresponding to the upper-most end of the mass-radius curve.
    Any additional mass accumulated beyond $\mtov$ introduces a radial instability that leads to the collapse of the neutron star into a black hole.
    The heaviest neutron star  discovered to date is J0740+6620 with mass of $2.14^{+0.10}_{-0.09}\,\msun$ (Fig.~\ref{fig:Intro:mass-radius}, pale yellow region)~\cite{Cromartie2020}.
    Any equation of state that does not enter this region can be eliminated as non-physical. \par

    An equation of state can further classified by its dimensionless compactness, $C\equiv GM/Rc^2$, where $M$ and $R$ are the neutron star mass and radius, respectively.
    Large values of $C$, corresponding to soft equations of state, are shown with curves towards the left of Fig.~\ref{fig:Intro:mass-radius}, whereas stiffer equations of state, with smaller $C$, are located towards the right of Fig.~\ref{fig:Intro:mass-radius}.
    The black region in Fig.~\ref{fig:Intro:mass-radius}, where $C\geqslant 0.5$, is excluded to neutron stars as the Schwarzschild radius is larger than the prospective neutron star radius.\footnote{The Schwarzschild radius, $Rs = 2GM/c^2$, for a non-rotating black hole corresponds to $C=0.5$ which corresponds to a gradient on the Mass-radius diagram of $c^2/2G \approx 0.339 \msun/\mathrm{km}$.}
    % $M[\msun]/R[km] = c^2/2G\msun$
    The grey region is excluded by causality, where the speed of sound in the neutron star exceeds $c$.
    Given these constraints, the radius of neutron stars are $\gtrsim 10\,$km with $\mtov \gtrsim 2.2\, \msun$.

% \par

    % Another unique property of neutron stars is that they are opaque to neutrinos.

% ***************************************************************************************************************************************************************
    \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
    Binary neutron star mergers emit vast amounts of energy in gravitational waves.
    The estimated radiated energy for the GW170817 merger was $\gtrsim 5\times 10^{52}\,$ergs~\cite{GW170817Detection}, equivalent to $\sim 1\%$ of the total rest mass of the two neutron stars.
    For a source at a distance of $D$ from the observer, the amplitude of gravitational waves are  $\propto 1/D$, whereas the amplitude of electromagnetic waves are $\propto 1/D^2$.
    Additionally, electromagnetic waves can be attenuated by intervening media between the source and the detector whereas gravitational waves are transparent to the intervening media.
    Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
    % A sentence here about the lack of opacity to GW as opposed to EM?

    % If a BNS system of Rinit is less than tmerge is left undisturbed then the BNS can merge.
    % Formation channels.
    % If a binary neutron star system is formed with a separation less than the cr
    % From GW170817multi:
    % This motivated the development of accurate models for the two-body, general-relativistic dynamics (Blanchet et al. 1995;Buonanno & Damour 1999; Pretorius 2005; Baker et al. 2006; Campanelli et al. 2006; Blanchet 2014) that

    % Damour2012 in terms of tidal coupling constant, $\kappa_2^T$
    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.~\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}, can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
    \begin{align}
        h_{22}(f) & \equiv \frac{\mathcal{A}}{D}\mathcal{M}^{5/6}f^{-7/6} \exp(i\Psi(f)), \label{eq:Intro:StationaryPhaseApproxH22} \\
        \Psi(f) & = 2\pi f t_c - \phi_c - \frac{\pi}{4} +\frac{3}{4}\left(8\pi\mathcal{M}f\right)^{-5/3}, \label{eq:Intro:StationaryPhaseApproxPsi} 
    \end{align}
    where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
    Here, $f$ is the frequency~(Eq.~\ref{eq:Intro:FreqEvolution}), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system. Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane. Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
    Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
    Solving eqs.~\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution} for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
    \par
    % The base post-Newtonian expansion parameter is $\sqrt{M/r}$ but it is more convenient to use $\left(\pi M f\right)^{1/3}$ as the post-Newtonian expansion parameter.
    % These two parameters are equal up to and excluding terms of order $(M/r)^{3/2}$.
    The validity for these equations can be extended to times closer to the merger by adding relativistic corrections calculated by post-Newtonian expansion and effective-one-body dynamics~\cite{Cutler1994,Blanchet1995, Buonanno1999,Droz1999,Damour2012,Yagi2013,Blanchet2014}. % to Eq.~\ref{eq:Intro:StationaryPhaseApproxPsi}
    These corrections introduce dependence on the masses, spins, and tidal deformabilities of the individual neutron stars.
    This allows the measurement of these system properties in addition to the chirp mass, given a gravitational-wave inspiral detection.
    However, the accuracy of the individual neutron star properties is significantly less than the accuracy of the chirp mass because the relativistic corrections required to measure these properties are fairly weak. 
    This can be seen by looking at the parameters measured from GW170817.

    \par
    The best measured quantities of GW170817 are the chirp mass, $\mathcal{M}=1.186\pm 0.001\,\msun$, and the total mass, $M=2.73^{+0.04}_{-0.01}\,\msun$~\cite{GW170817Properties}.
    The measurements that rely more on relativistic corrections are less constrained: the component masses are $M_1 = 1.48\pm~0.12\,\msun$ and $M_2=1.26\pm~0.10\,\msun$, at 90\% credible intervals~\cite{GW170817Properties}.\footnote{These posteriors assume low-spin priors that are expected from Galactic neutron star binaries. Posteriors were also taken assuming more agnostic spin priors, see Ref.~\cite{GW170817Properties} for further details.}
    The mass-weighted quadrupolar tidal deformability was measured as $\tilde{\Lambda}=300^{+420}_{-230}$ at 90\% credible interval~\cite{GW170817Properties}.\par

    A coincident short gamma-ray burst was detected 1.7\,s after the inferred merger time for the GW170817 gravitational-wave event~\cite{Goldstein2017,GW170817Detection}. 
    A subsequent search found an optical candidate, SSS17a/AT2017gfo, near galaxy NGC~4993 at a distance of $\simeq 40\,$Mpc, within 11 hours of the merger~\cite{Coulter2017}.
    % i-band Swope      
    Other searches made detections from radio to X-rays~\cite[e.g.,][]{Troja2017,Nicholl2017,Chornock2017,Margutti2017,Alexander2017}.
    The first detectable X-rays were found with the Chandra X-ray Observatory nine days after the merger, which was consistent with an off-axis short gamma-ray burst~\cite{Troja2017}. \par

    The two favoured scenarios for the launch of GRB 170817A involve a central engine of a black hole with an accretion torus~\cite[e.g.,][]{Metzger2018,Gill2019,Murguia-Berthier2021}, or a central engine of a highly magnetised post-merger neutron star~\cite[e.g.,][]{Yu2018}.
    Early lanthanide-poor kilonova observations from AT2017gfo suggest that the short gamma-ray burst was launched following the collapse of the post-merger remnant at $\sim 0.1-1\,\s$ after the merger~\cite{Metzger2018}.
    Modelling the mass of the kilnova ejecta suggests that the post-merger remnant collapsed at $\sim 1\,\s$~\cite{Gill2019}.
    Jet structure together with afterglow observations were used to conclude that the post-merger remnant collapsed at $\sim 1-1.7\,\s$~\cite{Murguia-Berthier2021}.
    However, a stable neutron star remnant was found to be consistent with modelling of the AT2017gfo kilonova if the kilonova was powered mostly by radioactivity early on with additional energy from the neutron star remnant after $\sim 3\,$days~\cite{Yu2018}. \par

    Using gravitational waves to measure if and when the post-merger remnant of GW170817 collapsed to a black hole would greatly aid the modelling of GRB 170817A/AT2017gfo.
    Unfortunately, existing gravitational-wave detectors are not sensitive enough in the $\kHz$ range (see Fig.~\ref{fig:Intro:ASDs} and Sec.~\ref{sec:Intro:StrainAndSpectra}) and the collapse time is one of the most difficult parameters to predict with numerical-relativity simulations.

    % Metzger2018: Blue KN eject proposed cause: neutrino-irradiated, magnetically driver outflow from temporarily stable HMNS consistent with $\tc \sim 0.1-1\,\s$
    % Gill2019: Assume jet powered by rotating BH with $M_{\mathrm{ej,blue}} \approx 0.025\,\msun$ then $\tc \sim 1\,\s$
    % Murguia-Berthier2021: Assume jet powered by rotating BH: $\tc \approx1-1.7\,\s$
    % Yu2018: long lived stable NS

    % From Zhang2019:
    % There are two explanations for the total delay time of $\sim 1.7\,$s of GRB 170817A after the merger of GW170817~\cite{Zhang2019}.
    % The first involves a fireball where the emission is most likely from the photosphere radius. 
    % Assuming that the delay time from breakout to GRB emission is $\sim\,2\s$, similar to the duration of the GRB emission, the energy estimate for GRB 170817A suggests a Lorentz factor of $\sim 4$.
    % However, radio afterglow modelling suggests that the Lorentz factor should be $\gg 4$ for $t < 10\,$days after the merger. % Zhang2019[70]
    % This implies that the delay time from breakout to GRB emission is $\ll 2\,\s$ and that the total delay time is mostly attributed to the time to launch a jet or the time for breakout to occur. 
    % This scenario can still work if a post-merger remnant survives $\sim\,1\s$ before launching a jet.
    % However, this means that similarity between the total delay time of $\sim\,1.7\s$ and the GRB burst durations of $\sim\,2\s$ are coincidental as they are produced by different mechanisms. \par
    % The second explanation assumes that the majority of the total delay time is accounted by the time it takes from the breakout of the jet to the emission of the GRB~\cite{Zhang2018}.
    % This can be estimated by the width of the first pulse in GRB 170817A which was $\sim 1\,\s$ but could have been longer due to the weak nature of the burst.
    % This scenario does not depend on the type of post-merger remnant which may be either a neutron star or a black hole.

    % [Go through above references in the first blue paragraph]}

%     \begin{itemize}
%         \item tidal/inspiral methods
%         \item GW170817
%         \item other detections
%         \item Merger outcomes

%     \end{itemize}

% ***************************************************************************************************************************************************************
    \section{Numerical relativity simulations}

    Numerical-relativity simulations are required for determining the dynamics of the late inspiral of the binary neutron star and the post-merger remnant.
    To perform a numerical-relativity simulation, the equation of state must be defined, a set of initial conditions must be evaluated in general relativity, and then Einstein's field equations must be evolved.
    Each of these three requirements make these simulations difficult.
    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.

    \par
% Good overall references: Daverio2018, Dietrich2016

% ******************************************************************************************************************************************************    
    % Thesis: Maione2017
    % Two different conformal Z4 formulations have been studied: the CCZ4 formulation [Alic2012, Alic2013], which retains the full equations and in this way is a covariant formulation, at the price of greater differences respect to BSSN-OK and the need to introduce an additional parameter k3 to be able to evolve also spacetimes containing black holes and the Z4c [Gundlach2005, Bernuzzi2010, Hilditch2013] formulation, which retains only the principal part of the equations, in order to be as close as possible to BSSN-OK, but is not fully covariant and needs the introduction of ad-hoc constraint preserving boundary conditions in order to avoid constraint violations being reflected back inside the grid at the outer boundaries [Ruiz2011]. 
% ******************************************************************************************************************************************************    
    The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
    Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
    The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
    This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
    Quasiequilibrium configurations are also generated by \texttt{COCAL}.
    This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
    % Tsokaros2016: also KADATH [36], and SCRID [37,38]
    Once the initial conditions have been approximated then the space-time can be evolved.

    Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    %, and conformally flat (Isenberg-Wilson-Matthews)~\cite{Isenberg2008}. %Conformal and Covariant Z4 (CCZ4)~\cite{Bona2003,Bona2004,Alic2012,Alic2013},  \par
    The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
    This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
    Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this. \par

    To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}. % Dissipative hyperbolic driver from Alcubierre2001
    The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
    This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
    Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.  \par

    % The CCZ4 formalism is a fully convariant extension of the Z4 formalism with the exception of black hole simulations where covariance cannot be maintained. ~\cite{Alic2012}

    % Alic2013: The main advantage of the CCZ4 formulation over the BSSNOK formulation is its ability to dynamically evolve the constraints equations and to couple them with a constraint-damping scheme that leads to a rapid suppression of the violations that are inevitably produced in numerical simulations.
    % Takami2015: We also note that although we have recently developed a new code, i.e., WHISKYTHC, able to reach high-order convergence [21,22], we have not employed it here mostly because our focus is on the post-merger dynamics, where the large scale shocks reduce the accuracy of both codes to first order, and for which WHISKY appears to be slightly more robust.

    % Tsokaros2016: In this way we find that although global quantities of the system, like the mass and angular momentum, have differences of the order of ≲0.02%, local quantities, like rest-mass density, extrinsic curvature or angular velocity, show pointwise differences that are much larger, of the order of ≲1%. These local differences are then responsible for a ≲0.02%, local quantities, like rest-mass density, extrinsic curvature or angular velocity, show pointwise differences that are much larger, of the order of ≲1%. These local differences are then responsible for a dephasing in the gravitational waves at the merger time (after approximately three orbits) of ∼1.4 radians

    General-relativisitic hydrodynamic methods include codes such as \texttt{Whisky}~\cite{Baiotti2005}, \texttt{WhiskyTHC}~\cite{Baiotti2005,Radice2014,Radice2014b,Radice2015}, and \texttt{BAM}~\cite{Brugmann2004,Brugmann2008,Thierfelder2011,Dietrich2015,Bernuzzi2016b,Dietrich2019a}. 
    \texttt{WhiskyMHD} is a magneto-hydrodynamic extension to \texttt{Whisky}~\cite{Giacomazzo2007,Giacomazzo2011,Giacomazzo2013}, allowing for the introduction of magnetic fields.
    In this work we use gravitational-wave signals extracted from general-relativisitic hydrodynamic simulations from Refs.~\cite{Bernuzzi2014,Rezzolla2016,Radice2016,Dietrich2017b,Radice2017,Radice2017a,Radice2018,Dietrich2018}. 
    Refs.~\cite{Bernuzzi2014,Dietrich2017b} use the \texttt{BAM}  evolution code with the Z4c formulation, Refs.~\cite{Radice2016,Radice2017,Radice2017a,Radice2018} use the \texttt{WhiskyTHC} evolution code with the Z4c formulation, and Ref.~\cite{Rezzolla2016} uses the \texttt{Whisky} evolution code with the BSSNOK formulation. \par % Check metric evolution code for Rezzolla2016, have traced back to Baiotti2008 and it looks like BSSNOK

    The aforementioned tools allow the simulation of the late-time merger and post-merger remnant.
    However, there are a number of caveats that need to be considered.
    Firstly, the simulations are expensive, for example, 32 binary neutron star simulations took  $\sim 3\times 10^6\,$CPU hrs to complete~\cite{Takami2015}.
    Because of this, the direct use of numerical-relativity simulations in detection and parameter estimation, where we need large quantities of waveforms, is unfeasible. \par
    Secondly, numerical-relativity simulations unavoidably include numerical viscosity.
    Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010}.
    Numerical viscosity was found to introduce phase errors in the late-stage merger and improvements in numerical methods were subsequently introduced to minimise these errors~\cite{Radice2014}.
    The effects of numerical viscosity on the post-merger remnant is less clear, though it was found that numerical viscosity could alter both the neutrino luminosity~\cite{Sekiguchi2016} and the presence of one-armed spiral instability~\cite{Radice2016}.\par

    Thirdly, because of the complexity of the full general-relativity simulations, it is difficult to include all the necessary physics into simulations.
    Most simulations are performed with general-relativistic hydrodynamics, some also incorporate neutrino effects~\cite[e.g.,][]{Sekiguchi2011,Perego2014,Foucart2016,Radice2016a,Zappa2018}, viscous effects~\cite[e.g.,][]{Shapiro2000,Duez2004, Radice2017,Shibata2017}, and magnetic fields~\cite[e.g.,][]{Duez2006,Duez2006a,Siegel2013,Siegel2014,Giacomazzo2011,Kiuchia2012,Giacomazzo2013,Kiuchi2014,Giacomazzo2015}. \par

    Finally, simulations of the remnant are dependent on the spatial resolution.
    Generally, increasing the spatial resolution leads to more accurate simulations at the cost of simulation time.
    Ref.~\cite{Dietrich2018} state that their post-merger simulations from  the \texttt{CoRe} database are accurate enough to infer the energy and frequency content of the post-merger remnant.
    This is an important point: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.

    % Baiotti2017: A perhaps disappointing note that needs to be made, however, is that despite this abundance of numerical codes, systematic comparisons of them are seldom done, and the only cases we are aware of are the quantitative comparisons between the SACRA and Whisky codes [160, 490] and between the BAM and SpEC codes [110]; such comparisons are tedious and computationally expensive but important for gauging the systematic errors that the various codes can introduce.

% ***************************************************************************************************************************************************************
    \section{Post-merger remnants} \label{sec:Intro:PMR}
%     \begin{itemize}
%         \item Supramassive, Hypermassive properties
%         \item observational methods, 
%     \end{itemize}

    Neutron star mergers can result in four outcomes, depending on the total mass of the post-merger remnant and the maximum non-rotating neutron star mass, $\mtov$.
    The four different merger remnants in descending mass order are: black holes, hypermassive neutron stars, supramassive neutron stars, and stable neutron stars.
    With a total mass of $\gtrsim 1.5\,\mtov$, a black hole is promptly formed~\cite{Weih2018}.
    A hypermassive neutron star is formed when the mass is between approximately $1.2\,\mtov$ and $1.5\mtov$~\cite{Breu2016,Weih2018}.
    A supramassive neutron star is formed if the mass is between $1.0\,\mtov$ and $1.2\mtov$, and a stable neutron star is formed when the total mass is $\leq 1.0\,\mtov$.
    A remnant that has not collapsed into a black hole is referred to as a surviving remnant.\par

    Newly-formed post-merger remnants are differentially rotating unless they promptly collapse to a black hole within $\sim 2\,$ms.
    The differential rotation is quenched by gravitational-wave emission, magnetic braking, viscous forces, and neutrino cooling.
    If the remnant is a hypermassive neutron star then it will collapse into a black hole when the differential rotation is reduced below a critical value~\cite{Weih2018} and thermal support is lost~\cite[e.g.,][]{Bauswein2010,Kaplan2014}.  \par

    If the differentially-rotating remnant is a supramassive, or stable neutron star, then the remnant will evolve into a rigidly-rotating neutron star.
    A supramassive neutron star is supported against collapse by rigid rotation and will not collapse until the angular momentum drops below a critical value~\cite{Breu2016}.
    If the remnant has a mass corresponding to a stable neutron star, then it will remain stable for all levels of rigid rotation.\par

    Given a post-merger remnant mass, the deciding factor that discriminates between supramassive neutron stars and hypermassive neutron stars is $\mtov$.
    References~\cite{Margalit2017,Rezzolla2018,Khadkikar2021} found upper limits on $\mtov$ of $\lesssim 2.2\,\msun$ using numerical-relativity simulations with multi-messenger observations of GW170817~\cite{Margalit2017,Rezzolla2018}, and finite temperature modelling of matter~\cite{Khadkikar2021}. 
    Upper limits of $\lesssim 2.3\,\msun$ for $\mtov$ were found using angular momentum and energy considerations~\cite{Shibata2019}, and X-ray afterglow modelling~\cite{Sarin2020}, respectively. \par
    % Khadkikar2021 hot equation of state => wider uncertainties, relaxed constraints
    % Margalit2017: Multimessenger of GW170817 => M, R 
    % Rezzolla2018: Multimessenger of GW170817
    % Shibata2019: Angular momentum and energy conservation considerations
    % Sarin2020: X-ray afterglow
    Using $\mtov\approx 2.3\,\msun$ as a fiducial value, which is also consistent with the mass range of J0740+6620~\cite{Cromartie2020}, sets the threshold between supramassive and hypermassive neutron stars at $\approx 2.8\,\msun$. 
    This is very close to the total mass of the progenitor neutron stars for GW170817, $2.73^{+0.04}_{-0.01}\,\msun$~\cite{GW170817Properties},  which suggests that, on gravitational-wave data alone, the remnant of GW170817 could have been either a hypermassive neutron star or a supramassive neutron star.\par

    Most binary neutron star simulations have been performed with equal-mass, or near equal-mass, progenitors.
    Immediately after the merger, for a surviving remnant, the two inner cores of the progenitor neutron stars collide and bounce, while the surrounding material is differentially rotating~\cite{Takami2015}.
    Shock heating increases the temperature of the remnant up to $\sim 100\,$MeV~\cite{Perego2019}, exceeding the neutrino trapping temperature of 5-10\,MeV~\cite{Alford2018b}.   
    The remnant continues to differentially rotate as the cores merge in $\sim 2-5\,$ms.
    During the differential-rotation stage, the remnant is strongly emitting gravitational waves through f-mode oscillations (see Sec.~\ref{sec:Intro:StrainAndSpectra}).
 \par

    One picture for the evolution of differentially-rotating remnants can be drawn from Ref.~\cite{Kaplan2014}, where they show the influence of thermal pressure on equilibrium models for a $2.9\,\msun$ remnant with temperatures ranging up to 40\,MeV.
    The authors find a maximum core density where the remnant collapses which depends on the amount of differential rotation and is independent of temperature.
    They also show that the remnant reaches the critical density by losing gravitational mass while approximately conserving the baryonic mass.
    The gravitational mass is lost by: secular evolution with constant differential rotation, reduction in the amount of differential rotation, and loss of thermal support.
    % page 5 \Kaplan2014
    % Gravitational mass $m_g = m$ in Eq.~\ref{eq:Intro:TOV:mass_gradient}-\ref{eq:Intro:TOV:pressure_gradient} 
    % with the density $\rho$ in the same equations.
    % To calculate baryonic mass use $\frac{dM_b}{dr}$ = \frac{4\pi r^2 /rho}{1 + \epsilon}\left(1-\frac{2m_g}{r}\right)
    This additional thermal support extends the lifetime of the surviving post-merger remnant~\cite[e.g.,][]{Bauswein2010}.
 \par

    Reference~\cite{Paschalidis2012} estimated important timescales for the evolution of the post-merger remnant.
    The cooling timescale, estimated by assuming that neutrinos are trapped inside the post-merger remnant and escape by diffusion, is given by
    \begin{align}
        t_{\mathrm{cool}} & \approx 770 \left(\frac{M}{2.7\,\msun}\right)\left(\frac{R}{10\,\km}\right)^{-1}\left(\frac{E_\nu}{10\,\MeV}\right)^2\,\mathrm{ms}, \label{eq:Intro:NeutrinoCooling}        
    \end{align}    
    % 400 * (20/10) / (2.8/2.7) * (2.8/2.7)  *(M/2.8) (Ev/10)^2 / (R/20) = 400 (M/2.7) (Ev/10)^2 / ((R/20) * (20/10))
    where $E_\nu$ is the typical neutrino energy.
    We estimate a cooling time of $\sim 7$\,s by using a typical neutrino energy of $\sim 30\,\MeV$, found in numerical-relativity simulations that implement neutrino trapping~\cite{Sumiyoshi2020}, though this is reasonably uncertain due to the scaling with $E_\nu^2$.\par

    % Palenzuela2015 Ev~10 -20 MeV, Sumiyoshi2020 30 MeV
    The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
    \begin{align}
        t_{\mathrm{GW}} & \approx 150 \left(\frac{M}{2.7\,\msun}\right)^{-1}\left(\frac{R}{10\,\km}\right)^{-2}\left(\frac{f_2}{3.3\,\kHz} \right)^{-4}\left(\frac{e}{0.2} \right)^{-2}\,\mathrm{ms}, \label{eq:Intro:GWAngularMomentumLoss}        
    \end{align}
    where $f_2$ is the dominant post-merger oscillation frequency (see Sec.~\ref{sec:Intro:StrainAndSpectra}) and $e\equiv(a-b)/R$ is the ellipticity of the hypermassive neutron star. 
    Here, $a$ and $b$ are the semi-major and semi-minor axes of the neutron star, respectively, such that $R=(a+b)/2$.
    The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.\par
    % GRLES Simulations of Neutron Star Mergers: https://www.youtube.com/watch?v=Nppxi9E1ljg David Radice's youtube channel

    % In comparison, the emission of gravitational waves due to ellipticity and f-modes is approximated by:~\cite{Becker2009}
    % \begin{align}
    %     h_{\mathrm{GW}} \approx 6\times10^{-29} \left(\frac{M}{2.7\,\msun}\right)\left(\frac{R}{10\,\km}\right)^{2}\left(\frac{f_2}{3.3\,\kHz} \right)^2\left(\frac{d}{40\,\Mpc} \right)^{-1}\left(\frac{e}{0.2} \right), \label{eq:Intro:GWFModes}        
    % \end{align}
    % which has been scaled to a distance of $40\,\Mpc$, the distance of GW170817.
    Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
    % Rezzolla2010 long early simulation showing conservation of M_adm and J_gw
    Although Eq.~\ref{eq:Intro:GWAngularMomentumLoss} shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
    Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.

    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.
    % Andy comment on length

%     \begin{itemize}
%         \item Show Eos difference in GW strain
%         \item remnant astroseismology and frequency response
%         \item Quasi-universal relationships
%         \item Hot cold Eos
%     \end{itemize}

    \section{Gravitational-wave strain and spectra} \label{sec:Intro:StrainAndSpectra}

    Figure~\ref{fig:Intro:GWStrain} shows a typical gravitational-wave strain from a numerical-relativity simulation.
    The left panel shows the inspiral and post-merger components of the gravitational-wave strain for the plus polarisation, and the right panel shows a zoom of the post-merger component only.
    The numerical-relativity simulation uses the SLy equation of state with equal-mass $1.35\,\msun$ neutron stars ~\cite{Radice2016,Dietrich2018}.
    The gravitational-wave strain has a post-merger signal to noise ratio of 50 for a three detector network of LIGO Hanford, Livingston, and Virgo at advanced design sensitivity.
    The merger time is defined where $h_+(t)^2 +h_\times(t)^2$ reaches its first maximum.
    The inspiral gravitational wave $(t<0)$ shows a chirping sinusoid, increasing with frequency and amplitude until the neutron stars merge.
    The post-merger gravitational-wave strain $(t>0)$  has a much higher frequency than the inspiral component and has more structure in the frequency domain (see below).
    As mentioned in Sec.~\ref{sec:Intro:PMR}, the gravitational-wave strain in the first $\sim 2-5\,$ms, is more erratic due to  interactions of the two inner cores. 
    The interactions of the inner cores settle down after this time resulting in a somewhat monotonic f-mode oscillation in the gravitational-wave strain. 
    \par
\begin{figure}[H]
    \centering
    \begin{minipage}{0.49\textwidth}
        \centering
\includegraphics[width=\textwidth]{{Thesis_inspiral_SLy_NRonly_plus_time}.pdf}

    \end{minipage}\hfill
    \begin{minipage}{0.49\textwidth}
        \centering
        \includegraphics[width=\textwidth]{{Thesis_SLy_NRonly_plus_time}.pdf}
    \end{minipage}
        \caption{Plus polarisation of the gravitational-wave strain for inspiral and post-merger (left), and post-merger only (right). Generated from a numerical-relativity simulation using the SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016,Dietrich2018} for a post-merger signal-to-noise ratio of 50.} 
        \label{fig:Intro:GWStrain}
\end{figure}    
\begin{figure}[H]
    \centering
    \includegraphics[width=0.75\textwidth]{{Thesis_SLy_NRonly_plus_freq}.pdf}
    \caption{Frequency domain representation of the post-merger gravitational-wave strain shown on the right panel of Fig.~\ref{fig:Intro:GWStrain}. %The numerical-relativity simulation uses the SLy equation of state with equal-mass $1.35\,\msun$ neutron stars injected at a post-merger signal-to-noise ratio of 50. 
    Four main spectral peaks are visible in the gravitational-wave spectrum with the dominant peak, $\fpeak$, has the largest amplitude at $\sim 3300\,\Hz$. The sensitivity curves for Advanced LIGO and Advanced Virgo are shown as dashed black, and dotted black,  respectively.} 
        \label{fig:Intro:GWFreq}
\end{figure}

    The frequency response corresponding to the post-merger gravitational-wave strain in the right panel of Fig.~\ref{fig:Intro:GWStrain} is shown in Fig.~\ref{fig:Intro:GWFreq}.
    The amplitude is  $|\tilde{h}(f)|\sqrt{f}$, where $\tilde{h}(f)$ is the Fourier transform of the gravitational-wave strain. 
    The post-merger gravitational-wave spectrum is dominated by primary post-merger frequency, $\fpeak$, the main f\Hyphdash*mode oscillation of the remnant, which is $\sim 3.3\,\kHz$ for this simulation~\cite[e.g.,][]{Shibata1992, Zhuge1994,Stergioulas2011}. \par

    % fmode Shibata1992, Zhuge1994
    %fspiral, f_2_0, f_2+0, fpeak
        %  The peaks at frequencies $f_{2-0},f_{2+0}$ may result from coupling between a quasi-radial oscillation mode and $f_{\mathrm{peak}}$~\cite{Stergioulas2011}.
        % The peak at frequency $f_{\mathrm{spiral}}$ may result from the slower rotation-rate of tidally-deformed matter at the outer edges of the post-merger remnant~\cite{Bauswein2015}.
    Two additional oscillation modes were found in simulations that were caused by the interaction between the main f\Hyphdash*mode oscillation ($m=2$) and the $m=0$ radial pressure mode~\cite{Stergioulas2011}.
    % Reference~\cite{Stergioulas2011} proposed that the peaks above and below $\fpeak$ can be explained by the interaction between the dynamics associated with $\fpeak$ ($m=2$) and the $m=0$ radial pressure mode.
    This results in a modulated signal that should be present in the frequency response at frequencies $\fpeak \pm f_{m=0}$, designated as $\ftwominuszero$ and $\ftwopluszero$, respectively.
    %\par
    The $\ftwominuszero$ peak was found to be particularly strong in spectra for high-mass mergers as long as prompt collapse does not occur~\cite{Bauswein2015}.
    Another oscillation mode, labelled $\fspiral$, where $\ftwominuszero < \fspiral < \fpeak$~\cite{Bauswein2015}, was found in the spectra for low-mass mergers.
    This mode was generated by two antipodal bulges at the outer edge of the merger remnant that rotated at a slower frequency than $\fpeak$.
    The antipodal bulges were, in turn, caused by strong tidal deformation in the early post-merger stage.
    For higher-mass mergers that do not experience prompt collapse, the peaks associated with $\ftwominuszero$ dominate over peaks associated with $\fspiral$ in the spectra~\cite{Bauswein2015}.
    In the mass region between these extremes, both $\fspiral$ and $\ftwominuszero$ can be present in the spectra of these post-merger remnants. \par

    In a different interpretation of the post-merger spectra, the three main peaks were designated as $(f_1, f_2, f_3)$ in ascending frequency with $f_2 = \fpeak$~\cite{Takami2014,Takami2015,Rezzolla2016}.
    A quasi-universal relationship was found between $f_1$ and a number of progenitor system properties, including the compactness $C$ and the dimensionless quadrupolar tidal deformability $\Lambda$~\cite{Takami2015}.
    The spectral peak, $f_1$, was found to coincide with $\fspiral$ in spectrograms for stiff equations of state but was different otherwise~\cite{Rezzolla2016}.
    The power content associated with $\ftwominuszero$ was found to be extremely small in the spectrograms  even though $\ftwominuszero$ can easily be measured in simulations.
    References~\cite{Bauswein2019a,Vretinaris2019} support the interpretation of Ref.~\cite{Bauswein2015}, citing that a larger simulation set shows that $f_1$ is actually a combination of $\fspiral$ or $\ftwominuszero$ whereas a smaller simulation set, as in Refs.~\cite{Takami2014,Takami2015,Rezzolla2016}, is not large enough to highlight this.
\par

% \begin{figure}[H]
%     \centering
%     \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{{thesis_GW_spectrogram_5x7T_reduced_no_lines}.png}}%
%     \caption{Spectrograms (Figure from Ref.~\cite{Rezzolla2016}\protect\footnotemark) showing the frequency evolution of the post-merger remnant over time for equal-mass $1.35\,\msun$ neutron star mergers using five different equations of state at a distance of $50\,\Mpc$.  The amplitude of the spectrogram is indicated by the colour bar at the top of the image. The dominant post-merger oscillation frequency,  $\fpeak$, is shown as a dark-green solid line.} \label{fig:Intro:Rezzolla2016Spectrograms}
% \end{figure}
%     Figure~\ref{fig:Intro:Rezzolla2016Spectrograms} \footnotetext{This figure is a  subset of Fig.~4 from Ref.~\cite{Rezzolla2016}\label{foot:Intro:Rezzolla2016imageref}.} shows spectrograms for the numerical-relativity simulations of the post-merger gravitational-wave strain at $50\,\Mpc$ using five different equations of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Rezzolla2016}.\footnote{See footnote \footref{foot:Intro:Rezzolla2016imageref}.}
%     The equations of state progress from the stiffest, GNH3, in the left panel, to the softest, APR4, in the right panel.
%     The dominant post-merger frequency, $\fpeak$, is shown as a solid dark-green line. 
%     The magnitude of the spectrogram for a given frequency and time is indicated by the colour bar above the image.
%     In all panels except for APR4, there is some positive drift in the dominant post-merger frequency over time.
%     Furthermore, the spectrogram for GNH3 shows that there is also some frequency drift for the lower frequency peak at $\sim~1500\Hyphdash2000\,\Hz$.\footnote{Also note that the large frequency increase in the ALF2 spectrogram at $\sim 20\,\ms$ is due to the collapse of the remnant to a black hole.}

    Two models of the time-domain gravitational-wave strain were used for the basis for our model in Chapter~\ref{chapter:DetectionPE}.
    The first model used a third-order exponentially-damped sinusoid and successfully matched a simulation using DD2 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Bauswein2016}.
    The second model implemented a quadratic frequency-drift term over time for $\fpeak$ and used this model  to measure the neutron star radius~\cite{Bose2018}.
    Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
      \par

    With this in mind, we create a hybrid model of Refs.~\cite{Bauswein2016} and \cite{Bose2018} which we use in Chapters~\ref{chapter:DetectionPE} and~\ref{chapter:CollapseTime}. 
    We implement a linear frequency-drift term, as opposed to the quadratic frequency-drift term used in Ref.~\cite{Bose2018}.
    Similarly to Ref.~\cite{Bauswein2016}, we use an identical third-order system, to which we add the linear frequency-drift term.
    The model, outlined in full in Chapter~\ref{chapter:DetectionPE}, is $H w_j \exp(-t/T_j) \cos \left(2\pi f_j t\left[1+\alpha_j t\right]+\psi_j \right)$.
    The success of this model is indicated by fitting factors between posterior waveforms and the numerical-relativity simulation of $\gtrsim 0.92$ with most values $\sim 0.95$.
    \par

    A number of challenges are presented by the gravitational-wave strain generated with numerical-relativity simulations.
    Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
    % Bernuzzi2012: 2-4rad compared to post-Newtonian Taylor T4 at the time of merger, this is wrong it is the pN approximate that has error and this work is only interested up to contact.
    % Radice2014b: WhiskyTHC 0.6 rad between low and high resolutions
    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
    This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

    Another challenge presented by the gravitational-wave strain generated with numerical-relativity simulations is the collapse time of the remnant.
    As an example, consider the following simulations from the \texttt{CoRe} gravitational-wave database using the SLy equation of state~\cite{Dietrich2018,Radice2016}.
    These simulations are designated as THC:0036:R0$x$ for $x \in [1-4]$ with grid sizes of 148\,m, 222\,m, 295\,m, and 369\,m, respectively~\cite{Radice2016}.
    The finest resolution of 148\,m has a collapse time of $\sim 15\,\ms$, the remaining three simulations have collapse times of $\sim 17, \sim 24,$ and $\sim 60\,\ms$, respectively.
    Therefore, as the resolution becomes finer to reduce the phase error in the post-merger simulations, the amount of numerical-viscosity correspondingly increases,  resulting in an earlier collapse of the post-merger remnant.
    \par

    It may be the case that numerical-relativity simulations will not be fully calibrated to the true dynamics of post-merger remnants until we directly measure their gravitational waves.
    This is another motivation for simple models that are able to match the gravitational-wave strain from numerical-relativity simulations while remaining flexible to variations beyond the simulation outputs.
    In particular, models that can measure the collapse time of the remnant are important as numerical-relativity simulations are unable to predict this feature well. \par

%         \item Show Eos difference in GW strain
%         \item remnant astroseismology and frequency response

%  East2019a, Hotokezaka->2016PhRvD..93f4082H, Alic2013

    % ordered: Andersson1998b, Bauswein2012, Bauswein2012a, Yagi2013,Yagi2013a, Hotokezaka2013

    % others: Andersson1998b, Bauswein2012, Bernuzzi2014, Rezzolla2018, Iosif2020, Rezzolla2016, Bose2018, Sun2020

    % Bauswein2012,Bauswein2012a,Hotokezaka2013,Bernuzzi2014,Takami2014,Bernuzzi2015,Bauswein2015,Takami2015,Rezzolla2016,Bauswein2019

    % Andersson1998b: time constant for f-mode maps o compactness

    Quasi-universal relationships have been found relating the properties of the post-merger remnant and the progenitor neutron stars. 
    Perturbations in general relativity were used to find relationships between the frequency of neutron star f-modes and both the neutron star density and compactness~\cite{Andersson1998b}.
    The relationship between the f-mode frequency and the neutron star density was later refined using conformally-flat solutions of the Einstein field equations, and additional relationships were found between the neutron star radius and the f-mode frequency~\cite{Bauswein2012,Bauswein2012a}.
    Simulations of perturbed slowly-rotating neutron stars highlighted relationships between the neutron star moment of inertia, the spin-induced quadrupole moment, and the dimensionless tidal deformability~\cite{Yagi2013,Yagi2013a}. \par
%     \begin{figure}[H]
%     \centering
%         \includegraphics[scale=1.0]{{LambdaF2}.png}

%         \caption{Quasi-universal relationship between the dominant post-merger oscillation frequency, shown here as $f_2$, and the dimensionless quadrupolar tidal deformability, $\Lambda$, shown here as $(\lambda/\bar{M}^5)^{1/5}$~(Image Ref.~\cite{Takami2015}). Each colour shows a different equation of state.} 
%         \label{fig:Intro:LambdaFpeak}

% \end{figure}
    Relationships were then found between the progenitor neutron star ($l=2$) dimensionless quadrupolar tidal deformability, $\Lambda$, and a number of spectral features.
    The maximum gravitational wave frequency at the time of merger, $\fmax$, was closely related to $\Lambda^{1/5}$ for equal-mass neutron stars~\cite{Read2013}.
    A linear correlation was subsequently found between $\Lambda^{1/5}$ and the dominant post-merger f-mode frequency, $\fpeak$.
    We use this linear relationship in the model discussed in Chapter~\ref{chapter:ComputingFastReliable} to re-centre the gravitational-wave spectra after alignment.

    \section{Post-merger gravitational-wave models} 
    \label{sec:Intro:Models}
    In this section we briefly introduce the three main papers in Chapters~\ref{chapter:ComputingFastReliable}-\ref{chapter:CollapseTime} of this thesis. 
    A common theme in all three papers is that we need methods for generating fast binary neutron star post-merger waveforms for use in the detection or parameter estimation of post-merger remnants.
    In Chapter~\ref{chapter:ComputingFastReliable} we use a model that is trained on both the gravitational-wave spectra and the neutron star system parameters to generate post-merger spectra.
    In Chapter~\ref{chapter:DetectionPE} we develop a simple analytic model to generate time-domain gravitational-wave strain and test the model with Bayesian inference on numerical-relativity simulations.
    In Chapter~\ref{chapter:CollapseTime} we develop a model to measure the collapse time of post-merger remnants and find the distance at which we can detect the post-merger collapse for different gravitational-wave detector networks.

    In Chapter ~\ref{chapter:ComputingFastReliable}, we develop a way to generate gravitational-wave spectra in a fraction of a second.
    We use a hierarchical model that is trained on 35 numerical-relativity spectra from equal-mass neutron star post-merger simulations.
    The progenitor neutron star system is defined by the mass of the individual neutron stars, $M$, and the equation of state is defined by the compactness of the neutron stars, $C$, and the tidal coupling constant, $\kappa_2^T$.
    These three parameters, together with the gravitational-wave spectra, are sufficient to train our model. \par

    The model performance is tested with leave-one-out cross-validation and we measure noise-weighted fitting-factors of $0.95\pm 0.05$ between the inferred spectra and the spectra under-test.
    This can be compared to $0.93\pm 0.05$ which is the noise-weighted fitting-factor between the spectra of two identical neutron star systems using different numerical-relativity codes.
    We also show that we can use the trained model to perform parameter estimation for $\kappa_2^T$ with the noise-weighted fitting-factor as a proxy likelihood.

    In Chapter \ref{chapter:DetectionPE}, we use a third-order exponentially-damped sinusoid with a linear frequency-drift to model the post-merger time-domain gravitational-wave strain.
    We perform parameter estimation with this model to generate posterior waveforms for nine post-merger numerical-relativity simulations.
    The  noise-weighted fitting-factors between the numerical-relativity simulations and the posterior waveforms are $\sim 0.95$.
    The dominant post-merger oscillation frequency, $\fpeak$, can be constrained to $\pm_{1.2}^{1.4}\%$ for a post-merger signal-to-noise ratio of 15 and $\pm_{0.2}^{0.3}\%$ for post-merger signal-to-noise ratios of 50.
    We can constrain $\kappa_2^T$ at a post-merger signal-to-noise ratio of 15 to $\pm^{9}_{12}\%$, and $\pm 5\%$ at post-merger signal-to-noise ratios of 50,  using the hierarchical model. \par

    In Chapter~\ref{chapter:CollapseTime}, we extend the model from Chapter \ref{chapter:DetectionPE} to allow the collapse of the gravitational-wave strain.
    We use post-merger gravitational-wave strain from numerical-relativity simulations of equal-mass $1.35\,\msun$ neutron stars with LS220 and SLy equations of state.
    We then perform parameter estimation on the collapse time with four different gravitational-wave interferometer configurations. \par

    We find that we need a network of Einstein Telescope with Cosmic Explorer to detect a post-merger that collapses $\sim10\,\ms$ after the merger.
    Two A+ detectors at design sensitivity can measure a post-merger remnant that collapses $\sim\!10\,\mathrm{ms}$  after the merger to a distance of $\lesssim 10\,\Mpc$.
    If the proposed Neutron star Extreme Matter Observatory is operational, then this distance increases to $\sim\!18\Hyphdash*26\,\Mpc$ thereby increasing the effective volume and hence detection rate by a factor of $\sim 30$.

%      and \ref{chapter:CollapseTime} use Bayesian inference in a gravitational-wave framework using the \texttt{Bilby} package and \texttt{Dynesty} nested-sampler~\cite{Ashton2019,Speagle2019}.
%     The base model for both of these chapters is a third-order exponentially-damped sinusoid with a frequency-drift term, Eqs.~\ref{eq:htotal2}-\ref{eq:hmode2}.
%     In Chapter~\ref{chapter:DetectionPE}, we use this model for detection and parameter estimation of post-merger gravitational-wave strain.
%     One major difference between this model and the hierarchical model is that the former is able to model the time domain of the gravitational-wave strain whereas the latter is only able to model the amplitude of the gravitational-wave spectra.
%     In Chapter~\ref{chapter:CollapseTime}, we extend the model from Chapter~\ref{chapter:DetectionPE} to allow the measurement of the collapse time of the gravitational-wave strain.

%     In Chapter~\ref{chapter:DetectionPE}, we use a $0.2\,\ms$ rising-edge Tukey window applied to the time-domain post-merger gravitational-wave strain to reduce artefacts due to Gibbs phenomenon.
%     Gibbs phenomenon can cause unwanted oscillations in the time domain due to discontinuities in the signal amplitude~\cite{Gibbs1899}.
%     The window rise-time is kept small so that we can measure the performance of the model to the coalescence time (Figs.~\ref{fig:OF_MF_SNR_FF5ms}-\ref{fig:TCOALESCENCE}) without unduly filtering out early post-merger signals.

%     As each of the three modes of the base model are analytically identical, sampling of the model can be difficult.
%     This is because there are six permutations of model parameters for every set of parameters.
%     For example, if the posterior has support around frequencies $(f_0,f_1,f_2)=(1200, 1500, 3300)$ then all six permutations of these frequencies are also valid solutions for the posterior, including $(f_0,f_1,f_2)=(3300, 1200, 1500)$.
%     This means that the posterior volume is effectively six times larger, making it much more difficult to sample. \par

%     In Chapter~\ref{chapter:DetectionPE}, we find that we can remove the redundant permutations by ordering the maximum spectral amplitudes for each mode.
%     The maximum spectral amplitude for the $j$th mode is given by $H_j = \max(|\tilde{h}_j(f)|\sqrt{f})$ and we enforce a descending order such that $H_0>H_1>H_2$  (see also Eq.~\ref{eq:modesorting2}). 
%     This ensures that the dominant post-merger f-mode peak, $\fpeak$, corresponds to mode zero (ie. $\fpeak=f_0$).
%     We implement the mode sorting with constrained priors in \texttt{Bilby}.
%     To find the posteriors, we use \texttt{Bilby} and \texttt{Dynesty} to perform nested sampling with 750 live points.
%     If the number of live points is too small, then the sampler will loose valid parts of the posterior volume\footnote{If there are not enough live points within a particular mode then all live points can leave this mode by random chance. This is called mode `die-off'. The simplest solution to this is to increase the total number of live points.}, whereas an excessively large number of live points will take too long for the sampler to complete.
%  \par

%     Adding a collapse-time parameter to the post-merger gravitational-wave model makes convergence of the sampler more difficult.
%     This is because the signal amplitude can be quite small at the time of collapse and the dimensionality of the system is increased by one.\footnote{The dimensionality would actually be increased by two, as we introduce two parameters: $\tc$ and $\tw$. For convergence reasons we fix $\tw=2\,\ms$. See Chapter~\ref{chapter:CollapseTime} for further information on this.}
%     In response to this, in Chapter~\ref{chapter:CollapseTime}, we find that a number of additional measures are required to aid convergence. 
%     Firstly, we sort the mode frequencies  in descending order, $f_0 > f_1 > f_2$, (Eq.~\ref{eq:freqsorting3}) in addition to the mode amplitude sorting used in Chapter~\ref{chapter:DetectionPE}.
%     This limits our posteriors to spectral peaks at frequencies $\lesssim \fpeak=f_0$.
%     Secondly, we implement a $2\,\ms$ rising-edge Tukey window to reduce sensitivity to the gravitational-wave strain for $\lesssim 2\,\ms$.
%     Finally, we double the number of live points to 1500 to aid convergence of the sampler. \par

%     However, to aid performance we implement mode sorting by  overriding the \texttt{log\_likelihood\_ratio} method of  \texttt{bilby.gw.likelihood.GravitationalWaveTransient}.
%     When the maximum spectral amplitudes are not in descending order, the log likelihood method returns \texttt{-1e200}, effectively removing that point from the posterior. 
%     This is also implemented by  overriding the \texttt{log\_likelihood\_ratio} method of  \texttt{bilby.gw.likelihood.GravitationalWaveTransient}.
%     If the frequencies are not arranged in descending order then a log likelihood value of \texttt{-1e200} is returned and the spectral amplitudes are not calculated.
% \par

%     In Chapters~\ref{chapter:DetectionPE} and ~\ref{chapter:CollapseTime}, the total duration of the signal segment is $150\,\ms$ with a sampling frequency of $16384\,\Hz$.
%     Furthermore, we use a rising-edge Tukey window beginning at the merger time, $t\equiv 0$ to reduce artefacts due to Gibbs phenomenon. 
%     In Chapter~\ref{chapter:DetectionPE}, we use a $0.2\,\ms$ rising-edge Tukey window so that we can measure the performance of the model to the coalescence time (Figs.~\ref{fig:OF_MF_SNR_FF5ms}-\ref{fig:TCOALESCENCE}) without unduly filtering out early post-merger signals.
%     For~\ref{chapter:CollapseTime}, as we are interested in measuring the collapse time for longer signals, $\gtrsim 5\,\ms$, we use a longer $2\,\ms$ rising-edge Tukey window.
%     The parameters we use for \texttt{Dynesty}/\texttt{Bilby} in Chapter~\ref{chapter:DetectionPE} are:  $\texttt{npoints} = 750$, $\texttt{walks} = 150$.
%     Whereas it was necessary to use slower, more robust parameters for Chapter~\ref{chapter:CollapseTime}: $\texttt{npoints} = 1500$, $\texttt{walks} = 150$, $\texttt{nact} = 10$, $\texttt{maxmcmc} = 15000$.

 \par

    % We encounter problems with initial analytic injections in Chapter~\ref{chapter:CollapseTime} for collapse time runs.
    % When we perform injections with all parameters fixed except for the collapse time and the exponential decay time-constant, with no injected noise and a small signal-to-noise ratio, we find that the collapse time posterior is biased\footnote{For zero noise injections, the gravitational-wave likelihood is still weighted by the power-spectral-density even though no Gaussian noise is added to the signal. This is why the signal-to-noise ratio has meaning for zero noise injections.}.
    % We also find that the two-dimensional posteriors are not biased and that the bias is introduced when the one-dimensional posteriors are calculated.
    % We also found that we could not simultaneously recover both $\tc$ and $\tw$ even in loud signals with zero injected noise and found that using a $\tw=2\,\ms$ was a good compromise allowing successful recovery for the collapse time.

%     \begin{itemize}
%         \item Machine learning
%         \item Validation
%         \item Bayesian inference
%         \item nested sampling         
%     \end{itemize}

\end{document}

### 2021-10-11 papers/paper1.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}

% \documentclass[
% % preprint,
% % prx,
% % prl,
% % twocolumn,
% %onecolumn,%!TEX encoding = UTF-8 Unicode
% nofootinbib,
% % ,secularistic%
% tightenlines %
% floatfix,
% showpacs ,amssymb, aps,% linenumbers
% superscriptaddress
% ]{revtex4-1}
% \documentclass[../main.tex]{subfiles}
% \usepackage{amsmath,slashed}
% \usepackage[left]{lineno}

% \usepackage{comment}
% \usepackage{graphicx}
% \usepackage{amsmath,amssymb}
% \usepackage{bm}
% \usepackage[shortlabels]{enumitem}
% \usepackage{float,soul}
% \usepackage{natbib}
% \usepackage{wasysym}
% \usepackage{aas_macros}
% \graphicspath{{images/}}
% \RequirePackage{lineno}
% \usepackage[usenames,dvipsnames]{color}

% % \newcommand{\red}[1]{\textcolor{red}{#1}}
% % \newcommand{\pdl}[1]{\textcolor{red}{[PDL: #1]}}
% % \newcommand{\pje}[1]{\textcolor{Purple}{[PJE: #1]}}
% % \newcommand{\blue}[1]{\textcolor{blue}{#1}}
% \DeclareMathOperator*{\argmin}{arg\,min}
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Vic 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}

\begin{document}

\begingroup
\clearpage% Manually insert \clearpage
\let\clearpage\relax% Remove \clearpage functionality
\vspace*{-2cm}% Insert needed vertical retraction

\chapter{Computing Fast and Reliable Gravitational Waveforms of Binary Neutron Star Merger Remnants}
\label{chapter:ComputingFastReliable}
\endgroup

\chapterauthor{Paul J. Easter \\ Paul D. Lasky\\Andrew R. Casey\\Luciano Rezzolla\\Kentaro Takami}
% \email{paul.easter@monash.edu}
% \author{Paul D. Lasky}
% \email{paul.lasky@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Andrew R. Casey}
% \email{andrew.casey@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \affiliation{Faculty of Information Technology, Monash University, VIC 3800, Australia}
% \author{Luciano Rezzolla}
% \affiliation{Institut f\"ur Theoretische Physik, Max-von-Laue-Stra\ss e 1, 60438 Frankfurt, Germany}
% \author{Kentaro Takami}
% \affiliation{Kobe City College of Technology, 651-2194 Kobe, Japan}

\section*{Abstract}
	Gravitational waves have been detected from the inspiral of a binary neutron-star, GW170817, which allowed constraints to be placed on the neutron star equation of state. The equation of state can be further constrained if gravitational waves from a post-merger remnant are detected. Post-merger waveforms are currently generated by numerical-relativity simulations, which are computationally expensive. Here we introduce a hierarchical model trained on numerical-relativity simulations, which can generate reliable post-merger spectra in a fraction of a second. Our spectra have mean fitting factors of 0.95, which compares to a fitting factor of 0.93 between different numerical-relativity codes that simulate the same physical system. This method is the first step towards generating large template banks of spectra for use in post-merger detection and parameter estimation.

\section{Introduction}
    Gravitational waves have been  observed from the inspiral of binary neutron star merger GW170817~\cite{GW170817Detection}. This allowed limits to be placed on the neutron star tidal deformability ~(see~e.g.,~\cite{GW170817Detection,Annala2018,Radice2018,Most2018,De2018,GW170817Properties}). However, due to lack of detector sensitivity at high frequencies, the merger and post-merger signals were not detected~\cite{GW170817Properties,GW170817Postmerger1,GW170817Postmerger2}. Post-merger gravitational-waves from a binary neutron star merger could be detected with a signal-to-noise ratio of 5 at a distance of $\sim$25-40\,Mpc with Advanced LIGO at design sensitivity~\cite{Takami2014}.  The physics of the post-merger remnant is of particular interest as it probes the neutron star equation of state at significantly higher temperatures than the progenitor stars.\par 
    The detection and characterisation of a post-merger remnant is aided by a large bank of gravitational-wave strain waveforms. Generating such waveforms is currently computationally expensive, and there are only of order 100 in existence. In this work we make a step towards generating a large template bank of post-merger spectra by training a hierarchical model on a set of numerical-relativity spectra.\par    
    There has been significant research applied to the relationship between post-merger numerical-relativity simulations, the corresponding spectrum of the gravitational-wave strain, and the neutron star equation of state~(e.g.,~\cite{Stergioulas2011,Bauswein2012,Read2013,Hotokezaka2013,Takami2014,Bernuzzi2015,Takami2015,Clark2016postmerger,Rezzolla2016,Radice2017,Dietrich2017b,Radice2017a,Pietri2018,Dietrich2018,Dietrich2018a,Bose2018}). There are many degrees of freedom for each simulation, which include the neutron star system parameters, equation of state, and  simulation parameters (e.g., spacetime evolution formalism~\cite{Baiotti2017}, resolution), as well as parameters related to magnetic fields and neutrinos. We choose to use a set of numerical-relativity simulations that are homogeneous, eliminating unwanted variations between simulations with different parameters. To achieve this, we use a subset of 35 waveforms from Ref.~\cite{Rezzolla2016} consisting of identical simulation parameters with variations in the neutron star mass and equation of state only. To obtain consistent spectra, we select waveforms that have a fixed time-step and almost identical length. \par
    Ref.~\cite{Clark2016postmerger} showed that dimensional reduction of post-merger waveforms is possible by performing principal component analysis after aligning the maximum of each gravitational-wave strain spectra in the frequency domain (see also \cite{Bose2018}). We use a similar method of frequency shifting in our model. We introduce a hierarchical model that trains on existing numerical-relativity post-merger simulations, and can produce new, accurate spectra in a fraction of a second. This is the first step towards making large template banks of post-merger spectra suitable for detection and parameter estimation which could compliment existing unmodelled searches for post-merger remnants~\cite{GW170817Properties,Klimenko2016,Chatziioannou2017}. \par
\begin{figure*}
    \includegraphics[width=0.94\textwidth]{NewAllMatches35withCoRe.pdf} 
    \caption{ Reconstructed gravitational-wave spectra generated with leave-one-out cross-validation (solid red) and original numerical-relativity spectra~\cite{Rezzolla2016} (dashed black), scaled to a distance of 50 Mpc. Each column represents a different equation of state and each row represents a different neutron star mass, increasing towards the bottom. The one-sigma uncertainty in the spectra is shaded in light red for each prediction.  The Advanced LIGO noise amplitude spectral density (dotted black curve)~\cite{PSD:aLIGO} is shown on all subplots. A numerical-relativity spectrum generated from~\cite{Dietrich2018} is shown (dashed dot blue curve) in the last row (equal mass 1.35\,M$_\odot$) for SLy~\cite{Radice2016} (fourth column) equation of state.}
    \label{fig:GoodPlotsWithWaveformGeneration}
\end{figure*} 
    Simulation of the post-merger phase of binary neutron star mergers is significantly more complicated than the inspiral phase due to the complex physics including shock heating and nonlinear mode coupling. Additional effects, such as neutrino cooling and magnetic fields, are not expected to yield substantial modifications to the locations of the spectral peaks (see e.g. \cite{Giacomazzo2011,Sekiguchi2016,Kawamura2016}), while the role of viscous effects is still a matter of debate \cite{Alford2018}. The accuracy of the resulting simulations can be limited by the trade off between computational constraints and higher resolutions~\cite{Baiotti2017}. This is particularly true for the phase evolution of the post-merger simulations which do not necessarily converge~\cite{Dietrich2018}. However, the power-spectral content is convergent for sufficiently high resolutions (e.g.,~\cite{Takami2015,Dietrich2018}). Our model is representative up to the validity of the numerical-relativity simulations that it is based upon. With this in mind, we wish to encourage further research into numerical-relativity simulations of post-merger remnants to increase the available number of waveforms and to enable further cross-checking between codes.\par
\section{Methodology}
    \sloppy We use 35 numerical-relativity simulations of binary neutron star mergers from~Ref.\cite{Rezzolla2016}, to which we refer to for details on the equations of state employed.  Each simulation consists of non-spinning, equal-mass progenitor neutron stars, with five different equations of state across the simulations.  We train our model on the amplitude  of the characteristic strain spectra, $h_c(f)~=~|\tilde{h}(f)|\sqrt{f}$. Here,  $\tilde{h}(f)$ is the Fourier transform of the plus polarisation of the post-merger gravitational-wave strain, $h_+(t>0)$. The plus and cross polarisations of the simulated gravitational-wave strain have almost identical spectral amplitude and have a phase offset of almost exactly $\pi/2$. We gain no extra information by including the cross polarisation. The merger time, $t=0$, is defined as the time where $h_+^2(t)~+~h_\times^2(t)$ reaches the first maximum. \par

    We use a hierarchical model to represent the amplitude spectra. Given a neutron star of mass $M$ and radius $R$, we assume the compactness of a neutron star, $C\equiv M/R$, in the $j$th simulation has a power-law dependence with the mass, $M$, over all equations of state
    \begin{align}
        C_j = \alpha_j M_j^{\beta_j}. 
        \label{eq:Ccalc1}
    \end{align}
        The validity of this model will be determined by how well we can match the numerical-relativity waveforms. The hyperparameters $\{\mathbf{a},\mathbf{b}\}$, and the quadrupolar tidal coupling constant, $\kappa_2^\tau$, determine the values of $\{\alpha,\beta\}$ as follows:
    \begin{align}
        \alpha_j & \sim\mathcal{N}(a_0 + a_1 \kappa_{2,j}^\tau,\sigma_\alpha^2),\label{eq:alphacalc1} \\ 
        \beta_j & \sim\mathcal{N}(b_0 + b_1 \kappa_{2,j}^\tau,\sigma_\beta^2),
        \label{eq:betacalc1}
    \end{align}
        where $\mathcal{N}(\mu,\sigma^2)$ is a Gaussian distribution of mean $\mu$ and variance $\sigma^2$. The quadrupolar tidal coupling constant, $\kappa_2^\tau$, is used due to its importance in the inspiral dynamics~\cite{Damour2010,Read2013,Takami2014,Bernuzzi2015,Takami2015,Rezzolla2016} and its correlation with the location of the main frequency peak of the post-merger spectrum~\cite{Bauswein2012,Takami2014,Bernuzzi2015, Takami2015}.\par
        All spectra in the training set, which excludes the spectrum under test when leave-one-out cross-validation is performed, are used to determine the hyperparameters $\{\mathbf{a},\mathbf{b}\}$ by a least squares fit. The amplitudes for each spectrum are frequency shifted so that the peak frequencies are aligned in a similar way to Refs. \cite{Clark2016postmerger,Bose2018}.  We then fit the aligned spectral amplitudes with a linear model
    \begin{align}
        (h_c)_{i,j} = \boldsymbol{\Theta}_{i}\mathbf{X}_j + \text{noise}, 
        \label{eq:GenerateWaveform1}
    \end{align}
        where the noise is modelled as intrinsic variance, $s_i^2$ for the $i$th frequency bin, $\boldsymbol{\Theta}_i$ is a vector of unknown coefficients,  and $\mathbf{X}_j$ is a design matrix of 
    \begin{align}
        \mathbf{X}_j=[1,\widehat{C}(M_j,\kappa^\tau_{2,j}),\widehat{M}_j,\widehat{\kappa^\tau_{2,}}_{j}]. \label{eq:DesignMatrix1}
    \end{align}
       \noindent The hats indicate the whitened transformations of the neutron star parameters such that $\widehat{\mathbf{x}}\sim~ \mathcal{N}(0,1)=( \mathbf{x}-\mu_x)/\sigma_x$ where $\mu_x$ and $\sigma_x^2$ are the mean and variance of $\mathbf{x}$ respectively. The compactness parameter can be generated from Eq.~\ref{eq:Ccalc1}-\ref{eq:betacalc1} after determining the values for $a_0, a_1, b_0$, and $b_1$. Spectra can then be trivially generated given any mass, quadrupolar tidal coupling constant and frequency shift. The frequency shift can be determined from the value of the quadrupolar tidal coupling constant~\cite{Bernuzzi2015,Takami2015}. \par
       We perform leave-one-out cross-validation to test the performance of the model. We do this by excluding the spectrum under test and its associated parameters from the training set. In doing so, the spectra generated during leave-one-out cross-validation represent an extrapolation by the model  and the fitting factors are therefore conservative.
       \par
    \begin{figure}[H]
          \includegraphics[width=0.99\columnwidth]{newFittingFactorHistogram35.pdf}  
          \caption{Histogram of fitting factors determined by comparing numerical-relativity spectra with spectra generated by our model using leave-one-out cross-validation.}
          \label{fig:FittingFactorsHistogram}
    \end{figure}   
    \begin{figure*}
        \includegraphics[width=\textwidth]{NewGeneratedWaveforms35.pdf} 
        \caption{ Spectra generated by the model when trained on all the numerical-relativity spectra (red). The uncertainties in the spectra are shown in light red. The Advanced LIGO noise amplitude spectral density (dotted black) is shown on all subplots.}
        \label{fig:GeneratedWaveforms}
    \end{figure*}      
       We perform spectral comparisons using the following noise-weighted fitting factor, or overlap~\cite{Apostolatos95}
        \begin{equation}
        	FF(h_{1},h_{2})\equiv\frac{\left<h_{1}|h_{2}\right>}{\sqrt{\left<h_{1}|h_{1}\right>\left<h_{2}|h_{2}\right>}}.\label{eq:FF1}
        \end{equation}
        Here, the inner product is defined by
        \begin{align}
        	\left<h_1|h_2\right>\equiv4 \int df\frac{|\tilde{h}_1(f)|\ |\tilde{h}_2(f)|}{S_h(f)},
        \end{align}
        where $S_h$ is the noise power spectral  density. Throughout this article we use the \texttt{ZERO\char`_DET\char`_high\char`_P} file from~\cite{PSD:aLIGO} to determine the amplitude of $S_h$. The resultant fitting factor is similar to the standard fitting factor except that it operates on the Fourier amplitude only. A fitting factor of one represents a perfect match. The fitting factor represents the loss incurred to the  optimal signal-to-noise ratio due to mismatch in the model spectra, where the optimal signal-to-noise ratio is given by $\rho_{opt}=\sqrt{	\left<h|h\right>}$.\par
        While it is known that smooth relationships exist between various system properties (e.g., mass, tidal parameters, etc.) and post-merger waveform spectral features~\cite{Bauswein2012,Takami2014,Bernuzzi2015,Takami2015,Bauswein2015,Rezzolla2016}, no such relationships exist for the phase information~(see also~\cite{Messenger2014}). Empirically, while we find good training fits using our model on the spectral content of the waveforms (see below), we are not able to train confidently on the full time series including both phase and amplitude as the phase evolves too quickly between adjacent simulations.  We discuss this in more detail below. \par

    \section{Results}
        Following the training step, we use Eq.~(\ref{eq:GenerateWaveform1}) to generate spectra.  Figure~\ref{fig:GoodPlotsWithWaveformGeneration} shows how well our generated spectra match the original numerical-relativity spectra. The original spectra are shown as dashed black curves, the cross-validation spectra are shown as red curves, and the one-sigma model uncertainty is shown in shaded light red. All spectra are scaled to a distance of 50 Mpc. The Advanced LIGO noise amplitude spectral density is shown as the dotted black curves~\cite{PSD:aLIGO}.  We fit the large-scale structure of the numerical-relativity peaks well with some deviations in the small-scale structure. \par Figure~\ref{fig:FittingFactorsHistogram} shows a histogram of the noise-weighted fitting factor, Eq.~(\ref{eq:FF1}), between our cross-validated model prediction and the corresponding numerical-relativity spectra for all tested waveforms. The resulting histogram has a mean of 0.95 with a standard deviation of 0.03. \par
        To place the above results in context, we calculate the nearest neighbour fitting factor of the numerical-relativity spectra. We measure the fitting factor for the spectrum under test against all other spectra, and report the largest fitting factor. We obtain nearest neighbour fitting factors for the numerical-relativity spectra of 0.93 with a standard deviation of 0.05. The fitting factors generated by our model compare favourably with this result. Additionally, our model is capable of generating spectra given the required input parameters, whereas a nearest neighbour interpolation would not be capable of this.\par

        As a additional baseline value for comparison, we compare fitting factors between the numerical-relativity spectra used in this paper~\cite{Rezzolla2016}, and those produced with other codes~\cite{Dietrich2018}. Notwithstanding the fact that post-merger waveforms can differ with resolution even when using the same code, we assume that the waveforms have similar truncation errors and compare one set of spectra using equations of state SLy~\cite[THC:0036:R03]{Radice2016} for equal-mass binaries with $M=1.35\,M_\odot$.

        The spectrum  for the comparison waveform is plotted (blue dashed dot curve) in the last row and fourth column of Fig.~\ref{fig:GoodPlotsWithWaveformGeneration}, corresponding to the SLy equation of state. This spectrum can be compared to the black dashed waveform from~\cite{Rezzolla2016} in the same panel. While amplitude offsets do not change the fitting factor, differences in the frequency and the shape of the peaks do. The fitting factor between these two waveforms is 0.93. This can be attributed to the difference in the relative shapes of the two main peaks between the simulations in~\cite{Rezzolla2016} and~\cite{Radice2016}.\par
        This comparison indicates that our fitting factors are comparable to the fitting factors between different numerical-relativity codes. We note that numerical simulations are indeed accurate enough for understanding the main structures (e.g., positions of dominant peaks) and their relationship with bulk properties of the remnant. However, our method is limited by the accuracy of the numerical-relativity simulations, which are in turn limited by computational capabilities; we discuss the implications of this below.  \par
        To evaluate how the generated spectra vary, we train on all numerical-relativity spectra and generate a grid of model spectra. We generate spectra at five equally spaced mass and quadrupolar tidal coupling constant values. The mass ranges from $M=1.25\,M_\odot$ to 1.35\,$M_\odot$, and  the quadrupolar tidal coupling constant varies from $\kappa_2^\tau = 50$ to 350. The generated spectra are shown in Fig.~\ref{fig:GeneratedWaveforms} as the red curves,  the one-sigma model uncertainty as light red shading, and the Advanced LIGO noise curve as dotted black. Each of these spectra take a fraction of a second to evaluate. We show these spectra to indicate what we can hope to achieve by implementing these models in full parameter estimation. \par

        In Fig.~\ref{fig:GeneratedFittingFactor} we compare the fitting factor between a spectrum generated with $M=$~1.3\,$M_\odot$, $\kappa_2^\tau=$~100 against spectra generated at other parameter values. We choose $\kappa_2^\tau=$~100 (for an equal-mass system this corresponds to $\Lambda=530$) to be consistent with tidal deformability values, $\Lambda$, determined in \cite{GW170817Detection,Annala2018,Radice2018,Most2018,De2018,GW170817Properties}, under the simplified assumption that $\kappa_2^\tau\approx \frac{3}{16}\Lambda$ (noting that the equality holds for equal-mass progenitors). The location of the reference spectrum is shown with the black cross. This provides the first indication of whether this model could recover the mass and quadrupolar tidal coupling constant when trained on sufficient numerical-relativity simulations. The peak of the contour plot around the reference waveform shows that this model is selective, and may be used for parameter estimation and/or detection in the future. However, this is a task for future work and will require full Bayesian analysis with a noise implementation.
 \begin{figure}[H]
 \begin{center}
    \includegraphics[width=8cm]{NewGeneratedFittingFactor90plus35Kappa100.pdf}
        \caption{Fitting factor between spectra generated using our hierarchical model with 1.30\,$M_{\odot}$ and $\kappa_2^\tau=$100 (black cross), against a grid of mass and quadrupolar tidal coupling constant values.}
        \label{fig:GeneratedFittingFactor}

 \end{center}
\end{figure} 
        It is possible that a phase transition may occur in the remnant during the post-merger phase due to an increase in density~(e.g.~\cite{Most2018,Bauswein2018}). This could be detected by comparing the inferred values of the quadrupolar tidal coupling constant from the inspiral and post-merger phases. Simulations that compare hydrodynamic models with and without quark deconfinement show that the presence of quark deconfinement causes no change in the inspiral gravitational-wave signal due to the low densities involved. This is contrasted to the post-merger phase where the densities are greater, leading to a higher proportion of deconfined quarks when modelled, which in turn leads to a change in the post-merger gravitational-wave strain~\cite{Most2018b}. Phase transitions due to quark deconfinement could be detected by training our model on simulations that disallow quark deconfinement. If a post-merger signal was detected that contained the signature of quark-deconfinement then this would result in an inconsistency between the tidal coupling constant inferred from the post-merger gravitational-wave spectra and the inspiral phase. We note that other mechanisms, either physical or arising from errors in the modelling, could also cause such frequency shifts. We leave the exploration of this as future work. \par
    \section{Discussion}
        In this paper, we use a hierarchical model to generate binary neutron star  post-merger spectra by training on spectra generated with numerical-relativity simulations. Our trained model allows us to generate spectra in ${\sim}10^{-4}$ seconds, which significantly reduces the computational effort required to populate a template bank of spectra. We obtain noise-weighted, amplitude-only fitting-factors across all tested spectra, with a mean of 0.95 and a standard deviation of 0.03. This compares favourably to a post-merger fitting factor of 0.93 between different numerical-relativity codes.\par
        Training on the phase of the post-merger spectra will allow fitting factor comparisons with both the amplitude and phase information, as well as the generation of time-based waveforms. In addition, it will provide insight on the number of numerical-relativity simulations required to achieve a complete and accurate database. While obtaining information on the phase evolution is in principle possible, see, e.g.,~\cite{Bose2018}, this also requires a systematic investigation that goes well beyond the scope of this work. Without the phase information, a matched filter search is less sensitive, but it is still possible to design such a search using only the signal amplitudes.
        \par
        Results based on our trained model suggest that the model is selective and could potentially be used in parameter estimation of detected events. If posteriors for the mass and tidal coupling constant are able to be determined, then it is a simple step to calculate the posteriors for the compactness, Eq.~(\ref{eq:Ccalc1}), and the radius of the neutron star. This will be confirmed in future work using a Bayesian framework. Parameter estimation of the post-merger spectra could set bounds on the post-merger quadrupolar tidal coupling constant, allowing comparison with the inspiral value. This could determine whether a phase change in the equation of state is present~\cite{Bauswein2018,Most2018b}.\par
        To be valuable in search and parameter-estimation studies, our model must be extended to include individual values of mass, spins, compactness and quadrupolar tidal deformabilities for each progenitor. This paper trained on waveforms with progenitor $\kappa^\tau_2$ values ranging from 50 to 350, and equal-mass progenitors from 1.2 to 1.35\,M$_\odot$. The merger of two progenitors with the same parameters $(C,M,\kappa_2^\tau)$, but different equations of state could produce different output spectra, for example if one equation of state had a lower maximum, non-rotating neutron star mass. This could be a problem for the existing model and an additional input parameter may be needed to remove the degeneracy in the output spectra. This is not necessary for the numerical-relativity simulations that are used in the paper, as there is enough variation in the three parameters to define the output spectra. The model may be also expanded to include simulations with black hole progenitors. In this case the dominant post-merger frequency, corresponding to black hole ring-down, would likely be too high for detection with advanced LIGO. These changes can be introduced given enough numerical-relativity simulations to cover the required ranges of parameter values. The placement of numerical-relativity simulations to enable this is a subject of future work. Our method may eventually provide an additional tool to aid in the detection of short-term post-merger neutron star remnants, supplementing the existing tools~\cite{Klimenko2016,Chatziioannou2017}.
\section*{Acknowledgments}
        PDL is supported through Australian Research Council (ARC) Future Fellowship FT160100112 and ARC Discovery Project DP180103155, ARC is supported in part by the Australian Research Council through Discovery Grant DP160100637. We thank James Clark and Eric Thrane for their feedback.

\end{document}

### 2021-10-11 papers/paper2.tex
Initial Overleaf Import

\documentclass[../Thesis.tex]{subfiles}
% \documentclass[
% reprint,
% % prx,
% prd,
% twocolumn,
% % onecolumn,%!TEX encoding = UTF-8 Unicode
% nofootinbib,
% % ,secularistic%
% tightenlines %
% floatfix,
%  amsmath,
% showpacs ,amssymb, aps,% linenumber
% superscriptaddress
% ]{revtex4-2}
% \usepackage{amsmath,slashed}
% \usepackage[left]{lineno}
% \usepackage{units}
% \usepackage{comment}
% \usepackage{graphicx}
% \usepackage{amsmath,amssymb}
% \usepackage{support-caption}
% \usepackage{subcaption}
% \usepackage{bm}
% \usepackage[shortlabels]{enumitem}
% \usepackage{float}
% \usepackage{soul}
% \usepackage{natbib}
% \usepackage{wasysym}
% \usepackage{aas_macros}
% \graphicspath{{images/}}
% \RequirePackage{lineno}
% \usepackage[usenames,dvipsnames,table]{xcolor}
% \usepackage{appendix}
% \usepackage{extdash}
% \usepackage{colortbl}
% % \usepackage{titlesec}

% \definecolor{lightblue}{HTML}{1F77B4}
% \newcommand{\SLyWaveform}[1]{\textcolor{lightblue}{#1}}
% \newcommand{\red}[1]{\textcolor{red}{#1}}
% \newcommand{\pdl}[1]{\textcolor{red}{[PDL: #1]}}
% \newcommand{\pje}[1]{\textcolor{Purple}{[PJE: #1]}}
% \newcommand{\blue}[1]{\textcolor{blue}{#1}}
% \DeclareMathOperator*{\argmin}{arg\,min}
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Vic 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
% \newcommand{\AstroThreeD}{ARC Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), Canberra, ACT 2611, Australia}
% \newcommand{\GeorgiaTech}{Center for Relativistic Astrophysics and School of Physics, Georgia Institute of Technology, Atlanta, Georgia 30332, USA}
% \newcommand{\Flatiron}{Center for Computational Astrophysics, Flatiron Institute, 162 5th Ave, New York, NY 10010}
% \captionsetup{belowskip=-8pt,aboveskip=0pt}
% \titlespacing*{\section}
% {0pt}{5.5ex plus 1ex minus .2ex}{4.3ex plus .2ex}

\definecolor{Waveform1}{HTML}{1f77b4}
\definecolor{Waveform2}{HTML}{ff7f0e}
\definecolor{Waveform3}{HTML}{2ca02c}
\definecolor{Waveform4}{HTML}{d62728}
\definecolor{Waveform5}{HTML}{9467bd}
\definecolor{Waveform6}{HTML}{8c564b}
\definecolor{Waveform7}{HTML}{e377c2}
\definecolor{Waveform8}{HTML}{7f7f7f}
\definecolor{Waveform9}{HTML}{bcbd22}

\begin{document}

\begingroup
\clearpage% Manually insert \clearpage
\let\clearpage\relax% Remove \clearpage functionality
\vspace*{-2cm}% Insert needed vertical retraction

\chapter{Detection and parameter estimation of binary neutron star merger remnants}
\label{chapter:DetectionPE}
\endgroup

\chapterauthor{Paul J. Easter\\Sudarshan Ghonge\\Paul D. Lasky\\Andrew R. Casey\\James A. Clark\\Francisco Hernandez Vivanco\\Katerina Chatziioannou}
% \email{paul.easter@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Sudarshan Ghonge}
% \affiliation{\GeorgiaTech}
% \author{Paul D. Lasky}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Andrew R. Casey}
% \affiliation{\SPA}
% \affiliation{\AstroThreeD}
% \author{James A. Clark}
% \affiliation{\GeorgiaTech}
% \author{Francisco Hernandez Vivanco}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Katerina Chatziioannou}
% \affiliation{\Flatiron}
% \pacs{
% }

\section*{Abstract}
    Detection and parameter estimation of binary neutron star merger remnants can shed light on the physics of hot matter at supranuclear densities.
    Here we develop a fast, simple model that can generate gravitational waveforms, and show it can be used for both detection and parameter estimation of post-merger remnants.
    The model consists of three exponentially-damped sinusoids with a linear frequency-drift term.
    We test the model against nine equal-mass numerical-relativity simulations selected for emission of gravitational waves for $\gtrsim25$\,ms.
    The median fitting factors between the model waveforms and numerical-relativity simulations exceed 0.90.
    We detect remnants at a post-merger signal-to-noise ratio of $\ge 7$ using a Bayes-factor detection statistic with a threshold of 3000.
    We can constrain the primary post-merger frequency to $\pm_{1.2}^{1.4}\%$ at post-merger signal-to-noise ratios of 15 with an increase in precision to $\pm_{0.2}^{0.3}\%$ for post-merger signal-to-noise ratios of 50.
    The tidal coupling constant can be constrained to $\pm^{9}_{12}\%$ at post-merger signal-to-noise ratios of 15, and $\pm 5\%$ at post-merger signal-to-noise ratios of 50 using a hierarchical inference model.

% \maketitle
    \section{Introduction}
    \label{sec:introduction2}
    Gravitational waves have been directly detected from the inspiral of binary neutron star mergers~\cite{GW170817Detection,GW190425Detection}. 
    The  post-merger remnant may promptly collapse into a black hole, or form a hot, differentially-rotating neutron star~\cite{Baumgarte2000,Shibata2000bar}, which emits gravitational waves~\cite[e.g.,][]{Nakamura1994,New1997,Rasio1999,Shibata2005}.
    Numerical-relativity simulations of post-merger remnants show relationships between the gravitational-wave spectra and a number of progenitor properties through quasi-universal relationships~\cite[e.g.,][]{Bauswein2012,Bauswein2012a,Hotokezaka2013,Bernuzzi2014,Takami2014,Bernuzzi2015,Bauswein2015,Takami2015,Rezzolla2016,Bauswein2019}.
    Of particular interest is the relationship between the progenitor tidal coupling constant and the primary post-merger oscillation frequency for baryonic equations of state~\cite{Takami2015,Bernuzzi2015,Rezzolla2016,Breschi2019}, which can be used to place constraints on the tidal coupling constant.\par

    Gravitational-wave spectra generated from numerical-relativity simulations show consistent  features related to the dynamics of the surviving remnant.
    A dominant peak, designated as $f_{\mathrm{peak}}$~\cite[][]{Oechslin2007a}, is produced by the fundamental oscillations of the bar-mode deformed post-merger remnant~\cite[e.g.,][]{Shibata2003a,Rezzolla2010,Giacomazzo2011,Hotokezaka2011,Bauswein2012}.
    The frequencies of four possible peaks can be labelled as $(f_{2-0}, f_{\mathrm{spiral}},f_{\mathrm{peak}},f_{2+0})$ in ascending order~\cite{Bauswein2015}. 
    The peaks at frequencies $f_{2-0},f_{2+0}$ may result from coupling between a quasi-radial oscillation mode and $f_{\mathrm{peak}}$~\cite{Stergioulas2011}.
    The peak at frequency $f_{\mathrm{spiral}}$ may result from the slower rotation-rate of tidally-deformed matter at the outer edges of the post-merger remnant~\cite{Bauswein2015}.
    See Refs.~\cite{Takami2014,Takami2015} for an alternative proposed explanation of the frequency peaks.
\par
\begin{figure*}
        \centering
        \includegraphics[scale=0.38]{{time_freq_0ms_time_freq}.png}
        \caption
        {Waveform reconstruction of numerical-relativity post-merger signal injections. Top panels: time (left) and frequency (right) domain reconstructions of a numerical-relativity simulation using the SLy equation of state with equal-mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform SLy-M1.350-$\Lambda$390). The post-merger waveform (black curve) is injected at a post-merger signal-to-noise ratio of 50. The reconstructed waveforms are shown in blue. Bottom panels: same as the top panels except the injected waveform is using the LS220 equation of state with equal mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform LS220-M1.350-$\Lambda$684). The reconstructed waveforms are shown in orange. Noise sensitivity curves are shown for Advanced LIGO (dashed black) and Advanced Virgo (dotted black) for plots on the right.} 
        \label{fig:TimeFrequencyResponse}
\end{figure*}
    In this paper, we develop a Bayesian detection and parameter-inference pipeline.
    Normally these pipelines require a large bank of waveforms.
    Numerical-relativity simulations cannot be used to generate these waveforms as each simulation requires~${\sim\,\mathcal{O}(10^5)}$ CPU hours to complete~\cite{Takami2015}.
    We develop a fast, simple model of gravitational waves for post-merger remnants that phenomenologically incorporates the main frequencies previously mentioned.
    Our model produces waveforms in a time-frame that is suitable for use in detection and parameter estimation of binary neutron star post-merger remnants.\par

    We match nine equal-mass numerical-relativity waveforms with gravitational-wave emission for $\gtrsim 25$\,ms with fitting factors of 0.92-0.97.
    This model addresses the two restrictions that prevent matched filtering of post-merger gravitational-wave strain: computational time and poor fitting factors.
    Our model is derived from a hybrid of the two models outlined in Refs.~\cite{Bauswein2016} and~\cite{Bose2018}. 
    Our model is agnostic to the locations of the frequency peaks and uses Bayesian statistics to determine the actual peak frequencies.
    Furthermore, the addition of a frequency drift term allows for secular changes in the frequency peak locations.
    With post-merger signal-to-noise ratios of $\ge 15$, the model can localise the primary post-merger frequency to $\pm_{1.2}^{1.4}\%$ at 95\% confidence, reducing to $\pm_{0.2}^{0.3}\%$ at post-merger signal-to-noise ratio of 50.
    Using the hierarchical model developed in Ref.~\cite{Easter2019} we can then constrain the tidal parameters and compactness of the progenitor neutron stars.
    The tidal coupling constant is constrained to $\pm_{12}^{9}\%$ at post-merger signal-to-noise ratios of 15 for a 95\% confidence interval. 
    At post-merger signal-to-noise ratios of 50 this tightens to $\pm 5 \%$. \par

    In Section~\ref{sec:methodology2} we outline the model and associated methods used in this paper. 
    In Section~\ref{sec:modelfit2} we validate the model fits in the time and frequency domains and quantify the goodness of the fits.
    In Section~\ref{sec:sensitivity2} we use a Bayes factor detection statistic to determine at what post-merger signal-to-noise ratios a detection occurs and test how the model performs due to uncertainty in the inspiral coalescence time.
    In Section~\ref{sec:ParameterEstimation2} we calculate posteriors of the dominant post-merger frequency and introduce the hierarchical model from Ref.~\cite{Easter2019} to find the equation of state parameters for the progenitors.
    We find constraints on both the tidal coupling constant and the compactness of the progenitors.

    \section{Methodology}\label{sec:methodology2}
    We adopt a model for the post-merger gravitational-wave signal consisting of three exponentially damped sinusoids~\cite{Bauswein2016} with additional linear frequency drift terms~\cite{Bose2018}. 
    The plus, $h_{+}(\boldsymbol{\theta},t)$, polarisation of the gravitational-wave strain is extracted from the right circular polarisation, $h(\boldsymbol{\theta},t)$,  as follows:
\begin{align}
        h(\boldsymbol{\theta},t) & =  h_{+}(\boldsymbol{\theta},t) - \mathrm{i}\, h_{\times}(\boldsymbol{\theta},t) \label{eq:htotal2}\\
        & =  \sum_{j=0}^{2}h_{j,+}(\boldsymbol{\theta},t) - \mathrm{i}\, h_{j,\times}(\boldsymbol{\theta},t),\label{eq:hsum2} \\
        h_{j,+}(\boldsymbol{\theta},t) & = 
        H w_j \exp\left[-\frac{t}{T_j}\right] \cos \left(2\pi f_j t\left[1+\alpha_j t\right]+\psi_j \right). \label{eq:hmode2}
\end{align}
        Here, ${\boldsymbol{\theta} = \{H,w_j,T_j,f_j,\alpha_j,\psi_j :  j \in [0,2] \}}$ are the model parameters where $H$ is the amplitude scaling factor and $w_j$ is the relative scaling factor for each mode, $j\in [0,2]$, such that $\sum_j w_j = 1$. 
        The initial frequency of each mode is given by $f_j$, $T_j$ are the damping times, $\psi_j$ are the initial phases, and $\alpha_j$ are the frequency drift terms. 
        The time, $t$, is defined such that $t=0$ corresponds to the coalescence time when the maximum of $h_+^2(\boldsymbol{\theta},t)+h_\times^2(\boldsymbol{\theta},t)$ occurs~\cite[e.g.,][]{Read2013,Takami2015,Rezzolla2016,Easter2019}.
        The cross polarisation of the $j$th mode is generated by a $\pi/2$ phase shift on $h_{j,+}(\boldsymbol{\theta},t)$.
        Setting $\alpha_j=0$ allows detection of signals corresponding to the cross-polarisation model in Ref.~\cite{Bauswein2016}. 
        These equations are a subset of the plus polarisation model in Ref.~\cite{Bose2018} with the quadratic drift term set to zero and no explicit modulation of spectral peaks.
        \par

        We use nine post-merger numerical-relativity simulations from Ref.~\cite{Dietrich2018} (see Appendix~\ref{appendix:a2} for details), selecting only simulations with equal-mass progenitors where a nascent neutron star
        survives for at least $\sim\unit[25]{ms}$.
        For equal-mass systems, the tidal parameter of the neutron stars is related to the dimensionless compactness, $C=GM/(Rc^2)$, and the second Love number, $k_2$, as follows:
        \begin{align}
            \tilde{\Lambda} = \frac{2}{3} k_2 C^{-5}, \\
            \kappa_2^{\mathrm{T}} = \frac {1}{8}k_2 C^{-5},
        \end{align}
        where $\tilde{\Lambda}$ is the quadrupolar tidal deformability and $\kappa_2^{\mathrm{T}}$ is the total quadrupolar tidal coupling constant. 
        Here, $M$ is the neutron star mass, $R$ is the neutron star radius, $G$ is the gravitational constant, and $c$ is the speed of light. 
        The tidal properties of the progenitors can be estimated from the dominant post-merger frequency using relations found from numerical-relativity simulations with baryonic equations of state~\cite{Takami2015,Bauswein2019} (although see Refs.~\cite{Most2018b,Most2018,Bauswein2019} for the consequences of a phase transition to strange matter). 
        We discuss this more in Section~\ref{sec:ParameterEstimation2}.      \par

        We inject numerical-relativity waveforms at various post-merger signal-to-noise ratios into a three-detector network (LIGO Hanford, Livingston, and Virgo) at design sensitivity for each interferometer~\cite{PSD:aLIGO,PSD:aVirgo}. 
        We inject the post-merger signal at a fixed time and fixed sky position, assuming that we know the coalescence time from the inspiral stage. 
        In Section~\ref{sec:sensitivity2} we test this assumption by determining the uncertainty in the coalescence time for various signal-to-noise ratios.
        We use the {\textsc{Bilby}} package  \cite{Ashton2019} with the {\textsc{Dynesty}} sampler~\cite{Speagle2019} to sample posteriors, $p(\boldsymbol{\theta}|d)$, of the model parameters  using the likelihood, $\mathcal{L}(d | \boldsymbol{\theta})$, as follows:
        \begin{align}
        	p(\boldsymbol{\theta}|d) & =\dfrac{\mathcal{L}(d | \boldsymbol{\theta})\pi(\boldsymbol{\theta})}{\mathcal{Z}},\label{eq:posterior2}\\
        	\mathcal{Z} & =\int_{\boldsymbol{\theta}} d\boldsymbol{\theta} \mathcal{L}(d | \boldsymbol{\theta})\pi(\boldsymbol{\theta}),\label{eq:evidence2}\\
        	\mathcal{L}(d | \boldsymbol{\theta}) & \propto \exp\Big[-\Big<d(t)-h(\boldsymbol{\theta},t),d(t)-h(\boldsymbol{\theta},t)\Big>\Big]\label{eq:likelihood2}. 
        \end{align}        
        Here, $d(t)=s(t)+n(t)$ is the numerical-relativity waveform, $s(t)$, injected into noise, $n(t)$. 
        We simulate ten different Gaussian noise realisations with {\textsc{Bilby}}, to examine the response of the model to variations in detector noise. 
        We limit this to ten noise realisations to keep the computation time manageable.
        The priors on the model parameters are $\pi(\boldsymbol{\theta})$. 
        The noise-weighted inner product in Eq.~\ref{eq:likelihood2} is defined by:
        \begin{align}
        	\left<h_1,h_2\right>\equiv4\, \mathrm{Re} \int df\frac{\tilde{h}_1(f)\tilde{h}^{*}_2(f)}{S_h(f)},\label{eq:inner_product2}
        \end{align}
    where $S_h$ is the detector's noise power spectral density.
    We use a sampling frequency of 16384\,Hz to eliminate aliasing of the upper sidebands. 
    We use constrained priors to sort the maximum amplitude for $\tilde{h}_{j,+}(\boldsymbol{\theta},f)$, such that  $|\tilde{h}_{j,+}(\boldsymbol{\theta},f)|_{max} > |\tilde{h}_{j+1,+}(\boldsymbol{\theta},f)|_{max}$.
    This ensures that the mode zero ($j=0$) exponentially damped sinusoid corresponds to the dominant post-merger frequency. 
    Full details on the priors are given in Appendix~\ref{appendix:b2}. 
    The optimal post-merger signal-to-noise ratio, $\rho_{\mathrm{opt}}$, is calculated from the quadrature sum of the optimal post-merger signal-to-noise ratio for each of the three detectors, $\rho_{\mathrm{opt},i}$ as follows:
\begin{equation}
    \rho_{\mathrm{opt}}^2 = \sum_{i \in  \mathrm{HLV}} \rho_{\mathrm{opt},i}^2\quad,\label{eq:OptimalSNR2}
\end{equation}
    for $t \geq 0$.
    The matched filter signal-to-noise ratio for a single detector is given by:
\begin{equation}
    \rho_{mf} = \frac{\left<d, h(\boldsymbol{\theta})\right>}{\left< h(\boldsymbol{\theta}),h(\boldsymbol{\theta})\right>^\frac{1}{2}}\quad.\label{eq:MatchedFilterSNR2}
\end{equation} \par

\section{Model Validation}\label{sec:modelfit2}
     Figure~\ref{fig:TimeFrequencyResponse} shows the posterior waveforms in the time and frequency domain for the plus polarisation of two numerical-relativity post-merger simulations. The two gravitational-wave simulations, SLy-M1.350-$\Lambda$390 (THC:0036:R03, top) and LS220-M1.350-$\Lambda$684 (THC:0019:R05, bottom) are injected at a post-merger signal-to-noise ratio of 50. 
     These waveforms are chosen for compatibility with the inferred properties of $\Lambda$ from GW170817~\cite{GW170817Detection,Annala2018,Radice2018,Most2018,De2018,GW170817Properties}. 
     SLy-M1.350-$\Lambda$390 is a simulation of equal progenitor mass $1.35\,M_\odot$ neutron stars with tidal deformability, $\tilde{\Lambda}=390.1$ ($\kappa_2^{\mathrm{T}}=73.14$) and SLy equation of state. Similarly, LS220-M1.350-$\Lambda$684 has masses of $1.35\,M_\odot$, $\tilde{\Lambda}=683.8$ ($\kappa_2^{\mathrm{T}}=128.2$) and LS220 equation of state. \par
     We generate posterior waveforms by randomly drawing samples from the posterior distribution $p(\boldsymbol{\theta}|d)$. The posterior waveforms are shown as  blue (top, SLy-M1.350-$\Lambda$390) and orange (bottom, LS220-M1.350-$\Lambda$684)  curves in Fig.~\ref{fig:TimeFrequencyResponse}. 
     The solid black curves show the injected numerical-relativity waveforms. 
     As can be seen in the time-response plots (Fig.~\ref{fig:TimeFrequencyResponse}, left), the posterior samples are tightly clustered around the numerical-relativity simulations, particularly for the first $\sim\,15\,$ms. 
     The phase of waveform SLy-M1.350-$\Lambda$390 is lost after $\sim\,15\,$ms (Fig.~\ref{fig:TimeFrequencyResponse}, upper left) though the majority of the spectral content is contained in the first $5\,$ms (see Fig.~\ref{fig:OF_MF_SNR_FF5ms}).
     We note that accumulated phase errors in numerical-relativity simulations increase over time.\par

     The frequency-response plots  are shown on the right side of Fig.~\ref{fig:TimeFrequencyResponse}, along with the amplitude spectral density of Advanced LIGO (dashed black curve) and Advanced Virgo (dotted black curve) at design sensitivity. 
     The primary frequency peaks are well recovered for both reference waveforms. 
     Two low frequency peaks of SLy-M1.350-$\Lambda$390 are resolved in preference to the upper frequency peak, whereas only one low frequency peak is resolved for LS220-M1.350-$\Lambda$684. The other two modes are located at the main frequency peak of LS220-M1.350-$\Lambda$684.\par

     To measure the extent of the waveform mismatch, we calculate the noise-weighted fitting factor between the injected numerical-relativity waveform, $d(t)$, and the posterior waveform, $h(\boldsymbol{\theta},t)$, ~\cite{Apostolatos95}:
    \begin{equation}
    	\mathcal{F}(d(t),h(\boldsymbol{\theta},t))\equiv\frac{\left<d(t)|h(\boldsymbol{\theta},t)\right>}{\sqrt{\left<d(t)|d(t)\right>\left<h(\boldsymbol{\theta},t)|h(\boldsymbol{\theta},t)\right>}}.\label{eq:FF2}
    \end{equation}
    The fitting factor, calculated with noise from one detector at Advanced LIGO design sensitivity~\cite{PSD:aLIGO}, quantifies the loss in signal-to-noise due to signal mismatch in relation to an optimal signal-to-noise ratio, Eq.~\ref{eq:OptimalSNR2}. \par

    The median fitting factors are $0.92$ and $0.95$, for SLy-M1.350-$\Lambda$390 and LS220-M1.350-$\Lambda$684, respectively. As the detection rate scales as $\mathcal{F}^3$~\cite{Apostolatos95}, the reduction in detection rate due to the above mismatch is 22\% and 14\% respectively for these two waveforms. \par
        \begin{figure}[H]
        \centering
        \includegraphics[scale=0.5]{NRviolinFittingFactorsV2_0msFF.pdf}
        \caption{Fitting-factor distributions, $\mathcal{F}(d(t),h(\boldsymbol{\theta},t))$, for each post-merger numerical-relativity waveform. 
        The signal-to-noise ratio for the post-merger gravitational-wave strain for each waveform is 50. 
        The upper and lower horizontal bars represent 99.7\% confidence intervals. 
        The central horizontal bar shows the median value. 
        The thick vertical line shows the 95\% confidence intervals. 
        The median fitting factors range between 0.92 to 0.97 which corresponds to a reduction in detection rate from $22\%$ down to $9\%$ due to mismatch with the numerical-relativity injections.} 
        \label{fig:FittingFactors}
    \end{figure} 
    The fitting factors for all nine numerical-relativity simulations are shown in Fig.~\ref{fig:FittingFactors}, with each simulation represented by a  different colour.
    Ten different Gaussian noise realisations are used for each numerical-relativity simulation.
    The 99.7\% confidence intervals for the fitting factors are shown by the upper and lower horizontal bars.
    The median value is shown by the central horizontal bar, and 95\% confidence intervals are indicated by thick vertical bars.
    Finally, the distribution of the fitting factors are shown by the width of the shaded areas.
    The lowest fitting factors, for simulation, SLy-M1.350-$\Lambda$390, have an average match of 0.92. 
    Other numerical-relativity injections have fitting factors of $0.95-0.97$. 
    The injection with the softest equation of state under-performs the other injections. 
    This is due to complex dynamics of the nascent neutron star in the first $\sim 2\,$ms.
\par

\section{Sensitivity}\label{sec:sensitivity2}
    We calculate the Bayes factor between the signal hypothesis and a noise hypothesis to evaluate the sensitivity of our model. 
    We do this by injecting  the post-merger signal SLy-M1.350-$\Lambda$390 into ten different noise realisations at various signal-to-noise ratios. 
    The results are shown in Fig.~\ref{fig:BFvsSNR}.
    The distribution of the natural logarithm of the Bayes factor, $\ln(\mathcal{BF})$, is shown for each post-merger signal-to-noise ratio along with the 99.7\% confidence intervals (upper and lower horizontal bars) and the median value (middle horizontal bar). 
    We define that strong evidence for a signal hypothesis over a noise hypothesis corresponds to a Bayes factor exceeding 3000 ($\ln(\mathcal{BF})> 8.0$)~\cite[e.g.,][]{Jeffreys61}. 
    In this case a signal hypothesis is 3000 times more likely than a noise hypothesis.
    This occurs with post-merger signal-to-noise ratios of $\gtrsim 10$ for all tested noise realisations.
    There are some noise realisations where the Bayes factor threshold is exceeded for post-merger signal-to-noise ratios of $\approx 7\Hyphdash*9$.\par
    \begin{figure}[H]
         \centering
         \includegraphics[scale=0.5]{NRviolinBayesFactorsV2_0ms.pdf}
         \caption{Natural logarithm of the Bayes factor comparing a signal hypothesis against a noise hypothesis plotted against the post-merger signal-to-noise ratio.
         The numerical-relativity waveform, SLy-M1.350-$\Lambda$390, is injected into ten different noise realisations at the specified signal-to-noise ratio.  
         The upper and lower horizontal bars show the 99.7\% confidence intervals of the log Bayes factor and the central horizontal bar shows the median value. 
         A post-merger signal-to-noise value of $\gtrsim 10$ is required to ensure strong evidence for a signal hypothesis ($\ln(\mathcal{BF})>8$) across all ten noise realisations. However, there are some noise realisations where $\ln(\mathcal{BF})>8$ occurs for post-merger signal-to-noise ratios as low as 7.}
         \label{fig:BFvsSNR}
     \end{figure}  
     An important consideration for our signal model is the uncertainty in the coalescence time as measured from the gravitational-wave inspiral signal. 
     This determines how close we can get to the true coalescence time for the binary neutron star merger.
     In Fig.~\ref{fig:OF_MF_SNR_FF5ms} we investigate the model performance to uncertainties in the coalescence time.
     We show how the fitting factor and matched-filter signal-to-noise ratio change when starting the adopted model at various times after the coalescence time.
     We multiply the numerical-relativity injection, $d(t)$, by the Heaviside step function, $\mathcal{H}(t-t_{delay})$, and evaluate the model, $h(\boldsymbol{\theta},t-t_{delay})$, for $t \geq t_{delay}$.
     The matched filter signal-to-noise ratio is calculated using Eq.~\ref{eq:MatchedFilterSNR2} with a single detector at Advanced LIGO sensitivity.
     We use numerical-relativity injection, SLy-M1.350-$\Lambda$390, selected due to compatibility with the tidal parameters inferred from GW170817.
     A delay time of zero includes the entire post-merger waveform, whereas a delay time of $2\,$ms excludes the first $2\,$ms of the injection after the coalescence time. 
     The fitting factor is lower ($\sim 0.91$) for small delay times and increases to $\sim\,0.96$ at $2\,$ms. 
     The fitting factor is lower in the first $2\,$ms due to complex dynamics of the nascent neutron star.
     In Fig.~\ref{fig:OF_MF_SNR_FF5ms}, the matched-filter signal-to-noise ratio is almost monotonically decreasing as expected. 
     Even though the fitting factors are lower at zero delay time, the matched-filter signal-to-noise ratio is at maximum. 
     Therefore, from a sensitivity perspective, a minimum delay time is preferred.
\begin{figure}[H]
         \centering
         \includegraphics[scale=0.5]{OF_MF_SNR_FF5ms.pdf}
         \caption{Variation of fitting factor (hatched blue) and matched filter  signal-to-noise ratio (solid blue) for differing values of the delay time after the time of coalescence. The shaded regions show 95\% confidence intervals. The optimal signal-to-noise ratio is also shown (solid black). Although the fitting factor is lower when the entire post-merger signal is used, the matched filter signal-to-noise ratio is largest. The fitting factor is lower for smaller delay-times due to the complex dynamics of the nascent neutron-star.}
     \label{fig:OF_MF_SNR_FF5ms}
     \end{figure}  
     To estimate the uncertainties of the time of coalescence of the inspiral signal as a function of the signal-to-noise ratio of the inspiral signal, we use a Fisher matrix approximation. 
     We assume that the signal parameters $\boldsymbol{\vartheta}$ follow a Gaussian distribution: 
\begin{equation}
    p \left( \Delta \boldsymbol{\vartheta} \right) \propto \exp \left[-\frac{1}{2}\Gamma_{ij}\Delta \vartheta^i \Delta \vartheta^j\right].
\end{equation}
    \sloppy Here, $\Delta \vartheta_i = \vartheta^i-\hat{\vartheta}^i$, $\hat{\vartheta}^i$ are the best fit inspiral parameters and $\Gamma_{ij} = (\partial h/ \partial \vartheta_i | \partial h/ \partial \vartheta_j)$ is the expected Fisher information matrix. 
    The estimated errors of the parameters, $\vartheta_i$, are obtained by taking the diagonal elements of the Fisher information matrix. 
    The relevant parameters within our approximation are $\boldsymbol{\vartheta} = (\mathcal{M}, q, \phi_c, \tilde{\Lambda}, t_c,H)$, where $\mathcal{M}$ is the chirp mass, $q$ is the mass ratio, $\phi_c$ is the phase of coalescence.
    The average-weighted tidal deformabilty is  $\tilde{\Lambda}$, $t_c$ is the time of coalescence and $H$ is the amplitude of the inspiral waveform.
    We calculate the errors on $\vartheta_i$ assuming an equal-mass 1.4\,$M_\odot$ non-rotating progenitor system. 
    The expected uncertainties for the coalescence time are shown in  Fig.~\ref{fig:TCOALESCENCE}. 
    The left axis shows the inspiral signal-to-noise ratio for an optimally oriented source into a two detector LIGO network at design sensitivity.
    We use Fig.~8 from Ref.~\cite{Martynov2019} to determine the luminosity distance, $D_L$, from the inspiral signal-to-noise ratio. 
    We calculate the product of $D_{L0}\approx 475$\,Mpc (at $z\approx 0.1$) with the corresponding inspiral signal-to-noise ratio, $\rho_{inspiral,0}\approx 7$.
    We inject the numerical-relativity post-merger waveform, SLy-M1.350-$\Lambda$390, at luminosity distance, $D_L = D_{L0}\left(\rho_{inspiral,0}/\rho_{inspiral}\right)$, and evaluate the post-merger signal-to-noise ratio using the Advanced LIGO and Virgo detector network at design sensitivity.
    The right axis in  Fig.~\ref{fig:TCOALESCENCE} shows the corresponding post-merger signal-to-noise ratio.\par
    \begin{figure}[H]
         \centering
         \includegraphics[scale=0.5]{TCOALESCENCE.pdf}
         \caption{Uncertainty in coalescence time plotted with signal-to-noise ratios. The coalescence time uncertainty has been determined by a Fisher matrix approximation. The left axis is the signal-to-noise ratio for a two detector network of Advanced LIGO at design sensitivity for a binary neutron star inspiral. The right axis is the post-merger signal-to-noise ratio for a three detector network of Advanced LIGO and Advanced Virgo at design sensitivity using numerical-relativity simulation SLy-M1.350-$\Lambda$390. Post-merger signal to noise ratios above 6.0 have coalescence time uncertainties of less than 0.1\,ms.}
         \label{fig:TCOALESCENCE}
     \end{figure}  
    For post-merger signal-to-noise ratios larger than 6, the uncertainty in the coalescence time is less than 0.1\,ms.
    This shows that, for post-merger signal-to-noise ratios of interest in this work, the coalescence time is similarly constrained. 
    The uncertainty in coalescence time can be related to Fig.~\ref{fig:OF_MF_SNR_FF5ms} to show that the resultant matched filter signal-to-noise ratio is not significantly reduced due to the uncertainty in the coalescence time.

\section{Parameter  estimation}\label{sec:ParameterEstimation2}
    Estimation of the primary post-merger frequency is another important indicator of the model performance.
    We estimate this by calculating posteriors of the peak frequency, $f_{\mathrm{peak}}$, of the dominant mode.
    Fig.~\ref{fig:BWFREQvsSNR} shows posteriors of $f_{\mathrm{peak}}$ as a function of post-merger signal-to-noise ratio.
    These have been calculated for an injection of SLy-M1.350-$\Lambda$390 at post-merger signal-to-noise ratios  of $\ge 9$.
    The noise realisation was kept the same for all injections. 
    Blue shading indicates regions of 95\% confidence intervals and the median values are shown as blue dots.
    The frequency corresponding to the maximum value of the characteristic strain spectrum of the  numerical-relativity signal, $|\tilde{s}_+(f)|\sqrt{f}$, is shown as a black horizontal line.
    This can be thought of as an approximation of the true injected value of $f_{\mathrm{peak}}$.
    The $f_{\mathrm{peak}}$ frequency is constrained within 95\% confidence intervals to $3310\pm^{46}_{38}$\,Hz at a post-merger signal-to-noise ratio of 15 which corresponds to $\pm^{1.4}_{1.2}\%$.
    At a post-merger signal-to-noise ratio of 50, the precision increases to $3296\pm^{11}_{8}$\,Hz $(\pm^{0.3}_{0.2}\%)$.
    The posteriors for $f_0, \alpha_0, f_1$ and $\alpha_1$, determined for all numerical-relativity injections at a post-merger signal-to-noise ratio of 50, are shown in Figs.~\ref{fig:Corner1}-\ref{fig:Corner9} in Appendix~\ref{appendix:c2}.\par
   \begin{figure}[H]
         \centering
         \includegraphics[scale=0.5]{BWPrimaryFrequency2sigma.pdf}
         \caption{Primary post-merger frequency comparison between our model (blue) and \textsc{BayesWave} (brown). The posteriors are plotted against the post-merger signal-to-noise ratio for injection SLy-M1.350-$\Lambda$390. The 95\% confidence intervals are shaded. The median points are shown as blue dots and brown crosses, for our model and \textsc{BayesWave}, respectively. The frequency corresponding to the peak of the spectral response of the injection is also shown (solid black line).
         \label{fig:BWFREQvsSNR}}
     \end{figure}
    We also analyse injections of SLy-M1.350-$\Lambda$390 using \textsc{BayesWave} \cite{Cornish2015,Littenberg2015}. \textsc{BayesWave} uses a variable number of Morlet-Gabor wavelets to model the signal, where both the number and the properties of the wavelets are marginalised over.
    This is an established method for post-merger studies~\cite{GW170817Properties,GW190425Detection}.
    References~\cite{Chatziioannou2017,Torres-Rivas2019} have performed simulations using \textsc{BayesWave} to infer the post-merger properties of binary neutron star mergers.
    We compute the posteriors of the spectral frequency peak, $f_{\mathrm{peak}}$, using \textsc{BayesWave} following Ref.~\cite{Chatziioannou2017}. Here, $f_{\mathrm{peak}}$, the frequency of the highest peak in the Fourier power spectrum of the signal, is determined for each sample from the \textsc{BayesWave} posterior. 
    For samples that do not have a peak, $f_{\mathrm{peak}}$ is computed using  random draws from its prior~\cite{Chatziioannou2017}.
    Figure~\ref{fig:BWFREQvsSNR} shows the 95\% confidence intervals of $f_{\mathrm{peak}}$ for each post-merger signal-to-noise ratio in brown. The median values are shown as brown crosses.
    The \textsc{BayesWave} frequency posteriors are consistent with Refs.~\cite{Chatziioannou2017,Torres-Rivas2019}.

    The posteriors for $f_{\mathrm{peak}}$ are similarly constrained for both \textsc{BayesWave} and our model for post-merger signal-to-noise ratios of $\gtrsim 20$. 
    \textsc{BayesWave} is more  constrained for post-merger signal-to-noise ratios of $\sim 9 - 15$.
    Both methods are able to recover the injected dominant post-merger frequency.
    \textsc{BayesWave} can generate very high fitting factors; the fitting factors for SLy-M1.350-$\Lambda390$ at a post-merger signal-to-noise ratio of 50 are $\approx 0.99$.
    The dimensionality of \textsc{BayesWave} is $\sim 90\  (\sim 18$ wavelets) at this post-merger signal-to-noise ratio.
    The dimensionality of our adopted model is 15 with fitting factors of $\approx 0.92$ for SLy-M1.350-$\Lambda390$.
    Furthermore, \textsc{BayesWave} can generalise to any signal (e.g. glitches).
    In contrast, our model has been developed to suit a post-merger gravitational-wave signal.
    The parameters in our model are interpretable: 
    for example, in Fig.~\ref{fig:Corner1}, the $\alpha_0$ value for SLy-M1.350-$\Lambda390$ is $-1.60\pm^{0.50}_{0.26}$ which
     shows that the frequency of the dominant gravitational-wave mode is decreasing.

    The hierarchical model from Ref.~\cite{Easter2019} allows a bidirectional relationship between equal-mass progenitor neutron star properties $(C,M,\kappa_2^\textsc{t})$ and numerical-relativity post-merger simulations. 
    This is achieved by a two step process. 
    Firstly, the progenitor properties are used to solve $\bar{C}(M,\kappa_2^\textsc{t})=C$ using a power-law relationship.  
    Secondly, the model parameters, $\boldsymbol{\Theta}$, are determined by solving $h_c = \boldsymbol{\Theta}\, \boldsymbol{X}(\bar{C}(M,\kappa_2^\textsc{t}),M,\kappa_2^\textsc{t})$, where $h_c$ is the numerical-relativity amplitude spectra for the characteristic strain $(h_c(f)=|\tilde{h}(f)|\sqrt{f})$.  
    Here, $\boldsymbol{X}(\bar{C} (M,\kappa_2^\textsc{t}),M,\kappa_2^\textsc{t})$ is a design matrix derived from the progenitor properties $M$ and $\kappa_2^\textsc{t}$.\par

    We use the posteriors from Section~\ref{sec:modelfit2} to calculate the amplitude of the characteristic spectrum $|\tilde{h}_{+}(\boldsymbol{\theta},f)|\sqrt{f}$ and use the trained model, $\boldsymbol{\Theta}$,  to determine the hierarchical model posteriors on $\kappa_2^\textsc{t}$ and $C$.
    The cross-polarisation waveforms are discarded because the hierarchical model only uses the magnitude of the spectra, and  $|\tilde{h}_{+}(\boldsymbol{\theta},f)|=|\tilde{h}_{\times}(\boldsymbol{\theta},f)|$.
    The hierarchical model, $\boldsymbol{\Theta}$, was previously trained on 35 numerical-relativity simulations from Ref.~\cite{Rezzolla2016}, a distinct set of numerical-relativity simulations to those used in this paper.
    Therefore, this is an out-of-sample model validation. \par
\begin{figure}[H]
         \centering
         \includegraphics[scale=0.5    ]{KappaPosterior.pdf}
         \caption{Tidal coupling constant posteriors versus post-merger signal-to-noise ratio for numerical-relativity waveform SLy-M1.350-$\Lambda$390. The tidal coupling constant is inferred from the hierarchical model~\cite{Easter2019} using the magnitude of the posterior waveforms, $|h_+(\boldsymbol{\theta},t)|$. The 68\% (dark blue) and 95\% (light blue) confidence intervals are shown along with the median values (blue dots). The true value for $\kappa_2^\textsc{t}$ is shown as the solid horizontal line. 
         The corresponding tidal deformability values are shown on the secondary vertical axis.}
         \label{fig:KappavsSNR}
     \end{figure}

   Figure~\ref{fig:KappavsSNR} shows the inferred posteriors for $\kappa_2^\textsc{t}$ with 68\% and 95\% confidence intervals in dark blue and light blue respectively. 
    The true injected value of $\kappa_2^\textsc{t}$, is shown by the horizontal solid black line and the median values as blue dots. 
    The  vertical axis shows both the quadrupolar tidal coupling constant (left axis) and the quadrupolar tidal deformability (right axis). 
    % The value of $\kappa_2^\textsc{t}$ inferred directly by the hierarchical model from the numerical-relativity simulation is shown by the horizontal black dotted line.
    The values inferred for the tidal parameters of the progenitor neutron stars are lower than the true value of the numerical-relativity injection, though the 95\% confidence interval excludes the true value only at a post-merger signal-to-noise ratio of 50.  
    The tidal coupling constant at a post-merger signal-to-noise ratio of 15 is  constrained at 95\% confidence intervals  to $68.5^{+5.9}_{-7.9}$, which tightens  to $68.5^{+3.4}_{-3.6}$ for  a post-merger signal-to-noise ratios of 50. \par
     There are a number of factors that will impact on the performance of the hierarchical model.
     Firstly, the numerical-relativity spectra from Ref.~\cite{Rezzolla2016}, which were used in Ref.~\cite{Easter2019} to train the model, are a distinct set of simulations to those in use in this paper~\cite{Dietrich2018}.
     Specifically, waveform SLy-M1.350-$\Lambda$390 is available in both sets of numerical-relativity simulations (Refs.~\cite{Dietrich2018} and ~\cite{Rezzolla2016}), and, although the primary post-merger peak occurs at the same frequency, the spectral response for the other frequencies are different (fitting factor $= 0.90$).
     Secondly, the simulation outputs can be dependent on the spatial and temporal resolution, which can lead to waveform changes related to parameters like collapse time, primary oscillation frequency and decay time constants. 
     To ensure the performance of the hierarchical model, the simulations used in the training set and test set should be consistent and not contradictory.
     Thirdly, the hierarchical model is an approximate model, and was  trained on 35 compatible waveforms; (for details, see Ref.~\cite{Easter2019}).
     Given the small training set, the hierarchical model results are consistent. 
     The complexity of the model could be increased as additional consistent numerical-relativity simulations become available. \par
      \begin{figure}[H]
         \centering
         \includegraphics[scale=0.5]{CompactnessPosterior.pdf}
         \caption{Compactness posteriors versus post-merger signal-to-noise ratio for numerical-relativity waveform SLy-M1.350-$\Lambda$390. The 68\% (dark blue) and 95\% (light blue) confidence regions are shown. The compactness is inferred from the hierarchical model~\cite{Easter2019}. The compactness inferred directly from the numerical-relativity waveform (dashed black line) and the compactness of the progenitor neutron stars (solid black line) are also shown. See the text for an explanation of the offset in these two values.}
         \label{fig:CompactnessvsSNR}
     \end{figure}  
     Hierarchical model posteriors for the compactness, $C$, are shown in Fig.~\ref{fig:CompactnessvsSNR}. 
     The 68\% and 95\% confidence intervals are shaded dark blue and light blue  respectively, and the median values are shown with blue dots.
     The true value corresponding to the injected numerical-relativity simulation is shown as a horizontal solid black line.
     The value inferred from the numerical-relativity simulation using the hierarchical model is shown as a horizontal black dashed line. 
     The hierarchical posteriors for $C$ are clustered around the value inferred directly from the hierarchical model for the numerical-relativity simulation.
     This is expected and due to the mismatch between the numerical-relativity injection and the set of numerical-relativity simulations used to train the hierarchical model.
     The compactness has been constrained to $0.162^{+0.007}_{-0.004}$ at post-merger signal-to-noise ratios of 15 tightening to  $0.164^{+0.002}_{-0.003}$ at signal-to-noise ratios of 50 to 95\% confidence intervals. 
     The posteriors for the compactness, $C$, only narrow moderately as the post-merger signal-to-noise ratio is increased.

\section{Discussion}
    We use an analytical model to characterise gravitational-wave strain from nine numerical-relativity simulations selected such that the post-merger oscillations persist for $\sim 25$\,ms. 
    The median noise-weighted fitting factors for the posterior waveforms range between $0.92\,$\nobreakdash-$\,0.97$ for injections with post-merger signal-to-noise ratios of 50. 
    This corresponds to a loss in detection rate of  $22\,$\nobreakdash-$\,12\%$ when compared to a signal without mismatch.
    We measure the Bayes factor in favour of signal detection with numerical-relativity simulation SLy-M1.350-$\Lambda$390 and find that successful detections occur with post-merger signal-to-noise ratios of $\ge 10$ with possible detections as low as  post-merger signal-to-noise ratios of 7, depending on the specific noise realisation. 
    This indicates that this model could be used for parameter estimation and detection if a post-merger signal louder than signal-to-noise ratio of 10 was  coincident with an inspiral detection. 
    We find that this corresponds to a distance of $\sim\,10$\,Mpc for an optimally oriented system using a three-detector network (LIGO Hanford, Livingston, and Virgo) at design sensitivity.
    \par
    We determine that starting the model at the time of coalescence results in the maximum matched filter signal-to-noise ratio even though the fitting factors are lower in the vicinity of the merger due to the dynamics of the nascent neutron star.
    We find that the uncertainty in the time of coalescence for the inspiral of the progenitor neutron stars is less that 0.1\,ms for a post-merger signal-to-noise ratio of $\ge\,6$ and show that this corresponds to a maximum matched-filter signal-to-noise ratio. \par

    The gravitational-wave strain of the inspiral can constrain the equation of state for the cold neutron star at the high inspiral signal-to-noise ratios ($\gtrsim 200$) required for post-merger detection of the remnant (see Fig.~\ref{fig:TCOALESCENCE}). 
    This can place additional constraints on the priors for the dominant post-merger frequency. 
    However, a phase transition in the hot post-merger remnant~\cite{Most2018b,Most2018,Bauswein2019}, and uncertainty in the numerical-relativity calculations due to computational trade-offs, may result in a post-merger gravitational-wave signal that is quantitatively different than the  numerical-relativity simulations.
    With this in mind, we assume a more general, agnostic set of priors (see Appendix~\ref{appendix:b2}).
    \par

    Using numerical-relativity waveform SLy-M1.350-$\Lambda$390, selected for its compatibility with $\Lambda$ values determined from GW170817~\cite[e.g.,][]{GW170817Detection,Annala2018,Radice2018,Most2018,De2018,GW170817Properties}, we constrain the primary post-merger frequency to a range of $3310$\,Hz$\pm_{1.2}^{1.4}\%$ for 95\% confidence intervals at post-merger signal-to-noise ratios of 15.
    The precision increases to $3296$\,Hz$\pm_{0.2}^{0.3}\%$ for post-merger signal-to-noise ratios of 50. \par
    We show that our model and \textsc{BayesWave} similarly constrain the dominant post-merger frequency, $f_{\mathrm{peak}}$, for post-merger signal-to-noise ratios of $\gtrsim 20$.
    For post-merger signal-to-noise ratio of $\sim 9 - 15$ \textsc{BayesWave} is better able to constrain $f_{\mathrm{peak}}$.
    We generate fitting factors of  $\approx 0.99$ using \textsc{BayesWave} for SLy-M1.350-$\Lambda$390 at a post-merger signal-to-noise ratio of 50.
    The corresponding fitting factors from our model are $\approx 0.92$.
    The dimensionality of the \textsc{BayesWave} posterior reconstruction is significantly larger than our analytic; $\sim 90$ dimensions for \textsc{BayesWave} cf. 15 for ours.
    Moreover, our adopted model is interpretable and can supply additional information about the individual modes (e.g. frequency drifts and exponential damping time constants).
     \par
    We use the hierarchical model from Ref.~\cite{Easter2019}, which has been trained on numerical-relativity waveforms from Ref.~\cite{Rezzolla2016}, to determine posteriors for $\kappa_2^\textsc{t}$ and $C$. 
    We obtain 95\% confidence intervals on $\kappa_2^\textsc{t}$ (and $\tilde{\Lambda}$) of $\pm^{9}_{12}\%$ at a post-merger signal-to-noise ratio of $15$ with increasing precision to $\pm\, 5\%$ at a post-merger signal-to-noise ratio of 50. 
    The 95\% confidence intervals on $C$ range from $\pm^{4.3}_{2.7}\%$ at post-merger signal-to-noise ratios of 15 to $\pm^{1.5}_{1.8}\%$ at post-merger signal-to-noise ratios of 50.
    However, the injected value for $C$ is outside the 95\% confidence interval. \par
    It should be noted that the inferred posteriors for $C$ are centred around the compactness value inferred directly from the numerical-relativity simulation.
    This indicates that the offset in the inferred compactness values is caused by the difference in the SLy numerical-relativity simulations between Ref.~\cite{Rezzolla2016} and Refs.~\cite{Dietrich2017b,Radice2016}.
    Contradictory training waveforms will increase uncertainties in the hierarchical model.
    The 35 numerical-relativity simulations used to train the hierarchical model were homogeneous, changing only the equation of state and the progenitor masses between simulations, keeping other simulation parameters the same. 
    It should also be emphasised that, because the numerical-relativity simulations are drawn from independent sources, the posteriors of $\kappa_2^\textsc{t} (\tilde{\Lambda})$ and $C$ are true out-of-sample estimates.
    We expect the estimates of $C$ and $\kappa_2^\textsc{t}$ to become more consistent with the injected value as the training set is increased in size and covers more system and progenitor properties. \par

    In addition to the aforementioned analytical models~\cite{Bauswein2016,Bose2018}, other work have generated analytical post-merger gravitational-wave  models. 
    In Ref.~\cite{Hotokezaka2013}, a model  was generated for the time-based amplitude and phase of the complex gravitational-wave strain using a smooth piece-wise function for the amplitude. 
    The time-based phase was fit by the combination of a polynomial and exponentially-damped sinusoid using an iterative CMA-ES (covariance matrix adaption evolution search) fitting algorithm.
    The maximum fitting factors were  calculated in the time domain without noise weighting and are not directly comparable to the noise-weighted fitting factors calculated with Eq.~\ref{eq:inner_product2}.
    The maximum fitting factors were  $\sim\!0.92\,$\nobreakdash-$\,0.98$ for 95\% of the 54 waveforms.\par

    A frequency-domain model was introduced in Ref.~\cite{Messenger2014} from analysing the major spectral peaks of the whitened power spectrum.
    The power of the dominant post-merger frequency peak was estimated by a trapezoidal structure and the model parameters were determined with a least-squares algorithm. 
    No fitting factors were calculated in this reference, as the goal was estimating source red-shifts.
    This model was extended in Ref.~\cite{Takami2015} to add a Gaussian component to the fundamental post-merger frequency using a nonlinear least-squares fit. 
    The goal of the fits in Ref.~\cite{Takami2015} were qualitative, rather than quantitative and no fitting factors were stated.\par

    The model used in Ref.~\cite{Bauswein2016} consists of three exponentially damped sinusoids centred at frequencies ($f_{2-0}, f_{\mathrm{spiral}}, f_{\mathrm{peak}})$ which are described in Section~\ref{sec:introduction2}.
    In contrast, the model introduced in Ref.~\cite{Bose2018}, consists of two exponentially damped sinusoids, the first centred on $f_1$ which is modulated by frequency $f_{1e}$, and the second is centred on the dominant post-merger frequency, $f_2$, with a linear and quadratic frequency drift terms. 
    This model produced fits of $\sim 80\Hyphdash*94\%$. 
    In Ref.\cite{Tsang2019}, a frequency-domain model was developed for a single damped-sinusoid.
    This model was based on three or six parameters and used Bayesian inference to estimate the parameters for 120 numerical-relativity simulations. 
    They obtained fitting factors of $\sim\!0.60\,$\nobreakdash-$\,0.98$. 
    Reference~\cite{Breschi2019} parameterised the instantaneous amplitude and phase of the time-based gravitational-wave strain using 172 numerical-relativity simulations. 
    Their model uses a rational-polynomial fit based on the progenitor properties ($M_1$, $M_2$, $\kappa_2^\textsc{t}$) derived in Refs.~\cite{Bernuzzi2015,Zappa2018,Dietrich2019}.
    They achieved fitting factors of $\sim\!0.30\,$\nobreakdash-$\,0.85$ in zero noise.\par
    The fitting factors obtained in our paper compare favourably to those listed above; our maximum fitting factors are above 0.93 for all waveforms \cite[cf.][]{Hotokezaka2013} and our minimum fitting factors are above 0.90 across all waveforms~\cite[cf.][]{Tsang2019,Breschi2019}. 
    The fitting factor is more sensitive to deviations in the time-based phase or Fourier phase response, than it is to amplitude deviations. 
    The fits in Ref.~\cite{Breschi2019} could possibly be improved by adding in more flexibility in the phase response.
    Our model bypasses the phase matching difficulty by directly fitting the phase with parameters, $(f_j, \alpha_j, \psi_j)$, from the injected signals from all three interferometers.
    Although Ref.~\cite{Tsang2019} does directly fit the phase, the first-order model is too restricted to obtain higher fitting factors and better results may be obtained by increasing the order of the model.
    The number of numerical-relativity simulations used in Refs.~\cite{Hotokezaka2013,Tsang2019,Breschi2019} was significantly larger than the nine simulations we use here, including additional mass ratios, spin configurations and eccentricity values. 
    We leave it for future work to expand the number of numerical-relativity simulations to test this model.\par

    Although numerical-relativity simulations currently provide the best estimate of the post-merger gravitational-wave strain, future post-merger signals may not be consistent with these state-of-the-art simulations.
    With this in mind, our model matches the numerical-relativity simulations well, but it is more flexible than these simulations.
    This is important because this method is a middle ground between  simulations of known waveforms, and more general (e.g. unmodelled excess power and \textsc{BayesWave}) methods. Nevertheless, numerical-relativity simulations are the primary method of investigating the dynamical physics of the post-merger region and  research into these simulations is vital.

\section*{Acknowledgments}
        P.D.L. is supported through Australian Research Council (ARC) Future Fellowship FT160100112, ARC Discovery Project DP180103155, and ARC Centre of Excellence CE170100004. A.R.C. is supported by ARC grant DE190100656. 
        S.G. and J.A.C. gratefully acknowledge the NSF for financial support from Grants No. PHY 1806580, PHY 1809572, and TG-PHY120016.
        The Flatiron Institute is supported by the Simons Foundation. 
        This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center (https://www.gw-openscience.org), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. LIGO is funded by the U.S. National Science Foundation. Virgo is funded by the French Centre National de Recherche Scientifique (CNRS), the Italian Istituto Nazionale della Fisica Nucleare (INFN) and the Dutch Nikhef, with contributions by Polish and Hungarian institutes.
        The authors are grateful for computational resources provided by the LIGO Laboratory and supported by National Science Foundation Grants PHY-0757058 and PHY-0823459. 
        This research was done using resources provided by the Open Science Grid \cite{Osg07, Osg09}, which is supported by the National Science Foundation award 1148698, and the U.S. Department of Energy's Office of Science.
        We are grateful to Sukanta Bose for valuable comments on the manuscript.

\raggedbottom
\begin{subappendices}
\renewcommand\thefigure{\thesection.\arabic{figure}} 
\renewcommand\thetable{\thesection.\arabic{table}} 
\setcounter{figure}{0} 
\setcounter{table}{0} 
\section{Numerical relativity simulations}
\label{appendix:a2}
We use nine simulations from the \texttt{CoRe} gravitational wave database~\cite{Dietrich2018} for binary neutron star mergers.
The simulations are listed by their equation of state, the progenitor mass, and the quadrupolar tidal deformability. 
We limit our simulations to those with equal-mass prognitors for compatibility with the heirarchical model in Ref.~\cite{Easter2019}. 
We choose simulations with the highest resolution such that the remnant was transmitting gravitational waves for $\sim 25$\,ms. 
In some cases increasing the resolution resulted in a reduced lifetime of the remnant. 
Table~\ref{tbl:NRwaveforms} shows the simulation designator for this paper, the name of the waveform in the \texttt{CoRe} database, and the citation for the associated simulation in the metadata (if available).
\begin{table}[H]
\centering
\caption{Numerical relativity simulations}\label{tbl:NRwaveforms}
 \begin{tabular}{|c|c|c|}
 \hline
 Designator & Simulation name & Citations \cite{Dietrich2018} \\  
 \hline
 \color{Waveform1}
 SLy-M1.350-$\Lambda$390 & \color{Waveform1} THC:0036:R03 & \color{Waveform1} \cite{Radice2016}  \\ 
 \hline
 \color{Waveform2}
 LS220-M1.350-$\Lambda$684 & \color{Waveform2} THC:0019:R05 & \color{Waveform2} \cite{Radice2017} \\
 \hline
  \color{Waveform3}
 MS1b-M1.500-$\Lambda$864 & \color{Waveform3} BAM:0088:R01 & \color{Waveform3} - \\
 \hline
  \color{Waveform4}
  BHBlp-M1.300-$\Lambda$1046 & \color{Waveform4} THC:0002:R01 & \color{Waveform4} \cite{Radice2017a,Radice2018} \\
 \hline
  \color{Waveform5}
 DD2-M1.250-$\Lambda$1295 & \color{Waveform5} THC:0011:R01 &
   \color{Waveform5} \cite{Radice2017a,Radice2018} \\
   \hline
    \color{Waveform6}
  MS1b-M1.375-$\Lambda$1389 & \color{Waveform6} BAM:0070:R01 &
   \color{Waveform6} \cite{Dietrich2017b} \\
   \hline
    \color{Waveform7}
   MS1b-M1.350-$\Lambda$1532 & \color{Waveform7} BAM:0065:R03 & \color{Waveform7}
  \cite{Bernuzzi2014} \\
  \hline
   \color{Waveform8}
  DD2-M1.200-$\Lambda$1612 & \color{Waveform8} THC:0010:R01 & 
  \color{Waveform8} \cite{Radice2017a,Radice2018} \\
  \hline
   \color{Waveform9}
  2H-M1.350-$\Lambda$2326 & \color{Waveform9} BAM:0002:R02 &
  \color{Waveform9} \cite{Bernuzzi2014}\\
 \hline
\end{tabular}
\end{table}
\section{Priors}
\label{appendix:b2}
    The priors are listed in Eqs.~\ref{eq:firstprior2}\Hyphdash*\ref{eq:w012} with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
    The mode number $j$ is limited to $\lbrace 0,1,2 \rbrace$ and the mode number $i$ is restricted to $\lbrace 0,1 \rbrace$. 
    The priors in Eqs.~\ref{eq:modesorting2}\Hyphdash*\ref{eq:w012} are constrained priors. 
    These restrictions are enforced in addition to the standard priors. The prior in Eq.~\ref{eq:modesorting2} ensures that the maximum spectral amplitude of each mode is decreasing. 
    This results in $f_0$ converging to the loudest peak. 
    \begin{eqnarray}
        \log_{10}{H} & \sim & \mathcal{U}(-24, -19)\label{eq:firstprior2}  \\
        f_j & \sim & \mathcal{U}(1000,5000) \\
        \log_{10}T_j & \sim & \mathcal{U}(-4.0,0.3)\\
        \psi_j & \sim & \mathcal{U}(-\pi,\pi) \\
        \alpha_j & \sim & \mathcal{U}(-6.4,6.4) \\
        w_i & \sim & \mathcal{U}(0.0,1.0)\label{eq:lastprior2}\\
        \log_{10}\left(\frac{\max|\tilde{h}_{j}(f)|_f}
        {\max|\tilde{h}_{j+1}(f)|_f}\right) & \sim & \mathcal{U}(0.0,10.0)\label{eq:modesorting2} \\
        w_0+w_1 & \sim & \mathcal{U}(0.0,1.0) \label{eq:w012}
    \end{eqnarray}
    $w_2$ is calculated as:
    \begin{equation}
        w_2 = 1 - w_0 - w_1,\label{eq:w0122}
    \end{equation}
    ensuring that $\sum_j w_j = 1$ and $w_2\in[0,1]$ as required. 
% \pagebreak
\section{Posteriors for all numerical-relativity injections}
\label{appendix:c2}  
Selected posteriors for all numerical-relativity simulations are shown in Fig.~\ref{fig:Corner1}-\ref{fig:Corner9}. 
The waveforms are injected at a post-merger signal-to-noise ratio of 50.
The posteriors shown are: $f_0, \alpha_0, f_1$ and $\alpha_1$. 
The posteriors are coloured as per Fig.~\ref{fig:FittingFactors} and Table~\ref{tbl:NRwaveforms}. 
The frequency drift term, $\alpha_0$, for the dominant post-merger frequency is negative for some simulations.
Most numerical-relativity simulations spectrograms show that the dominant frequency is constant or increases slightly over time, however, many simulations do show decreasing evolution of the dominant post-merger oscillation frequency~\cite[e.g.,][]{Foucart2016,Radice2017a,Dietrich2019b}.

\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{THC_0036_R03_0ms_corner}.pdf}
        \caption
        {Selected posteriors for numerical-relativity post-merger injection using the equation of state SLy with equal mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform SLy-M1.350-$\Lambda$390). The numerical-relativity simulation was injected at a post-merger signal-to-noise ratio of 50.} 
        \label{fig:Corner1}
\end{figure}

\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{THC_0019_R05_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state LS220 with equal mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform LS220-M1.350-$\Lambda$684).} 
        \label{fig:Corner2}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{BAM_0088_R01_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state MS1b with equal mass, $1.50\,\mathrm{M}_\odot$, neutron stars (waveform MS1b-M1.500-$\Lambda$864).} 
        \label{fig:Corner3}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{THC_0002_R01_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state BHBlp with equal mass, $1.30\,\mathrm{M}_\odot$, neutron stars (waveform BHBlp-M1.300-$\Lambda$1046).} 
        \label{fig:Corner4}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{THC_0011_R01_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state DD2 with equal mass, $1.25\,\mathrm{M}_\odot$, neutron stars (waveform DD2-M1.250-$\Lambda$1295).} 
        \label{fig:Corner5}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{BAM_0070_R01_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state MS1b with equal mass, $1.375\,\mathrm{M}_\odot$, neutron stars (waveform MS1b-M1.375-$\Lambda$1389).}
        \label{fig:Corner6}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{BAM_0065_R03_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state MS1b with equal mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform MS1b-M1.350-$\Lambda$1532).}
        \label{fig:Corner7}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{THC_0010_R01_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state DD2 with equal mass, $1.20\,\mathrm{M}_\odot$, neutron stars (waveform DD2-M1.200-$\Lambda$1612).}
        \label{fig:Corner8}
\end{figure}
\begin{figure}[H]
        \centering
        \includegraphics[scale=0.3]{{BAM_0002_R02_0ms_corner}.pdf}
        \caption
        {As per Fig.~\ref{fig:Corner1} using the equation of state 2H with equal mass, $1.35\,\mathrm{M}_\odot$, neutron stars (waveform 2H-M1.350-$\Lambda$2326).}
        \label{fig:Corner9}
\end{figure}
\end{subappendices}

\end{document}

### 2021-10-11 papers/paper3.tex
Initial Overleaf Import

% \documentclass[
% reprint,
% % prx,
% prd,
% twocolumn,
% % onecolumn,%!TEX encoding = UTF-8 Unicode
% nofootinbib,
% % ,secularistic%
% tightenlines %
% floatfix,
%  amsmath,
% showpacs ,amssymb, aps,% linenumber
% superscriptaddress
% ]{revtex4-2}
% \usepackage{amsmath,slashed}
% \usepackage[left]{lineno}
% \usepackage{units}
% \usepackage{comment}
% \usepackage{graphicx}
% \usepackage{amsmath,amssymb}
% \usepackage{support-caption}
% \usepackage{subcaption}
% \usepackage{bm}
% \usepackage[shortlabels]{enumitem}
% \usepackage{float}
% \usepackage{soul}
% \usepackage{natbib}
% \usepackage{wasysym}
% \usepackage{aas_macros}
% \graphicspath{{images/}}
% \RequirePackage{lineno}
% \usepackage[usenames,dvipsnames,table]{xcolor}
% \usepackage{appendix}
% \usepackage{extdash}
% \usepackage{colortbl}
% \usepackage{tikz}

% \usepackage{booktabs}

% \usepackage[printwatermark]{xwatermark}

% % \usepackage{titlesec}

% \definecolor{lightblue}{HTML}{1F77B4}
% \definecolor{grey}{HTML}{808080}
% \newcommand{\SLyWaveform}[1]{\textcolor{lightblue}{#1}}
% \newcommand{\grey}[1]{\textcolor{grey}{#1}}
% \newcommand{\red}[1]{\textcolor{red}{#1}}
% \newcommand{\pdl}[1]{\textcolor{red}{[PDL: #1]}}
% \newcommand{\pje}[1]{\textcolor{Purple}{[PJE: #1]}}
% \newcommand{\blue}[1]{\textcolor{blue}{#1}}

% \DeclareMathOperator*{\argmin}{arg\,min}
% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Vic 3800, Australia}
% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
% \newcommand{\AstroThreeD}{ARC Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), Canberra, ACT 2611, Australia}

% \captionsetup{belowskip=-8pt,aboveskip=0pt}
% % \titlespacing*{\section}
% % {0pt}{5.5ex plus 1ex minus .2ex}{4.3ex plus .2ex}

% \definecolor{Waveform1}{HTML}{1f77b4}
% \definecolor{Waveform2}{HTML}{ff7f0e}
% \definecolor{Waveform3}{HTML}{2ca02c}
% \definecolor{Waveform4}{HTML}{d62728}
% \definecolor{Waveform5}{HTML}{9467bd}
% \definecolor{Waveform6}{HTML}{8c564b}
% \definecolor{Waveform7}{HTML}{e377c2}
% \definecolor{Waveform8}{HTML}{7f7f7f}
% \definecolor{Waveform9}{HTML}{bcbd22}

% % \newsavebox\mybox
% % \savebox\mybox{\tikz[color=red,opacity=0.3]\node{DRAFT};}
% % \newwatermark*[
% %   allpages,
% %   angle=45,
% %   scale=6,
% %   xpos=-20,
% %   ypos=15
% % ]{\usebox\mybox}

\documentclass[../Thesis.tex]{subfiles}

\begin{document}

\begingroup
\clearpage% Manually insert \clearpage
\let\clearpage\relax% Remove \clearpage functionality
\vspace*{-2cm}% Insert needed vertical retraction

\chapter{Can we measure the collapse time of a post-merger remnant for a future GW170817-like event?}
\label{chapter:CollapseTime}
\endgroup

\chapterauthor{Paul J. Easter\\Paul D. Lasky\\Andrew R. Casey}

% \author{Paul J. Easter}
% \email{paul.easter@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Paul D. Lasky}
% \email{paul.lasky@monash.edu}
% \affiliation{\SPA}
% \affiliation{\OzGravMonash}
% \author{Andrew R. Casey}
% \email{andrew.casey@monash.edu}
% \affiliation{\SPA}
% \affiliation{\AstroThreeD}

% \pacs{
% }
% \enlargethispage{-3.2cm}
\section*{Abstract}

    Measuring the collapse time of a binary neutron star merger remnant can inform the physics of extreme matter and improve modelling of short gamma-ray bursts and associated kilonova.
    The lifetime of the post-merger remnant directly impacts the mechanisms available for the jet launch of short gamma-ray bursts.
    We develop and test a method to measure the collapse time of post-merger remnants.
    We show that for a GW170817-like event at $\sim\!40$\,Mpc, a network of Einstein Telescope with Cosmic Explorer is required to detect collapse times of $\sim\!10$\,ms.
    For a two-detector network at A+ design sensitivity, post-merger remnants with collapse times of $\sim\!10\,\mathrm{ms}$ must be $\lesssim 10$\,Mpc to be measureable.
    This increases to $\sim\!18\Hyphdash*26$\,Mpc if we include the proposed Neutron star Extreme Matter Observatory (NEMO), increasing the effective volume by a factor of $\sim\!30$.

% \maketitle
    \vspace{-0.25cm}
    \section{Introduction}
    \label{sec:introduction3}

    Measuring the lifetimes of binary neutron star post-merger remnants can help probe extreme matter at high temperature and densities and narrow down the physics of short gamma-ray bursts~\cite[e.g.,][]{Zhang2019,Ciolfi2020a}.
    These remnants may promptly collapse into a black hole or form a hot, differentially-rotating neutron star.
    If the total mass of a remnant is between 1.2 and 1.5 times the maximum non-rotating neutron star mass (the Tolman-Oppenheimer-Volkov mass), then the remnant is known as a hypermassive neutron star~\cite{Tolman1939,Oppenheimer1939,Breu2016,Weih2018,Baumgarte2000}, which is expected to collapse to form a black hole in a timescale from milliseconds to seconds \cite{Paschalidis2012}. 
    For smaller masses, the differentially-rotating remnant will evolve into rigidly-rotating neutron star after the differential rotation is quenched.
    The rigidly-rotating remnant will either collapse to a black hole, or form a stable neutron star, depending on the remnant mass.
    See Ref.~\cite{Sarin2021} for a recent review on the evolution of neutron star merger remnants.

    Determining the collapse times of hypermassive remnants can help narrow down the nature of the central engine for short gamma-ray-bursts. 
    Multi-messenger observations of binary neutron star merger GW170817 suggest that the remnant may have either collapsed to a black hole \cite[e.g.,][]{Metzger2018,Gill2019,Murguia-Berthier2021}, or formed a long-lived remnant~\cite[e.g.,][]{Yu2018}.
    Measuring the collapse time of a remnant may help determine the characteristic timescales associated with short gamma-ray-bursts, aiding the selection of the central engine (for a review see \cite{Zhang2019}).
    Furthermore, measuring the collapse time of the post-merger remnant may help constrain the quenching mechanism and physics of the differential rotation, which may reveal indicators towards the relative contribution of radiative (gravitational waves and neutrino) and dissipative (viscous, resistive and magnetic braking) processes within the remnant.\par

    The direct detection of gravitational waves from future neutron star merger remnants presents a great opportunity to constrain the collapse time.
    Numerical-relativity simulations of merger remnants show gravitational waves predominantly emitted from the fundamental f-mode oscillation of the remnant~\cite{Zhuge1994,Stergioulas2011}.
    Gravitational waves emitted from this mode occur at $\sim\!1.8\Hyphdash*4$\,kHz~\cite{Takami2015,Bernuzzi2015}.
    No post-merger remnant was detected for GW170817 by the LIGO and Virgo collaboration due to lack of sensitivity of the detectors at these frequencies. 
    However, increased sensitivity of gravitational-wave instruments and targeted high-frequency detectors  may enable future detections of post-merger remnants~\cite[e.g.,][]{Martynov2019,NEMO2020}. \par
    \begin{figure*}
        \centering
        \includegraphics[scale=0.28]{dec_tc10_dist_timeplot_panels.pdf}
        \caption{Time domain posterior waveforms for SLy equation of state with equal-mass 1.35M${_\odot}$ progenitor (left panels) and measured collapse times (right panels) for injections with collapse times $\tc=5$\,ms~(top), $\tc=10$\,ms~(centre), and $\tc=15$\,ms~(bottom) injected into 2A+ detector network. Left panels show the numerical-relativity injections that have been forced to collapse (black curves) with the posterior waveforms (shaded blue curves). The right panels show the collapse-time posterior distributions for each numerical-relativity injection.
        The solid vertical lines on all panels shows  $\tc$ for the injected waveform. The dashed vertical lines shows  $\tc+\tw$ where the injected signal drops to zero due to the induced collapse. Here we show three loud injections from our injection study where $\tc$ are clearly recovered. The injected distances are 5.93\,Mpc, 3.04\,Mpc, and 1.00\,Mpc, for 5\,ms,  10\,ms, and  15\,ms respectively. The full posteriors for the $\tc=10\,$ms collapse time   injection are shown in the appendix (Fig~\ref{fig:FullCorner}).} 
        \label{fig:TimeResponse}
\end   {figure*}
\pagebreak
    In this paper, we assess how well future networks of gravitational-wave detectors can measure the collapse time of a post-merger remnant.
    We extend a waveform model developed in Ref.~\cite{Easter2020} which was derived from Refs.~\cite{Bauswein2016,Bose2018}, to allow the measurement of the collapse time of the post-merger remnant.
    Using Bayesian inference, we inject numerical-relativity gravitational waveforms that are forced to collapse into different interferometer configurations to measure the maximum distance at which we can recover the collapse time.\par

    This paper is laid out as follows. In Section~\ref{sec:methodology3} we outline the extended waveform model along with the method of gravitational-wave injections.
    The results are outlined in Section~\ref{sec:results3} and their implications in Section~\ref{sec:discussion3}.

    \vspace{-0.25cm}
    \section{Methodology}\label{sec:methodology3}
    We use numerical-relativity waveforms with two different equations of state that we inject into Gaussian noise realisations determined by the gravitational-wave interferometer configuration.
    We modify the numerical-relativity waveforms to collapse at a given collapse time.
    We then use an analytic model to perform detection and parameter estimation to determine distributions of the model parameters.
    This model, based on two models in Refs.~\cite{Bauswein2016,Bose2018}, is outlined in Ref.~\cite{Easter2020}.
    See also Refs.~\cite{Tsang2019,Breschi2019} for alternative models of the post-merger gravitational-wave strain.

        The gravitational waves from the post-merger remnant are modelled as a third-order, exponentially-damped sinusoid with a linear frequency-drift term. The plus polarisation of the gravitational-wave strain, $h_{+}(\boldsymbol{\theta},t)$, is given by: 
\begin{align}
        h_{+}(\boldsymbol{\theta},t) & = \sum_{j} h_{j,+}(\boldsymbol{\theta},t),\nonumber\\
        h_{j,+}(\boldsymbol{\theta},t) & = 
        H w_j \exp\left[-\frac{t}{T_j}\right] \cos \left(2\pi f_j t\left[1+\alpha_j t\right]+\psi_j \right),\nonumber\\
        \boldsymbol{\theta} & = \{H,w_j,T_j,f_j,\alpha_j,\psi_j :  j \in [0,2] \},\label{eq:hplus3}
\end{align}
    where $\boldsymbol{\theta}$ are the model parameters and $\sum_{j} w_j=1$. 
    Here, $H$ is an overall amplitude scaling factor, $w_j$ is the relative amplitude of the $j$th mode, $T_j$ is the corresponding exponential damping time constant, and $f_j$ is the corresponding frequency.
    The linear frequency-drift term is $\alpha_j$ and the initial phase  is $\psi_j$.
    The time, $t$, is defined, such that $\max\left[h_+^2(\boldsymbol{\theta},t)+h_\times^2(\boldsymbol{\theta},t)\right]$ occurs when $t=0$. 
    The cross polarisation for Eq.~\ref{eq:hplus3} is found by applying a $\pi/2$ phase shift to $\psi_j$.
    More details on this model can be found in Ref.~\cite{Easter2020}.

    We extend this model by introducing a collapse in the gravitational-wave strain using the falling edge of a Tukey window.
    The falling edge starts at the collapse time, $\tc$, and by the time $t=\tc+\tw$ the gravitational-wave strain drops to zero (see Fig.~\ref{fig:TimeResponse}). 
    Here $\tw$ is the time taken for the gravitational-wave signal to completely decay.
    % to completely collapse once collapse begins
    We have chosen $\tw=2$\,ms from examining numerical-relativity simulations with collapsing remnants (e.g., simulation labels: \texttt{BAM:0109:R01}, \texttt{BAM:0110:R01}, \texttt{BAM:0111:R02}~\cite{Dietrich2018}, \texttt{BAM:0044:R02}, \texttt{BAM:0045:R01}~\cite{Dietrich2017a}, \texttt{BAM:0124:R01}~\cite{Dietrich2017b}).
    This model captures the reduction in amplitude of the post-merger gravitational-wave signal as the remnant collapses.
    The full gravitational-wave strain including the collapse of the remnant, $h_{\varphi,\mathrm{c}}(\boldsymbol{\tilde{\theta}},t)$, is given by:
\begin{align}
        & h_{\varphi,\mathrm{c}}(\boldsymbol{\tilde{\theta}},t)    =  h_{\varphi}(\boldsymbol{\theta},t)\mathcal{T}\left(t , \tc,\tw\right), \label{eq:hcol3}\\ 
        \mathcal{T} & (t,\tc,\tw) =    \nonumber  \\
        &\begin{cases} 
            1 &\text{if}\ t \leqslant  \tc \\ 
            \frac{1}{2}\left(\cos\left(\frac{\pi (t-\tc)}{\tw}\right) +1 \right)  &\text{if}\ \tc < t \leqslant \tc+\tw\ \text{.} \label{eq:tukeyfallingedge3} \\
            0  &\text{if}\  t > \tc+\tw\\ 
        \end{cases}
\end{align}
    Here, $\varphi\in\{+,\times\}$ are the plus and cross gravitational-wave polarisations, and $\boldsymbol{\tilde{\theta}}= \boldsymbol{\theta} \cup \{\tc,\tw\}$. \par

    To generate the collapsing gravitational-wave signal injection, $s_{\varphi,\mathrm{c}}(t)$, we apply the Tukey collapse window to the numerical-relativity simulation, $h_{\mathrm{NR},\varphi}(t)$, as follows: 
    \begin{align}
        s_{\varphi,\mathrm{c}}(t) & =  h_{\mathrm{NR},\varphi}(t)\mathcal{T}\left(t , \tc,\tw\right). \label{eq:scol3}
    \end{align}
    To ensure that $s_{\varphi,\mathrm{c}}(t)$ has the required collapse time, $h_{\mathrm{NR},\varphi}(t)$ must emit post-merger gravitational waves for $t>\tc+\tw$.
    In this paper we use two numerical-relativity simulations of binary neutron star mergers with equal-mass $1.35\,\mathrm{M}_\odot$ progenitors that emit gravitational waves for $\sim~25$\,ms and sample collapse times of 5, 10, and 15\,ms.
    The two simulations use the SLy equation of state~\cite{Douchin2001} (simulation label \texttt{THC:0036:R03} from Refs.~\cite{Dietrich2018,Radice2016}) and LS220 equation of state~\cite{Lattimer1991} (simulation label \texttt{THC:0019:R05} from Refs.~\cite{Dietrich2018,Radice2017}).
    The dimensionless tidal-deformabilities of the progenitor neutron stars for the two simulations are 390 and 684 respectively, which are consistent with tidal deformabilities inferred from GW170817~\cite{GW170817Properties}.
    The SLy equation of state is the softer of the two equations of state, with a lower tidal deformability, more compact remnant structure, and higher dominant oscillation frequency.\par

    We use four different detector networks in this injection study.
    Firstly, a two detector network at A+ design sensitivity (2A+) located at existing LIGO sites: Hanford, Washington; and Livingston, Louisiana~\cite{Abbott2020Prospects,PSD:Aplus}. 
    Secondly, we add the proposed Neutron star Extreme Matter Observatory (NEMO), located at Gingin, Western Australia to the first network~\cite{NEMO2020}.
    Thirdly, we use the Einstein Telescope (ET)~\cite{Hild2008,Punturo2010,Hild2011, PSD:ET}.
    And finally, the Einstein Telescope with an additional interferometer, located at Hanford, Washington, with a Cosmic Explorer (CE) power spectral density~\cite{Abbott2017b, Adhikari2019,PSD:CE}.
    We use fixed random seeds for Gaussian noise generation for each interferometer and inject numerical-relativity simulations at a sky position corresponding to a mean sky signal-to-noise ratio. 
    \par

    We use \textsc{Bilby}, a Bayesian inference package~\cite{Ashton2019}, to obtain posteriors, $p(\boldsymbol{\tilde{\theta}}~|~s_{\varphi,\mathrm{c}}(t))$ from a numerical relativity injection with enforced collapse starting at $t=\tc$ with width $\tw=2$\,ms.
    Posteriors are calculated using waveforms detailed in Eqs.~\ref{eq:hplus3}\Hyphdash*\ref{eq:tukeyfallingedge3}. 
    See Ref.~\cite{Easter2020} for further details of the gravitational-wave likelihood, and Appendix~\ref{appendix:a3} for additional information on the priors.
    \par
    \pagebreak
    To deem the collapse time as successfully recovered, we demand the following to hold: 
\begin{align}
    \tcl - 2\,\mathrm{ms} \leqslant \tc & \leqslant  \tcu + 2\,\mathrm{ms},\\
    \tcu - \tcl & \leqslant  5\,\mathrm{ms},
\end{align}
    where $\tc$ is the injected value. 
    Here $\tcu$ and $\tcl$ are the upper and lower 68 percentile credible intervals on the posterior,  $p(\tc)$.
    Finally, the Bayes Factor for the ratio of evidence for signal against evidence for noise must be $\gg 1$.
    The minimum successful Bayes Factor in favour of a signal over noise in this paper is $\sim\!500$.
    These requirements ensure that successfully recovered $\tc$ are within a few milliseconds of the  injected values. 
    Although this method is somewhat arbitrary, it successfully identifies injections where the collapse time is recovered.
    Furthermore, as the results are pessimistic, with detections not expected until Cosmic Explorer and Einstein Telescope are online, changing this selection criterior will not substantially change these results. \par

    We perform numerical-relativity injections with the full waveform, Eq.~\ref{eq:scol3}, at a grid of distances and apply the above rules to determine whether we successfully recover $\tc$.
\begin{figure*}
    \centering
    \includegraphics[scale=0.24]{MaximumDistancePanelLabelledGW170817.pdf}
    \caption{Maximum distances for which the collapse time can be measured for different interferometer networks and equations of state. The vertical axis shows different equations of state and interferometer configurations, and the maximum detection distance is shown on the horizontal axis. The numerical-relativity simulations are injected with equations of state SLy (upward pointing triangle) and LS220 (downward pointing triangle). The interferometer configurations are 2A+ (blue), 2A+ with the proposed Neutron star Extreme Matter Observatory (orange), Einstein Telescope (green), and Einstein Telescope with Cosmic Explorer~(red). The top panel shows collapse times of 5\,ms, the centre panel 10\,ms and the lower panel 15\,ms collapse times. The  luminosity distance for GW170817 (gravitational-wave only) is shown in shaded grey for comparison~\cite{GW170817multi}.} 
    \label{fig:CollapseTimeDistanceGrid}
\end{figure*} 
\vspace{-0.25cm}    
\section{Results}\label{sec:results3}
    We inject post-merger numerical-relativity waveforms modified to collapse at $t=\tc$ and sample posteriors for $\boldsymbol{\tilde{\theta}}$.
    We then calculate posterior waveforms, $h_{\varphi,c}(\boldsymbol{\tilde{\theta}},t)$, from Eqs.~\ref{eq:hcol3}\Hyphdash*\ref{eq:tukeyfallingedge3}.
    Fig.~\ref{fig:TimeResponse} shows example posterior waveforms for the plus polarisation (left panels) and collapse-time posterior distributions (right panels) for numerical-relativity injections using an SLy equation of state with equal-mass $1.35\,\mathrm{M}_\odot$ progenitor neutron stars.
    The three panels have $\tc=5$\,ms (upper panels), $\tc=10$\,ms (centre panels), and $\tc=15$\,ms (lower panels).
    The left panels show the numerical-relativity injection (plus polarisation) in black and the posterior waveforms in blue.
    The vertical black lines show the beginning, $t=\tc$ (solid), and the end, $t=\tc+\tw$ (dashed), of the collapse for the injected signal.
    We perform injections into the 2A+ detector network at a grid of distances  and from these injections we select three distances where we can clearly recover the collapse time (see Sec.~\ref{sec:methodology3}).
    The injection distances are 5.93\,Mpc for 5\,ms, 3.04\,Mpc for 10\,ms, and 1.00\,Mpc for 15\,ms.
    The right panels show posteriors, $p(\tc)$, in shaded blue, along with the true injected value, $\tc$, as solid vertical black lines.
    The model successfully recovers both the collapse time and the complex nature of the numerical-relativity injection, for all three injections.
    For reference, the full posteriors for $\tc=10\,$ms are shown in the appendix (Fig.~\ref{fig:FullCorner}).\par

\par
    In Fig.~\ref{fig:CollapseTimeDistanceGrid}, we show the maximum distance for which we can recover the collapse time for the post-merger remnant.
    The numerical-relativity injections are performed for SLy (upward triangles) and LS220 (downward triangles) equations of state.
    We inject into the following interferometers described in Sec.~\ref{sec:methodology3}: 1) 2A+ (blue), 2) 2A+ and the proposed NEMO (orange), 3) Einstein Telescope (green), and 4) Einstein Telescope with Cosmic Explorer (red).
    We calculate the signal-to-noise ratio at a fixed distance over the entire sky for each detector network.
    We then choose a sky position with a signal-to-noise ratio close to the mean all-sky signal-to-noise ratio, and perform all injections at this sky position for this detector network.
    The lower error bars show the largest distance where the collapse times are recovered, the upper error bars show the smallest distance where the collapse time recoveries fail, and the marker is placed in the midpoint between these distances.
    If a post-merger collapse event occurs at a sky location near the antenna pattern maximum then the maximum distance where we can measure $\tc$ will increase by a factor of $\sim\!0.6$.
    \par

    Detecting the collapse time of a GW170817-like event at a luminosity distance of $40^{+8}_{-14}$\,Mpc~\cite{GW170817multi} (Fig.~\ref{fig:CollapseTimeDistanceGrid}, shaded region, gravitational wave only) would require the combination of ET with CE for $\tc \sim\!10$\,ms,  or ET with $\tc \sim\!5$\,ms. 
    The detection distance for 2A+ with  $\tc\sim\!10$\,ms is $\sim\!6\Hyphdash*8$\,Mpc.  
    This reduces to $\sim\!1.5\Hyphdash*2.0$\,Mpc for $\tc \sim\!15$\,ms. 
    Adding the NEMO high frequency detector increases the detection distance to $\sim\!17\Hyphdash*31$\,Mpc  for $\tc \sim\!10$\,ms.
    The detection distances for $\tc\sim\!15$\,ms with 2A+ and NEMO interferometers are $\sim\!3$\,Mpc and $\sim\!8$\,Mpc, for SLy and LS220 injections, respectively.\par

    For most collapse times and interferometer configurations, the ratio of the detection distance for LS220 to SLy injections is around $\sim\!1.2\approx f_0(\mathrm{SLy}) / f_0(\mathrm{LS220})$ ($f_0$ is the dominant post-merger oscillation frequency) which is consistent with SLy being softer and more compact than the LS220 equation of state. 
    For $\tc\!\sim15$\,ms injections into either 2A+ with NEMO, ET, or ET with CE, interferometer networks, the ratio of the LS220 to SLy detection distance increases to $\sim\!3$.
    Specifically, for $\tc\!\sim15$\,ms injections, detection distances of $\sim\!8,16,23$\,Mpc are found for LS220 equation of state with interferometers 2A+ with NEMO, ET, and, ET with CE, respectively.
    The corresponding detection distances for SLy injections are $\sim\!3,5,7$\,Mpc. 
    \par
    For injections where we can recover the collapse time, the dominant post-merger frequency is well constrained at $f_0 \lesssim 2\%$, $f_0 \lesssim 1\%$, and $f_0 \lesssim 0.2\%$ for injections of $\tc=5$\,ms, $\tc=10$\,ms and $\tc=15$\,ms, respectively. 
    Finally, we find no significant correlations between $\tc$ and other model parameters.
    We also attempt to measure the maximum detection distance where we can recover $\tc=20$\,ms and find that limitations in the third-order exponentially-damped sinusoidal model, Eq.~\ref{eq:hplus3}, prevent recovery of such collapse times.
    For $\tc > 15$\,ms signals, the analytical model cannot successfully track the time-domain phase of the gravitational-wave strain.
    This limitation could be overcome by increasing the complexity of the frequency evolution in Eq.~\ref{eq:hplus3}, possibly introducing a quadratic frequency evolution term~\cite{Bose2018}.
    Additionally, unmodelled searches such as \textsc{BayesWave} could be modified to measure the collapse time of the remnant~\cite{Cornish2015,Littenberg2015,Chatziioannou2017,Torres-Rivas2019}.\par
\vspace{-0.25cm}
\section{Discussion}\label{sec:discussion3}

    We inject post-merger gravitational-wave signals that have been modified to collapse at varying distances into four different interferometer configurations: 2A+, 2A+ with NEMO, ET and ET with CE.
    We perform injections with collapse times of 5, 10, and 15\,ms, and recover collapse-time posteriors.
    The injected gravitational-wave strain is recovered with a third-order exponentially damped sinusoid with a linear frequency-drift term~\cite{Easter2020} that has been modified to collapse at $t=\tc$. \par

    To measure the collapse time of a post-merger remnant in a GW170817-like event (gravitational-wave only, luminosity distance of $40^{+8}_{-14}$\,Mpc~\cite{GW170817multi}), we find that we need interferometer configurations of either ET, or ET with CE, for $\tc \sim\!10$\,ms with the exclusion of ET with SLy equation of state.

    We show that, for each detector network, the maximum detection distance where we can measure 5\,ms collapse times is similar to the maximum detection distance for 10\,ms collapse times, with maximum detection distances of: $\sim\!6\Hyphdash*8$\,Mpc for 2A+, $\sim\!18\Hyphdash*26$\,Mpc for 2A+ with NEMO, $\sim\!23\Hyphdash*37$\,Mpc for ET, and $\sim\!33\Hyphdash*53$\,Mpc for ET with CE.\par

    We find that the stiffer equation of state, LS220, has more energy in the post-merger gravitational wave at larger times after the merger.
    This leads to larger maximum detection distances for LS220 equations of state relative to SLy injections for $\tc\sim\!15$\,ms.
    The maximum detection distance for each detector network for $\tc\sim\!15$\,ms are $\sim 1.5, 2.7, 5.4, 7.4$\,Mpc for SLy injections, and $\sim 2.0, 8.0, 16, 23$\,Mpc for LS220 injections, for 2A+, 2A+ with NEMO, ET, and ET with CE, detectors respectively.
    The above distances assume an injection at a sky position corresponding to an average signal-to-noise ratio over the entire sky.
    The detection distance would increase by a factor of $\sim\!0.6$ near an optimal sky position. \par

    We find that there are three predominant regions for detecting the collapse time. The first region, with small collapse times, is mainly limited by the Bayes Factor for the ratio of post-merger signal to noise. 
    For large collapse times, waveform systematics limit detections, specifically the inability of the model to track the phase of the gravitational-wave strain.
    Between these two regions the signal-to-noise ratio is the limiting factor.\par

    Ignoring waveform systematics, Ref.~\cite{Zhang2021} found that they could achieve a signal-to-noise ratio of 0.5\Hyphdash*8.6 for a collapse time of 10\,ms for a post-merger gravitational-wave signal at 50\,Mpc.
    The model used was a single-order damped sinusoid injected into a high-frequency detector.
    The authors used a TM1 equation of state with two equal-mass 1.35\,M$_\odot$ progenitors with a dominant post-merger frequency of $\approx 2.8$\,kHz which very similar to $f_0$ for the LS220 equation of state in this paper.
    We find in this paper that we require a post-merger signal-to-noise ratio of $\gtrsim 17$ to successfully recover $\tc\sim\!10$\,ms for LS220 equation of state when waveform systematics are considered.

    With an estimated binary neutron star merger rate of $320^{+490}_{-240}\,\mathrm{Gpc^{-3}yr^{-1}}$~\cite{PopGWTC2}, it is unlikely that the collapse time of a post-merger remnant will be detected before either Cosmic Explorer or Einstein Telescope are operating at design sensitivity.
    When Cosmic Explorer and Einstein Telescope are both operating we may detect post-merger collapse times of $\sim\!10$\,ms.
    If only Einstein Telescope is fully operating then we may potentially measure post-merger collapse times of $\sim\!10$\,ms except for soft equations of state like SLy.
    In the mean time we will need to rely on indirect estimates of the post-merger collapse time that depend on multi-messenger observations~\cite[e.g.,][]{Metzger2018,Gill2019,Murguia-Berthier2021,Yu2018}.
    However, if a GW170817-like event occurred near an optimal sky position there would be a 60\% increase in the detection distance.
    In this case $\tc \sim\!10$\,ms may be detectable for ET, and ET with CE, for both equations of state.
    Additionally, 2A+ with NEMO would also be detectable for $\tc \sim 10$\,ms in this situation.
    It may also be possible to detune the proposed NEMO high frequency detector to increase sensitivity in the post-merger frequency band.
    This could potentially increase the sensitivity of the NEMO detector by a factor of $\sim\!1.6$ which would be enough to allow the NEMO detector with 2A+ to detect a GW170817-like post-merger collapse for $\tc \sim\! 10$\,ms. \par
    Finally, these results are dependent on the decay characteristics of the numerical-relativity simulations used in this paper.
    If the amplitude of future post-merger gravitational-waves have significantly longer decay timescales than the numerical-relativity simulations used here, then it is conceivable that larger collapse times could be measured.
    However, in this case waveform systematics become more important and models will either need to successfully track the waveform phase, or rely on incoherent methods that are independent of the phase of the gravitational-wave strain, or use unmodelled coherent detection methods~\cite[e.g.,][]{GW170817Postmerger1,GW170817Postmerger2}.

\section{Acknowledgments}
        P.~D.~L. is supported through Australian Research Council (ARC) Future Fellowship FT160100112,  ARC Discovery Project DP180103155 and  ARC Centre of Excellence CE170100004. A.~R.~C. is supported in part by the Australian Research Council through a Discovery Early Career Researcher Award (DE190100656). Parts of this research were supported by the Australian Research Council Centre of Excellence for All Sky Astrophysics in 3 Dimensions (ASTRO 3D), through project number CE170100013. Parts of this work were performed on the OzSTAR national facility at Swinburne University of Technology. The OzSTAR program receives funding in part from the Astronomy National Collaborative Research Infrastructure Strategy (NCRIS) allocation provided by the Australian Government. The authors wish to acknowledge the public release of the Computational Relativity data at \url{http://www.computational-relativity.org}. We are grateful to Tim Dietrich for valuable comments on this manuscript.
% \bibliography{CollapseTimeBib}

% \pagebreak
% \raggedbottom
% \pagebreak
\begin{subappendices}

% \raggedbottom
\renewcommand\thefigure{\thesection.\arabic{figure}} 
\renewcommand\thetable{\thesection.\arabic{table}} 
\setcounter{figure}{0} 
\setcounter{table}{0}

% GW170817 40/+8/-14= 26-48

\section{Priors}
\label{appendix:a3}
    The priors are listed in Eqs.~\ref{eq:firstprior3}\Hyphdash*\ref{eq:w013} with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
    The mode number $j$ is limited to $\lbrace 0,1,2 \rbrace$. 
    The priors in Eqs.~\ref{eq:freqsorting3}\Hyphdash*\ref{eq:w013} are constraining priors. 
    These restrictions are enforced in addition to the standard priors. 
    The priors in Eqs.~\ref{eq:freqsorting3}\Hyphdash*\ref{eq:modesorting3} sort the maximum spectral amplitude of each mode which improves computational stability and mode identification.
    See Ref.~\cite{Easter2020} for more details on mode sorting.
    We find that correlations between $\tc$ , $\tw$, and $T_0$ (the exponential decay time-constant for mode zero) make it very difficult to recover all three parameters simultaneously, even with analytic injections into zero noise.
    Fixing $\tw=2$\,ms allows recovery of all other parameters in both analytical injections with zero noise, and numerical-relativity injections with Gaussian noise.
    % \vfill\null

    \begin{eqnarray}
        \log_{10}{H} & \sim & \mathcal{U}(-24, -17)\label{eq:firstprior3}  \\
        f_j & \sim & \mathcal{U}(1000,5000)\  [\mathrm{Hz}]\\
        \tc & \sim & \mathcal{U}(0,100)\  [\mathrm{ms}]\\
        \tw & = & 2\  [\mathrm{ms}]\\
        T_j & \sim & \mathcal{U}(0,100)\ [\mathrm{ms}]\\
        \psi_j & \sim & \mathcal{U}(-\pi,\pi) \\
        \alpha_j & \sim & \mathcal{U}(-6.4,6.4)\   [\mathrm{Hz}] \\
        w_0, w_1 & \sim & \mathcal{U}(0,1)\label{eq:lastprior3}\\
        f_{j}  & > & f_{j+1}\label{eq:freqsorting3}\\
        \max|\tilde{h}_{j}(f)|_f & > & \max|\tilde{h}_{j+1}(f)|_f \label{eq:modesorting3}\\
        w_0+w_1 & < & 1  \label{eq:w013}\\
        w_2 & = & 1 - w_0 - w_1 \label{eq:w23}
    \end{eqnarray}
    % \onecolumngrid
    \pagebreak
    % \vfill\null 
    % \raggedbottom
    % \pagebreak

\section{Example posteriors}
\label{appendix:b3}
    \vspace{-0.3cm}
    Figure~\ref{fig:FullCorner} shows the posteriors for a post-merger numerical-relativity injection with $\tc=10\,$ms and SLy equation of state with equal-mass $1.35\,\msun$ neutron stars. The injections are performed at a distance of 3.04\,Mpc into a detector network of 2A+.
    These posteriors correspond to the $\tc=10\,$ms time-domain signal and $p(\tc)$ in Fig.~\ref{fig:TimeResponse}.
    Orange lines on the bottom panels show the injected $\tc$.
    % The following uncertainties indicate the 68\% credible intervals.  
    The recovered collapse time is $\tc=10.1^{+0.3}_{-0.4}\,$ms. 
    The primary post-merger oscillation frequency is $f_0=3317\pm 11\,$Hz with an exponential decay time-constant of $T_0=6.2_{-0.6}^{+ 0.7}\,$ms. 
    The linear frequency-drift term for the fundamental frequency is $\alpha_0=-0.79\pm 0.31\,$Hz. The frequencies corresponding to the sub-dominant modes are $f_1=2880^{+65}_{-56}\,$Hz and $f_2=2488^{+32}_{-35}\,$Hz.

\vspace{-0.5cm}
\begin{figure}[h]
        \centering
        \includegraphics[scale=0.4]{{NRTCAnalyticRuns_index5.0_2xAplus_snr34.31_distance3.37_THC:0036:R03_tc10.0_ra2.37_dec1.49_seed9_znFalse_mod_corner}.pdf}
        \caption
        {Full posteriors corresponding to the $\tc=10\,$ms injection shown in Fig.~\ref{fig:TimeResponse}. The  numerical-relativity simulation is injected into a 2A+ detector network at a distance of 3.04\,Mpc with an SLy equation of state and equal-mass $1.35\,\msun$ neutron stars. Credible intervals of 68\% are shown as dashed blue lines in the one-dimensional posteriors. The injected collapse time is shown by the orange lines in the lower-most panels.} 
        \label{fig:FullCorner}
\end{figure}
\end{subappendices}
\end{document}

### 2021-10-12 chapters/conclusion.tex
removed unwanted files from tracking

@@
 
-    The lifetimes of post-merger remnants are highly uncertain with gravitational-wave timescales ranging from $0.15-12\,\s$ for a range of ellipticities and equations of state (Eq.~\ref{eq:Intro:GWAngularMomentumLoss}).
+    The lifetimes of post-merger remnants are highly uncertain with gravitational-wave timescales ranging from $0.15-12\,\s$ for a range of ellipticities and equations of state (Eq.(\ref{eq:Intro:GWAngularMomentumLoss})).
     If the above waveform systematics are successfully addressed then detection and parameter estimation studies could be performed with much longer collapse times. 

### 2021-10-12 papers/paper2.tex
removed unwanted files from tracking

@@
         The priors on the model parameters are $\pi(\boldsymbol{\theta})$. 
-        The noise-weighted inner product in Eq.~\ref{eq:likelihood2} is defined by:
+        The noise-weighted inner product in Eq.(\ref{eq:likelihood2}) is defined by:
         \begin{align}
@@
     \end{equation}
-    The fitting factor, calculated with noise from one detector at Advanced LIGO design sensitivity~\cite{PSD:aLIGO}, quantifies the loss in signal-to-noise due to signal mismatch in relation to an optimal signal-to-noise ratio, Eq.~\ref{eq:OptimalSNR2}.
+    The fitting factor, calculated with noise from one detector at Advanced LIGO design sensitivity~\cite{PSD:aLIGO}, quantifies the loss in signal-to-noise due to signal mismatch in relation to an optimal signal-to-noise ratio, Eq.(\ref{eq:OptimalSNR2}).
 \par
@@
      We multiply the numerical-relativity injection, $d(t)$, by the Heaviside step function, $\mathcal{H}(t-t_{delay})$, and evaluate the model, $h(\boldsymbol{\theta},t-t_{delay})$, for $t \geq t_{delay}$.
-     The matched filter signal-to-noise ratio is calculated using Eq.~\ref{eq:MatchedFilterSNR2} with a single detector at Advanced LIGO sensitivity.
+     The matched filter signal-to-noise ratio is calculated using Eq.(\ref{eq:MatchedFilterSNR2}) with a single detector at Advanced LIGO sensitivity.
      We use numerical-relativity injection, SLy-M1.350-$\Lambda$390, selected due to compatibility with the tidal parameters inferred from GW170817.
@@
     The time-based phase was fit by the combination of a polynomial and exponentially-damped sinusoid using an iterative CMA-ES (covariance matrix adaption evolution search) fitting algorithm.
-    The maximum fitting factors were  calculated in the time domain without noise weighting and are not directly comparable to the noise-weighted fitting factors calculated with Eq.~\ref{eq:inner_product2}.
+    The maximum fitting factors were  calculated in the time domain without noise weighting and are not directly comparable to the noise-weighted fitting factors calculated with Eq.(\ref{eq:inner_product2}).
     The maximum fitting factors were  $\sim\!0.92\,$\nobreakdash-$\,0.98$ for 95\% of the 54 waveforms.\par
@@
 \label{appendix:b2}
-    The priors are listed in Eqs.~\ref{eq:firstprior2}\Hyphdash*\ref{eq:w012} with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
+    The priors are listed in Eqs.(\ref{eq:firstprior2}\Hyphdash*\ref{eq:w012}) with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
     The mode number $j$ is limited to $\lbrace 0,1,2 \rbrace$ and the mode number $i$ is restricted to $\lbrace 0,1 \rbrace$. 
     The priors in Eqs.~\ref{eq:modesorting2}\Hyphdash*\ref{eq:w012} are constrained priors. 
-    These restrictions are enforced in addition to the standard priors. The prior in Eq.~\ref{eq:modesorting2} ensures that the maximum spectral amplitude of each mode is decreasing. 
+    These restrictions are enforced in addition to the standard priors. The prior in Eq.(\ref{eq:modesorting2}) ensures that the maximum spectral amplitude of each mode is decreasing. 
     This results in $f_0$ converging to the loudest peak. 

### 2021-10-12 papers/paper3.tex
removed unwanted files from tracking

@@
     The time, $t$, is defined, such that $\max\left[h_+^2(\boldsymbol{\theta},t)+h_\times^2(\boldsymbol{\theta},t)\right]$ occurs when $t=0$. 
-    The cross polarisation for Eq.~\ref{eq:hplus3} is found by applying a $\pi/2$ phase shift to $\psi_j$.
+    The cross polarisation for Eq.(\ref{eq:hplus3}) is found by applying a $\pi/2$ phase shift to $\psi_j$.
     More details on this model can be found in Ref.~\cite{Easter2020}.
@@
     We use \textsc{Bilby}, a Bayesian inference package~\cite{Ashton2019}, to obtain posteriors, $p(\boldsymbol{\tilde{\theta}}~|~s_{\varphi,\mathrm{c}}(t))$ from a numerical relativity injection with enforced collapse starting at $t=\tc$ with width $\tw=2$\,ms.
-    Posteriors are calculated using waveforms detailed in Eqs.~\ref{eq:hplus3}\Hyphdash*\ref{eq:tukeyfallingedge3}. 
+    Posteriors are calculated using waveforms detailed in Eqs.(\ref{eq:hplus3}\Hyphdash*\ref{eq:tukeyfallingedge3}). 
     See Ref.~\cite{Easter2020} for further details of the gravitational-wave likelihood, and Appendix~\ref{appendix:a3} for additional information on the priors.
@@
 
-    We perform numerical-relativity injections with the full waveform, Eq.~\ref{eq:scol3}, at a grid of distances and apply the above rules to determine whether we successfully recover $\tc$.
+    We perform numerical-relativity injections with the full waveform, Eq.(\ref{eq:scol3}), at a grid of distances and apply the above rules to determine whether we successfully recover $\tc$.
 \begin{figure*}
@@
     We inject post-merger numerical-relativity waveforms modified to collapse at $t=\tc$ and sample posteriors for $\boldsymbol{\tilde{\theta}}$.
-    We then calculate posterior waveforms, $h_{\varphi,c}(\boldsymbol{\tilde{\theta}},t)$, from Eqs.~\ref{eq:hcol3}\Hyphdash*\ref{eq:tukeyfallingedge3}.
+    We then calculate posterior waveforms, $h_{\varphi,c}(\boldsymbol{\tilde{\theta}},t)$, from Eqs.(\ref{eq:hcol3}\Hyphdash*\ref{eq:tukeyfallingedge3}).
     Fig.~\ref{fig:TimeResponse} shows example posterior waveforms for the plus polarisation (left panels) and collapse-time posterior distributions (right panels) for numerical-relativity injections using an SLy equation of state with equal-mass $1.35\,\mathrm{M}_\odot$ progenitor neutron stars.
@@
     Finally, we find no significant correlations between $\tc$ and other model parameters.
-    We also attempt to measure the maximum detection distance where we can recover $\tc=20$\,ms and find that limitations in the third-order exponentially-damped sinusoidal model, Eq.~\ref{eq:hplus3}, prevent recovery of such collapse times.
+    We also attempt to measure the maximum detection distance where we can recover $\tc=20$\,ms and find that limitations in the third-order exponentially-damped sinusoidal model, Eq.(\ref{eq:hplus3}), prevent recovery of such collapse times.
     For $\tc > 15$\,ms signals, the analytical model cannot successfully track the time-domain phase of the gravitational-wave strain.
-    This limitation could be overcome by increasing the complexity of the frequency evolution in Eq.~\ref{eq:hplus3}, possibly introducing a quadratic frequency evolution term~\cite{Bose2018}.
+    This limitation could be overcome by increasing the complexity of the frequency evolution in Eq.(\ref{eq:hplus3}), possibly introducing a quadratic frequency evolution term~\cite{Bose2018}.
     Additionally, unmodelled searches such as \textsc{BayesWave} could be modified to measure the collapse time of the remnant~\cite{Cornish2015,Littenberg2015,Chatziioannou2017,Torres-Rivas2019}.\par
@@
 \label{appendix:a3}
-    The priors are listed in Eqs.~\ref{eq:firstprior3}\Hyphdash*\ref{eq:w013} with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
+    The priors are listed in Eqs.(\ref{eq:firstprior3}\Hyphdash*\ref{eq:w013}) with $\mathcal{U}(a,b)$ representing a uniform prior distribution from $a$ to $b$. 
     The mode number $j$ is limited to $\lbrace 0,1,2 \rbrace$. 

### 2021-10-12 chapters/introduction.tex
removed unwanted files from tracking

@@
     Movement of matter in space can generate gravitational waves.
-    For simulation purposes, the movement of matter can be modelled by the Einstein field equations~Eq.~\ref{eq:Intro:GeqT} and perturbations of black holes~\cite[e.g.,][]{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}.
+    For simulation purposes, the movement of matter can be modelled by the Einstein field equations~Eq.(\ref{eq:Intro:GeqT}) and perturbations of black holes~\cite[e.g.,][]{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}.
     This methodology allows the calculation of multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$.
@@
     For most sources, gravitational waves are predominantly emitted in the $l=m=2$ mode. 
-    In this case Eq.~\ref{eq:Intro:MultipoleExpansion} evaluates to
+    In this case Eq.(\ref{eq:Intro:MultipoleExpansion}) evaluates to
      \begin{align}
@@
 
-    The energy loss due to gravitational waves can be calculated from Eq.~\ref{eq:Intro:MultipoleExpansion} as~\cite{Nagar2005}
+    The energy loss due to gravitational waves can be calculated from Eq.(\ref{eq:Intro:MultipoleExpansion}) as~\cite{Nagar2005}
      \begin{align}
@@
     where $\mathcal{M} \equiv \mu^{3/5} M^{2/5}$ is the chirp mass of the binary system. 
-    Using Eqs.~\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}, an expression for the frequency and the gravitational-wave strain magnitude can be calculated~~\cite[e.g.,][]{Cutler1994}:
+    Using Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}), an expression for the frequency and the gravitational-wave strain magnitude can be calculated~~\cite[e.g.,][]{Cutler1994}:
       \begin{align}
@@
     where $t$ is the time, $t_c$ is the time until coalescence, and $D$ is the luminosity distance to the source. 
-    The combination of Eq.~\ref{eq:Intro:chirp_f}~and~\ref{eq:Intro:chirp_ht} result in the chirp waveform detected in the inspiral of compact object mergers.
+    The combination of Eq.(\ref{eq:Intro:chirp_f})~and~Eq.(\ref{eq:Intro:chirp_ht}) result in the chirp waveform detected in the inspiral of compact object mergers.
     \par
@@
     % Damour2012 in terms of tidal coupling constant, $\kappa_2^T$
-    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.~\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}, can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
+    Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
     \begin{align}
@@
     where $h_{22}(f)$ is the Fourier transform of the $(l,m)=(2,2)$ gravitational-wave strain, and $\Psi(f)$ is the corresponding phase.
-    Here, $f$ is the frequency~(Eq.~\ref{eq:Intro:FreqEvolution}), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system.
+    Here, $f$ is the frequency~(Eq.(\ref{eq:Intro:FreqEvolution})), $\phi_c$ and $t_c$ are the phase and time of coalescence, and $\mathcal{A}$ is a factor that depends on the orientation of the system.\footnote{The orientation factor, $\mathcal{A}$, depends on angles intrinsic and extrinsic to the binary system.
 Intrinsic angles include the angles defining the relative orientation of the neutron star spins to the orbital plane.
@@
     Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
-    Solving eqs.~\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution} for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
+    Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
     \par
@@
     % Rezzolla2010 long early simulation showing conservation of M_adm and J_gw
-    Although Eq.~\ref{eq:Intro:GWAngularMomentumLoss} shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
+    Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
     Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.

### 2021-10-12 papers/paper1.tex
removed unwanted files from tracking

@@
+% !TeX root = ../Thesis.tex
 \documentclass[../Thesis.tex]{subfiles}
@@
        \noindent The hats indicate the whitened transformations of the neutron star parameters such that $\widehat{\mathbf{x}}\sim~ \mathcal{N}(0,1)=( \mathbf{x}-\mu_x)/\sigma_x$ where $\mu_x$ and $\sigma_x^2$ are the mean and variance of $\mathbf{x}$ respectively.
-The compactness parameter can be generated from Eq.~\ref{eq:Ccalc1}-\ref{eq:betacalc1} after determining the values for $a_0, a_1, b_0$, and $b_1$.
+The compactness parameter can be generated from Eqs.(\ref{eq:Ccalc1}-\ref{eq:betacalc1}) after determining the values for $a_0, a_1, b_0$, and $b_1$.
 Spectra can then be trivially generated given any mass, quadrupolar tidal coupling constant and frequency shift.
@@
     \section{Results}
-        Following the training step, we use Eq.~(\ref{eq:GenerateWaveform1}) to generate spectra.
+        Following the training step, we use Eq.(\ref{eq:GenerateWaveform1}) to generate spectra.
 Figure~\ref{fig:GoodPlotsWithWaveformGeneration} shows how well our generated spectra match the original numerical-relativity spectra.
@@
 We fit the large-scale structure of the numerical-relativity peaks well with some deviations in the small-scale structure.
-\par Figure~\ref{fig:FittingFactorsHistogram} shows a histogram of the noise-weighted fitting factor, Eq.~(\ref{eq:FF1}), between our cross-validated model prediction and the corresponding numerical-relativity spectra for all tested waveforms.
+\par Figure~\ref{fig:FittingFactorsHistogram} shows a histogram of the noise-weighted fitting factor, Eq.(\ref{eq:FF1}), between our cross-validated model prediction and the corresponding numerical-relativity spectra for all tested waveforms.
 The resulting histogram has a mean of 0.95 with a standard deviation of 0.03.
@@
         Results based on our trained model suggest that the model is selective and could potentially be used in parameter estimation of detected events.
-If posteriors for the mass and tidal coupling constant are able to be determined, then it is a simple step to calculate the posteriors for the compactness, Eq.~(\ref{eq:Ccalc1}), and the radius of the neutron star.
+If posteriors for the mass and tidal coupling constant are able to be determined, then it is a simple step to calculate the posteriors for the compactness, Eq.(\ref{eq:Ccalc1}), and the radius of the neutron star.
 This will be confirmed in future work using a Bayesian framework.

### 2021-10-12 chapters/introduction.tex
Removed unused comments

@@
 \addcontentsline{toc}{chapter}{Abstract}
-% Paper 1
-% Here we introduce a hierarchical model trained on numerical-relativity simulations, which can generate reliable post-merger spectra in a fraction of a second.
-Our spectra have mean fitting factors of 0.95, which compares to a fitting factor of 0.93 between different numerical-relativity codes that simulate the same physical system.
-This method is the first step towards generating large template banks of spectra for use in post-merger detection and parameter estimation.
-% Paper 2
-% Here we develop a fast, simple model that can generate gravitational waveforms, and show it can be used for both detection and parameter estimation of post-merger remnants.
-% The model consists of three exponentially-damped sinusoids with a linear frequency-drift term.
-% We test the model against nine equal-mass numerical-relativity simulations selected for emission of gravitational waves for $\gtrsim25$\,ms.
-% The median fitting factors between the model waveforms and numerical-relativity simulations exceed 0.90.
-% We detect remnants at a post-merger signal-to-noise ratio of $\ge 7$ using a Bayes-factor detection statistic with a threshold of 3000.
-% We can constrain the primary post-merger frequency to $\pm_{1.2}^{1.4}\%$ at post-merger signal-to-noise ratios of 15 with an increase in precision to $\pm_{0.2}^{0.3}\%$ for post-merger signal-to-noise ratios of 50.
-% The tidal coupling constant can be constrained to $\pm^{9}_{12}\%$ at post-merger signal-to-noise ratios of 15, and $\pm 5\%$ at post-merger signal-to-noise ratios of 50 using a hierarchical inference model.
-% Paper 3
-% We develop and test a method to measure the collapse time of post-merger remnants.
-% We show that for a GW170817-like event at $\sim\!40$\,Mpc, a network of Einstein Telescope with Cosmic Explorer is required to detect collapse times of $\sim\!10$\,ms.
-% For a two-detector network at A+ design sensitivity, post-merger remnants with collapse times of $\sim\!10\,\mathrm{ms}$ must be $\lesssim 10$\,Mpc to be measureable.
-% This increases to $\sim\!18\Hyphdash*26$\,Mpc if we include the proposed Neutron star Extreme Matter Observatory (NEMO), increasing the effective volume by a factor of $\sim\!30$.
-    % Surviving neutron star remnants from binary neutron star mergers are some of the most extreme forms of matter known in the Universe.
+
     Surviving remnants from binary neutron star mergers, remnants that have not yet collapsed to a black hole,  are some of the most extreme forms of matter known in the Universe.    
@@
     These multipoles can be evaluated by numerical-relativity simulations which allow the extraction of gravitational waves.
-% \redtext{Movement of matter in space generates mass multipoles that can be related to perturbations of black holes.
-%     These perturbations can be described in terms of the non-rotating metric for black holes, the Schwarzschild metric, and the Einstein field equations,~Eq.~\ref{eq:Intro:GeqT}.
-%     Non-spherical gauge-invariant perturbations, outlined in Ref.~\cite{Baiotti2008}, can generate master multipole functions of odd ($\mathcal{Q}^\times_{lm}$) and even ($\mathcal{Q}^+_{lm}$) parity, with angular indices of $(l,m)$, that can be estimated my numerical-relativity simulations (see also Refs.~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005}).}
-% \redtext{
-%     Gravitational waves can be extracted for a virtual observer on the 2-sphere of a Schwarzschild metric with an isotropic radius of $r$ from the source.
-%     The Schwarzschild radius for this metric is $r_S=r(1-M/2r)^2$, where $M$ is the enclosed mass~\cite[e.g.,][]{Baiotti2008}.
-%     Gravitational waves are extracted through non-spherical gauge-invariant perturbations of the Schwarzschild metric, resulting in multipoles  where $(l,m)$ are the angular indices~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}.\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}\par  % Matthews1962
-%     % Check out: Abrahams, Rezzolla, Rupright, Anderson, Anninos, Baumgarte, et al., Phys. Rev. Lett. 80, 1812 (1998) 
-%     % Rupright, Abrahams, Rezzolla, Phys. Rev. D 58, 044005 (1998).
-%     % Rezzolla, Abrahams, Matzner, Rupright, Shapiro, Phys. Rev. D 59, 064001 (1999).
-%     }
     The plus, $h_{+}$, and cross, $h_{\times}$, polarisations of the gravitational waves are related to the multipoles as follows~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}:\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}
@@
       f &  \propto \mathcal{M}^{- 5/8}(t_c-t)^{- 3/8},\label{eq:Intro:chirp_f}\\
-        % |h(t)| & \propto  \frac{\pi^{2/3}}{D}\left(\frac{576\mu^3 M^2}{5}\right)^{1/4}(t_c-t)^{-1/4}, \label{eq:Intro:chirp_ht_old} ,\\
         |h(t)| & \propto  \frac{\mathcal{M}^{5/4}}{D}(t_c-t)^{-1/4}, \label{eq:Intro:chirp_ht} 
-        % Calculated by Cutler1994 (2.12) + (2.11) + Chirp mass defn
     \end{align}
@@
     Another binary neutron star (possibly neutron star - black hole) merger, GW190425, was detected in 2019~\cite{GW190425Detection,GWTC2,PopGWTC2}.
-    %, along with another binary neutron star black hole binary with a high false alarm rate, GW190426\_152155.
 However, the luminosity distance of GW190425, $159^{+69}_{-72}\,$Mpc, and the broad sky localisation of the source, $8284\,\mathrm{deg}^2$, precluded any coincident electromagnetic observations~\cite{GW190425Detection}.
@@
     Therefore, to date, GW170817 is the only gravitational-wave signal that was observed with coincident electromagnetic radiation and will be discussed further in Section~\ref{sec:Intro:NeutronStarMergers}.
-    %, allowing the host galaxy to be located~\cite{GW170817multi}.
 
@@
     Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
-    % A sentence here about the lack of opacity to GW as opposed to EM?
-
-    % If a BNS system of Rinit is less than tmerge is left undisturbed then the BNS can merge.
-    % Formation channels.
-    % If a binary neutron star system is formed with a separation less than the cr
-    % From GW170817multi:
-    % This motivated the development of accurate models for the two-body, general-relativistic dynamics (Blanchet et al.
-1995;Buonanno & Damour 1999; Pretorius 2005; Baker et al.
-2006; Campanelli et al.
-2006; Blanchet 2014) that
-
-    % Damour2012 in terms of tidal coupling constant, $\kappa_2^T$
     Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
@@
 
-    % Metzger2018: Blue KN eject proposed cause: neutrino-irradiated, magnetically driver outflow from temporarily stable HMNS consistent with $\tc \sim 0.1-1\,\s$
-    % Gill2019: Assume jet powered by rotating BH with $M_{\mathrm{ej,blue}} \approx 0.025\,\msun$ then $\tc \sim 1\,\s$
-    % Murguia-Berthier2021: Assume jet powered by rotating BH: $\tc \approx1-1.7\,\s$
-    % Yu2018: long lived stable NS
-
-    % From Zhang2019:
-    % There are two explanations for the total delay time of $\sim 1.7\,$s of GRB 170817A after the merger of GW170817~\cite{Zhang2019}.
-    % The first involves a fireball where the emission is most likely from the photosphere radius. 
-    % Assuming that the delay time from breakout to GRB emission is $\sim\,2\s$, similar to the duration of the GRB emission, the energy estimate for GRB 170817A suggests a Lorentz factor of $\sim 4$.
-    % However, radio afterglow modelling suggests that the Lorentz factor should be $\gg 4$ for $t < 10\,$days after the merger. % Zhang2019[70]
-    % This implies that the delay time from breakout to GRB emission is $\ll 2\,\s$ and that the total delay time is mostly attributed to the time to launch a jet or the time for breakout to occur. 
-    % This scenario can still work if a post-merger remnant survives $\sim\,1\s$ before launching a jet.
-    % However, this means that similarity between the total delay time of $\sim\,1.7\s$ and the GRB burst durations of $\sim\,2\s$ are coincidental as they are produced by different mechanisms. \par
-    % The second explanation assumes that the majority of the total delay time is accounted by the time it takes from the breakout of the jet to the emission of the GRB~\cite{Zhang2018}.
-    % This can be estimated by the width of the first pulse in GRB 170817A which was $\sim 1\,\s$ but could have been longer due to the weak nature of the burst.
-    % This scenario does not depend on the type of post-merger remnant which may be either a neutron star or a black hole.
-
-    % [Go through above references in the first blue paragraph]}
-
-%     \begin{itemize}
-%         \item tidal/inspiral methods
-%         \item GW170817
-%         \item other detections
-%         \item Merger outcomes
-
-%     \end{itemize}
-
-% ***************************************************************************************************************************************************************
     \section{Numerical relativity simulations}
@@
     Each of these three requirements make these simulations difficult.
-    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.
+    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par
 
-    \par
-% Good overall references: Daverio2018, Dietrich2016
-
-% ******************************************************************************************************************************************************    
-    % Thesis: Maione2017
-    % Two different conformal Z4 formulations have been studied: the CCZ4 formulation [Alic2012, Alic2013], which retains the full equations and in this way is a covariant formulation, at the price of greater differences respect to BSSN-OK and the need to introduce an additional parameter k3 to be able to evolve also spacetimes containing black holes and the Z4c [Gundlach2005, Bernuzzi2010, Hilditch2013] formulation, which retains only the principal part of the equations, in order to be as close as possible to BSSN-OK, but is not fully covariant and needs the introduction of ad-hoc constraint preserving boundary conditions in order to avoid constraint violations being reflected back inside the grid at the outer boundaries [Ruiz2011].
-% ******************************************************************************************************************************************************    
     The initial conditions for binary neutron star mergers in general relativity are complex and a number of methods have been developed for this task.
@@
     This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
-    % Tsokaros2016: also KADATH [36], and SCRID [37,38]
     Once the initial conditions have been approximated then the space-time can be evolved.
@@
     Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}.
-    %, and conformally flat (Isenberg-Wilson-Matthews)~\cite{Isenberg2008}. %Conformal and Covariant Z4 (CCZ4)~\cite{Bona2003,Bona2004,Alic2012,Alic2013},  \par
     The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
@@
 
-    % The CCZ4 formalism is a fully convariant extension of the Z4 formalism with the exception of black hole simulations where covariance cannot be maintained. ~\cite{Alic2012}
-
-    % Alic2013: The main advantage of the CCZ4 formulation over the BSSNOK formulation is its ability to dynamically evolve the constraints equations and to couple them with a constraint-damping scheme that leads to a rapid suppression of the violations that are inevitably produced in numerical simulations.
-    % Takami2015: We also note that although we have recently developed a new code, i.e., WHISKYTHC, able to reach high-order convergence [21,22], we have not employed it here mostly because our focus is on the post-merger dynamics, where the large scale shocks reduce the accuracy of both codes to first order, and for which WHISKY appears to be slightly more robust.
-
-    % Tsokaros2016: In this way we find that although global quantities of the system, like the mass and angular momentum, have differences of the order of ≲0.02%, local quantities, like rest-mass density, extrinsic curvature or angular velocity, show pointwise differences that are much larger, of the order of ≲1%.
-These local differences are then responsible for a ≲0.02%, local quantities, like rest-mass density, extrinsic curvature or angular velocity, show pointwise differences that are much larger, of the order of ≲1%.
-These local differences are then responsible for a dephasing in the gravitational waves at the merger time (after approximately three orbits) of ∼1.4 radians
-
     General-relativisitic hydrodynamic methods include codes such as \texttt{Whisky}~\cite{Baiotti2005}, \texttt{WhiskyTHC}~\cite{Baiotti2005,Radice2014,Radice2014b,Radice2015}, and \texttt{BAM}~\cite{Brugmann2004,Brugmann2008,Thierfelder2011,Dietrich2015,Bernuzzi2016b,Dietrich2019a}.
@@
     Refs.~\cite{Bernuzzi2014,Dietrich2017b} use the \texttt{BAM}  evolution code with the Z4c formulation, Refs.~\cite{Radice2016,Radice2017,Radice2017a,Radice2018} use the \texttt{WhiskyTHC} evolution code with the Z4c formulation, and Ref.~\cite{Rezzolla2016} uses the \texttt{Whisky} evolution code with the BSSNOK formulation.
-\par % Check metric evolution code for Rezzolla2016, have traced back to Baiotti2008 and it looks like BSSNOK
+\par
 
@@
 
-    % Baiotti2017: A perhaps disappointing note that needs to be made, however, is that despite this abundance of numerical codes, systematic comparisons of them are seldom done, and the only cases we are aware of are the quantitative comparisons between the SACRA and Whisky codes [160, 490] and between the BAM and SpEC codes [110]; such comparisons are tedious and computationally expensive but important for gauging the systematic errors that the various codes can introduce.
-
-% ***************************************************************************************************************************************************************
     \section{Post-merger remnants} \label{sec:Intro:PMR}
-%     \begin{itemize}
-%         \item Supramassive, Hypermassive properties
-%         \item observational methods, 
-%     \end{itemize}
 
@@
     Upper limits of $\lesssim 2.3\,\msun$ for $\mtov$ were found using angular momentum and energy considerations~\cite{Shibata2019}, and X-ray afterglow modelling~\cite{Sarin2020}, respectively. \par
-    % Khadkikar2021 hot equation of state => wider uncertainties, relaxed constraints
-    % Margalit2017: Multimessenger of GW170817 => M, R 
-    % Rezzolla2018: Multimessenger of GW170817
-    % Shibata2019: Angular momentum and energy conservation considerations
-    % Sarin2020: X-ray afterglow
+
     Using $\mtov\approx 2.3\,\msun$ as a fiducial value, which is also consistent with the mass range of J0740+6620~\cite{Cromartie2020}, sets the threshold between supramassive and hypermassive neutron stars at $\approx 2.8\,\msun$.
@@
     The gravitational mass is lost by: secular evolution with constant differential rotation, reduction in the amount of differential rotation, and loss of thermal support.
-    % page 5 \Kaplan2014
-    % Gravitational mass $m_g = m$ in Eq.~\ref{eq:Intro:TOV:mass_gradient}-\ref{eq:Intro:TOV:pressure_gradient} 
-    % with the density $\rho$ in the same equations.
-    % To calculate baryonic mass use $\frac{dM_b}{dr}$ = \frac{4\pi r^2 /rho}{1 + \epsilon}\left(1-\frac{2m_g}{r}\right)
     This additional thermal support extends the lifetime of the surviving post-merger remnant~\cite[e.g.,][]{Bauswein2010}.
@@
     \end{align}
-    % 400 * (20/10) / (2.8/2.7) * (2.8/2.7)  *(M/2.8) (Ev/10)^2 / (R/20) = 400 (M/2.7) (Ev/10)^2 / ((R/20) * (20/10))
+
     where $E_\nu$ is the typical neutrino energy.
@@
 
-    % Palenzuela2015 Ev~10 -20 MeV, Sumiyoshi2020 30 MeV
     The timescale that sets the loss of angular momentum due to gravitational waves is given by~\cite{Paschalidis2012}
@@
     We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.\par
-    % GRLES Simulations of Neutron Star Mergers: https://www.youtube.com/watch?v=Nppxi9E1ljg David Radice's youtube channel
 
-    % In comparison, the emission of gravitational waves due to ellipticity and f-modes is approximated by:~\cite{Becker2009}
-    % \begin{align}
-    %     h_{\mathrm{GW}} \approx 6\times10^{-29} \left(\frac{M}{2.7\,\msun}\right)\left(\frac{R}{10\,\km}\right)^{2}\left(\frac{f_2}{3.3\,\kHz} \right)^2\left(\frac{d}{40\,\Mpc} \right)^{-1}\left(\frac{e}{0.2} \right), \label{eq:Intro:GWFModes}
-    % \end{align}
-    % which has been scaled to a distance of $40\,\Mpc$, the distance of GW170817.
     Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
-    % Rezzolla2010 long early simulation showing conservation of M_adm and J_gw
     Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
     Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
-
-    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.
-    % Andy comment on length
-
-%     \begin{itemize}
-%         \item Show Eos difference in GW strain
-%         \item remnant astroseismology and frequency response
-%         \item Quasi-universal relationships
-%         \item Hot cold Eos
-%     \end{itemize}
+    \redtext{Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.}
 
@@
     \caption{Frequency domain representation of the post-merger gravitational-wave strain shown on the right panel of Fig.~\ref{fig:Intro:GWStrain}. 
-%The numerical-relativity simulation uses the SLy equation of state with equal-mass $1.35\,\msun$ neutron stars injected at a post-merger signal-to-noise ratio of 50.
     Four main spectral peaks are visible in the gravitational-wave spectrum with the dominant peak, $\fpeak$, has the largest amplitude at $\sim 3300\,\Hz$.
@@
 
-    % fmode Shibata1992, Zhuge1994
-    %fspiral, f_2_0, f_2+0, fpeak
-        %  The peaks at frequencies $f_{2-0},f_{2+0}$ may result from coupling between a quasi-radial oscillation mode and $f_{\mathrm{peak}}$~\cite{Stergioulas2011}.
-        % The peak at frequency $f_{\mathrm{spiral}}$ may result from the slower rotation-rate of tidally-deformed matter at the outer edges of the post-merger remnant~\cite{Bauswein2015}.
     Two additional oscillation modes were found in simulations that were caused by the interaction between the main f\Hyphdash*mode oscillation ($m=2$) and the $m=0$ radial pressure mode~\cite{Stergioulas2011}.
-    % Reference~\cite{Stergioulas2011} proposed that the peaks above and below $\fpeak$ can be explained by the interaction between the dynamics associated with $\fpeak$ ($m=2$) and the $m=0$ radial pressure mode.
     This results in a modulated signal that should be present in the frequency response at frequencies $\fpeak \pm f_{m=0}$, designated as $\ftwominuszero$ and $\ftwopluszero$, respectively.
-    %\par
     The $\ftwominuszero$ peak was found to be particularly strong in spectra for high-mass mergers as long as prompt collapse does not occur~\cite{Bauswein2015}.
@@
 \par
-
-% \begin{figure}[H]
-%     \centering
-%     \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{{thesis_GW_spectrogram_5x7T_reduced_no_lines}.png}}%
-%     \caption{Spectrograms (Figure from Ref.~\cite{Rezzolla2016}\protect\footnotemark) showing the frequency evolution of the post-merger remnant over time for equal-mass $1.35\,\msun$ neutron star mergers using five different equations of state at a distance of $50\,\Mpc$.
-The amplitude of the spectrogram is indicated by the colour bar at the top of the image.
-The dominant post-merger oscillation frequency,  $\fpeak$, is shown as a dark-green solid line.} \label{fig:Intro:Rezzolla2016Spectrograms}
-% \end{figure}
-%     Figure~\ref{fig:Intro:Rezzolla2016Spectrograms} \footnotetext{This figure is a  subset of Fig.~4 from Ref.~\cite{Rezzolla2016}\label{foot:Intro:Rezzolla2016imageref}.} shows spectrograms for the numerical-relativity simulations of the post-merger gravitational-wave strain at $50\,\Mpc$ using five different equations of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Rezzolla2016}.\footnote{See footnote \footref{foot:Intro:Rezzolla2016imageref}.}
-%     The equations of state progress from the stiffest, GNH3, in the left panel, to the softest, APR4, in the right panel.
-%     The dominant post-merger frequency, $\fpeak$, is shown as a solid dark-green line. 
-%     The magnitude of the spectrogram for a given frequency and time is indicated by the colour bar above the image.
-%     In all panels except for APR4, there is some positive drift in the dominant post-merger frequency over time.
-%     Furthermore, the spectrogram for GNH3 shows that there is also some frequency drift for the lower frequency peak at $\sim~1500\Hyphdash2000\,\Hz$.\footnote{Also note that the large frequency increase in the ALF2 spectrogram at $\sim 20\,\ms$ is due to the collapse of the remnant to a black hole.}
 
@@
     Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
-    % Bernuzzi2012: 2-4rad compared to post-Newtonian Taylor T4 at the time of merger, this is wrong it is the pN approximate that has error and this work is only interested up to contact.
-    % Radice2014b: WhiskyTHC 0.6 rad between low and high resolutions
     Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
@@
 
-%         \item Show Eos difference in GW strain
-%         \item remnant astroseismology and frequency response
-
-%  East2019a, Hotokezaka->2016PhRvD..93f4082H, Alic2013
-
-    % ordered: Andersson1998b, Bauswein2012, Bauswein2012a, Yagi2013,Yagi2013a, Hotokezaka2013
-
-    % others: Andersson1998b, Bauswein2012, Bernuzzi2014, Rezzolla2018, Iosif2020, Rezzolla2016, Bose2018, Sun2020
-
-    % Bauswein2012,Bauswein2012a,Hotokezaka2013,Bernuzzi2014,Takami2014,Bernuzzi2015,Bauswein2015,Takami2015,Rezzolla2016,Bauswein2019
-
-    % Andersson1998b: time constant for f-mode maps o compactness
-
     Quasi-universal relationships have been found relating the properties of the post-merger remnant and the progenitor neutron stars. 
@@
 \par
-%     \begin{figure}[H]
-%     \centering
-%         \includegraphics[scale=1.0]{{LambdaF2}.png}
 
-%         \caption{Quasi-universal relationship between the dominant post-merger oscillation frequency, shown here as $f_2$, and the dimensionless quadrupolar tidal deformability, $\Lambda$, shown here as $(\lambda/\bar{M}^5)^{1/5}$~(Image Ref.~\cite{Takami2015}).
-Each colour shows a different equation of state.}
-%         \label{fig:Intro:LambdaFpeak}
-
-% \end{figure}
     Relationships were then found between the progenitor neutron star ($l=2$) dimensionless quadrupolar tidal deformability, $\Lambda$, and a number of spectral features.
@@
     If the proposed Neutron star Extreme Matter Observatory is operational, then this distance increases to $\sim\!18\Hyphdash*26\,\Mpc$ thereby increasing the effective volume and hence detection rate by a factor of $\sim 30$.
-
-%      and \ref{chapter:CollapseTime} use Bayesian inference in a gravitational-wave framework using the \texttt{Bilby} package and \texttt{Dynesty} nested-sampler~\cite{Ashton2019,Speagle2019}.
-%     The base model for both of these chapters is a third-order exponentially-damped sinusoid with a frequency-drift term, Eqs.~\ref{eq:htotal2}-\ref{eq:hmode2}.
-%     In Chapter~\ref{chapter:DetectionPE}, we use this model for detection and parameter estimation of post-merger gravitational-wave strain.
-%     One major difference between this model and the hierarchical model is that the former is able to model the time domain of the gravitational-wave strain whereas the latter is only able to model the amplitude of the gravitational-wave spectra.
-%     In Chapter~\ref{chapter:CollapseTime}, we extend the model from Chapter~\ref{chapter:DetectionPE} to allow the measurement of the collapse time of the gravitational-wave strain.
-
-%     In Chapter~\ref{chapter:DetectionPE}, we use a $0.2\,\ms$ rising-edge Tukey window applied to the time-domain post-merger gravitational-wave strain to reduce artefacts due to Gibbs phenomenon.
-%     Gibbs phenomenon can cause unwanted oscillations in the time domain due to discontinuities in the signal amplitude~\cite{Gibbs1899}.
-%     The window rise-time is kept small so that we can measure the performance of the model to the coalescence time (Figs.~\ref{fig:OF_MF_SNR_FF5ms}-\ref{fig:TCOALESCENCE}) without unduly filtering out early post-merger signals.
-
-%     As each of the three modes of the base model are analytically identical, sampling of the model can be difficult.
-%     This is because there are six permutations of model parameters for every set of parameters.
-%     For example, if the posterior has support around frequencies $(f_0,f_1,f_2)=(1200, 1500, 3300)$ then all six permutations of these frequencies are also valid solutions for the posterior, including $(f_0,f_1,f_2)=(3300, 1200, 1500)$.
-%     This means that the posterior volume is effectively six times larger, making it much more difficult to sample. \par
-
-%     In Chapter~\ref{chapter:DetectionPE}, we find that we can remove the redundant permutations by ordering the maximum spectral amplitudes for each mode.
-%     The maximum spectral amplitude for the $j$th mode is given by $H_j = \max(|\tilde{h}_j(f)|\sqrt{f})$ and we enforce a descending order such that $H_0>H_1>H_2$  (see also Eq.~\ref{eq:modesorting2}).
-%     This ensures that the dominant post-merger f-mode peak, $\fpeak$, corresponds to mode zero (ie. $\fpeak=f_0$).
-%     We implement the mode sorting with constrained priors in \texttt{Bilby}.
-%     To find the posteriors, we use \texttt{Bilby} and \texttt{Dynesty} to perform nested sampling with 750 live points.
-%     If the number of live points is too small, then the sampler will loose valid parts of the posterior volume\footnote{If there are not enough live points within a particular mode then all live points can leave this mode by random chance.
-This is called mode `die-off'.
-The simplest solution to this is to increase the total number of live points.}, whereas an excessively large number of live points will take too long for the sampler to complete.
-%  \par
-
-%     Adding a collapse-time parameter to the post-merger gravitational-wave model makes convergence of the sampler more difficult.
-%     This is because the signal amplitude can be quite small at the time of collapse and the dimensionality of the system is increased by one.\footnote{The dimensionality would actually be increased by two, as we introduce two parameters: $\tc$ and $\tw$.
-For convergence reasons we fix $\tw=2\,\ms$.
-See Chapter~\ref{chapter:CollapseTime} for further information on this.}
-%     In response to this, in Chapter~\ref{chapter:CollapseTime}, we find that a number of additional measures are required to aid convergence. 
-%     Firstly, we sort the mode frequencies  in descending order, $f_0 > f_1 > f_2$, (Eq.~\ref{eq:freqsorting3}) in addition to the mode amplitude sorting used in Chapter~\ref{chapter:DetectionPE}.
-%     This limits our posteriors to spectral peaks at frequencies $\lesssim \fpeak=f_0$.
-%     Secondly, we implement a $2\,\ms$ rising-edge Tukey window to reduce sensitivity to the gravitational-wave strain for $\lesssim 2\,\ms$.
-%     Finally, we double the number of live points to 1500 to aid convergence of the sampler. \par
-
-%     However, to aid performance we implement mode sorting by  overriding the \texttt{log\_likelihood\_ratio} method of  \texttt{bilby.gw.likelihood.GravitationalWaveTransient}.
-%     When the maximum spectral amplitudes are not in descending order, the log likelihood method returns \texttt{-1e200}, effectively removing that point from the posterior. 
-%     This is also implemented by  overriding the \texttt{log\_likelihood\_ratio} method of  \texttt{bilby.gw.likelihood.GravitationalWaveTransient}.
-%     If the frequencies are not arranged in descending order then a log likelihood value of \texttt{-1e200} is returned and the spectral amplitudes are not calculated.
-% \par
-
-%     In Chapters~\ref{chapter:DetectionPE} and ~\ref{chapter:CollapseTime}, the total duration of the signal segment is $150\,\ms$ with a sampling frequency of $16384\,\Hz$.
-%     Furthermore, we use a rising-edge Tukey window beginning at the merger time, $t\equiv 0$ to reduce artefacts due to Gibbs phenomenon. 
-%     In Chapter~\ref{chapter:DetectionPE}, we use a $0.2\,\ms$ rising-edge Tukey window so that we can measure the performance of the model to the coalescence time (Figs.~\ref{fig:OF_MF_SNR_FF5ms}-\ref{fig:TCOALESCENCE}) without unduly filtering out early post-merger signals.
-%     For~\ref{chapter:CollapseTime}, as we are interested in measuring the collapse time for longer signals, $\gtrsim 5\,\ms$, we use a longer $2\,\ms$ rising-edge Tukey window.
-%     The parameters we use for \texttt{Dynesty}/\texttt{Bilby} in Chapter~\ref{chapter:DetectionPE} are:  $\texttt{npoints} = 750$, $\texttt{walks} = 150$.
-%     Whereas it was necessary to use slower, more robust parameters for Chapter~\ref{chapter:CollapseTime}: $\texttt{npoints} = 1500$, $\texttt{walks} = 150$, $\texttt{nact} = 10$, $\texttt{maxmcmc} = 15000$.
-
 \par
 
-    % We encounter problems with initial analytic injections in Chapter~\ref{chapter:CollapseTime} for collapse time runs.
-    % When we perform injections with all parameters fixed except for the collapse time and the exponential decay time-constant, with no injected noise and a small signal-to-noise ratio, we find that the collapse time posterior is biased\footnote{For zero noise injections, the gravitational-wave likelihood is still weighted by the power-spectral-density even though no Gaussian noise is added to the signal.
-This is why the signal-to-noise ratio has meaning for zero noise injections.}.
-    % We also find that the two-dimensional posteriors are not biased and that the bias is introduced when the one-dimensional posteriors are calculated.
-    % We also found that we could not simultaneously recover both $\tc$ and $\tw$ even in loud signals with zero injected noise and found that using a $\tw=2\,\ms$ was a good compromise allowing successful recovery for the collapse time.
-
-%     \begin{itemize}
-%         \item Machine learning
-%         \item Validation
-%         \item Bayesian inference
-%         \item nested sampling         
-%     \end{itemize}
-
 \end{document}

### 2021-10-12 chapters/introduction.tex
Updates from Overleaf

@@
 
-% ***************************************************************************************************************************************************************
     \section{Gravitational waves} 
@@
     These multipoles can be evaluated by numerical-relativity simulations which allow the extraction of gravitational waves.
+
     The plus, $h_{+}$, and cross, $h_{\times}$, polarisations of the gravitational waves are related to the multipoles as follows~\cite{Moncrief1974,Abrahams1996,Andrade1999,Nagar2005,Baiotti2008}:\footnote{Please note that $\mathcal{Q}^+_{lm}$ and $\mathcal{Q}^\times_{lm}$ have different units, see Ref.~\cite{Nagar2005} for more information.}
@@
     Another binary neutron star (possibly neutron star - black hole) merger, GW190425, was detected in 2019~\cite{GW190425Detection,GWTC2,PopGWTC2}.
+
     However, the luminosity distance of GW190425, $159^{+69}_{-72}\,$Mpc, and the broad sky localisation of the source, $8284\,\mathrm{deg}^2$, precluded any coincident electromagnetic observations~\cite{GW190425Detection}.
@@
 
-% ***************************************************************************************************************************************************************
     \section{Neutron stars}
@@
 
-% \par
-
-    % Another unique property of neutron stars is that they are opaque to neutrinos.
-
-% ***************************************************************************************************************************************************************
     \section{Neutron star mergers} \label{sec:Intro:NeutronStarMergers}
@@
     Hence gravitational waves are detectable over much larger distances in comparison to electromagnetic radiation with the same source luminosity. \par
+
     Gravitational-wave strain emitted from the inspiral of binary neutron star mergers, approximated by Eqs.(\ref{eq:Intro:FreqEvolution}\Hyphdash*\ref{eq:Intro:chirp_ht}), can be recast in the frequency domain using the stationary phase approximation~\cite{Cutler1994}
@@
     \par
-    % The base post-Newtonian expansion parameter is $\sqrt{M/r}$ but it is more convenient to use $\left(\pi M f\right)^{1/3}$ as the post-Newtonian expansion parameter.
-    % These two parameters are equal up to and excluding terms of order $(M/r)^{3/2}$.
+
     The validity for these equations can be extended to times closer to the merger by adding relativistic corrections calculated by post-Newtonian expansion and effective-one-body dynamics~\cite{Cutler1994,Blanchet1995, Buonanno1999,Droz1999,Damour2012,Yagi2013,Blanchet2014}.
-% to Eq.~\ref{eq:Intro:StationaryPhaseApproxPsi}
     These corrections introduce dependence on the masses, spins, and tidal deformabilities of the individual neutron stars.
@@
     Each of these three requirements make these simulations difficult.
-    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par
+    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.
+
+    \par
 
@@
     This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
+
     Once the initial conditions have been approximated then the space-time can be evolved.
@@
     Space-time evolution methods include the following formulations: Arnowitt-Deser-Misner (ADM)~\cite{Arnowitt2008}\footnote{The original book reference for ADM formalism is out of print and the authors have listed the cited paper as an `intentially unretouched' alternative.},  Baumgarte-Shapiro-Shibata-Nakamura-Oohara-Kojima (BSSNOK)~\cite{Nakamura1987,Shibata1995,Baumgarte1999,Alcubierre2000} and Z4c~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}.
+
     The ADM formalism is a Hamiltonian method that splits the Einstein's field equations into separate time and space components.
@@
     To aid computation convergence, the BSSNOK method extended the ADM formalism by introducing a conformal factor to the three-metric and selecting a dissipative hyperbolic driver for the  shift vector~\cite{Nakamura1987,Shibata1995,Nakamura1999,Shibata1999,Baumgarte1999,Alcubierre2000,Alcubierre2001}.
-% Dissipative hyperbolic driver from Alcubierre2001
     The Z4c formalism modified the BSSNOK method by adding a four-vector of constraints, $Z_\mu$, to the Einstein field equations~\cite{Bona2003,Bona2004,Gundlach2005,Bernuzzi2010,Hilditch2013}. 
@@
     Most postmerger numerical-relativity simulations last for only tens of milliseconds, though some longer simulations have been performed between 50-100\,ms in length~\cite{Rezzolla2010,Ciolfi2017,Ciolfi2019,DePietri2020}.
+    % Rezzolla2010 long early simulation showing conservation of M_adm and J_gw
     Although Eq.(\ref{eq:Intro:GWAngularMomentumLoss}) shows an estimated emission time for gravitational waves of $\sim 150\,$ms, the uncertainty in the equation of state could place $f_2$ as low as $\sim 2.2\,\kHz$ and some simulations show that the ellipticity reduces over time~\cite{Takami2015,Shibata2017a}.
     Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
-    \redtext{Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.}
+
+    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.
 
@@
     Two additional oscillation modes were found in simulations that were caused by the interaction between the main f\Hyphdash*mode oscillation ($m=2$) and the $m=0$ radial pressure mode~\cite{Stergioulas2011}.
+
     This results in a modulated signal that should be present in the frequency response at frequencies $\fpeak \pm f_{m=0}$, designated as $\ftwominuszero$ and $\ftwopluszero$, respectively.
+    %\par
     The $\ftwominuszero$ peak was found to be particularly strong in spectra for high-mass mergers as long as prompt collapse does not occur~\cite{Bauswein2015}.
@@
     Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
+
     Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
@@
     If the proposed Neutron star Extreme Matter Observatory is operational, then this distance increases to $\sim\!18\Hyphdash*26\,\Mpc$ thereby increasing the effective volume and hence detection rate by a factor of $\sim 30$.
-\par
 

### 2021-10-12 chapters/introduction.tex
Changes due to examination comments

--- 2021-10-12 chapters/introduction.tex
@@
+%!TeX root = ../Thesis.tex
 \documentclass[../Thesis.tex]{subfiles}
@@
 Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
-    Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
+    \redtext{
+    % Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
+    Equations~(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid until close to the merger time.
+    The orbital separation, $a$, at the merger time can be approximated by the innermost stable circular orbit, where $a \sim 6M$.
+    Therefore Eqns.(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid for $a \gtrsim 6M$~\cite{Cutler1994}.}
     Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
     \par
-    % The base post-Newtonian expansion parameter is $\sqrt{M/r}$ but it is more convenient to use $\left(\pi M f\right)^{1/3}$ as the post-Newtonian expansion parameter.
-    % These two parameters are equal up to and excluding terms of order $(M/r)^{3/2}$.
+
     The validity for these equations can be extended to times closer to the merger by adding relativistic corrections calculated by post-Newtonian expansion and effective-one-body dynamics~\cite{Cutler1994,Blanchet1995, Buonanno1999,Droz1999,Damour2012,Yagi2013,Blanchet2014}.
@@
     Each of these three requirements make these simulations difficult.
-    The equation of state for a `normal' neutron star is still unknown, and the equation of state for a hot post-merger remnant is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par
+    \redtext{The equation of state for cold neutron stars is still unknown, and the equation of state for hot post-merger remnants is even more uncertain,} after all, this is the hottest and densest bulk matter in the Universe.\par
 
@@
     Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
-    The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
-    This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
-    Quasiequilibrium configurations are also generated by \texttt{COCAL}.
-    This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
+    % The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
+    % This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
+    \redtext{The \texttt{LORENE} code used multi-domain spectral-methods to solve for the initial conditions of neutron star binaries assuming a conformally-flat approximation~\cite{Gourgoulhon2001}.
+    This was achieved by constructing quasiequilibrium configurations, which are approximations of the inspiral dynamics using exactly circular orbits.}
+    % Quasiequilibrium configurations are also generated by \texttt{COCAL}.
+    % This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
+    \redtext{The \texttt{COCAL} code has also been used to generate quasiequilibrium initial conditions for binary neutron stars mergers.
+    Here, the field equations are expressed in elliptic form and are solved by multipole expansion of Green's integral formula~\cite{Uryu2012,Tsokaros2015}.}
     Once the initial conditions have been approximated then the space-time can be evolved.
@@
     This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
-    Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this.
-\par
+    % Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this.
+    \redtext{Unfortunately, as ADM numerical simulations were weakly hyperbolic in some guages, errors were found to accumulate over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a}.
+    To counter this, a number of different formulations were devised to improve this.}  \par
 

### 2021-10-13 chapters/introduction.tex
Changes due to examination comments continued

@@
     This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
-    Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.
+    % Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.
 \par
+    \redtext{The introduction of $Z_\mu$, given an approapriate gauge, enforces hyperbolic solutions of the Einstein field equations by placing additional constraints on the system, thereby improving convergence~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.} \par
 
@@
     Secondly, numerical-relativity simulations unavoidably include numerical viscosity.
-    Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010}.
+    \redtext{Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010,Bernuzzi2012,Bernuzzi2012b}.}
     Numerical viscosity was found to introduce phase errors in the late-stage merger and improvements in numerical methods were subsequently introduced to minimise these errors~\cite{Radice2014}.

### 2021-10-13 chapters/introduction.tex
Changes due to examination comments continued

@@
     Ref.~\cite{Dietrich2018} state that their post-merger simulations from  the \texttt{CoRe} database are accurate enough to infer the energy and frequency content of the post-merger remnant.
-    This is an important point: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.
+    \redtext{Additionally, a recent examination of waveform systematics found that existing waveform models, including numerical-relativity simulations, were not accurate enough to inform the equation of state at very loud SNRs ($\gtrsim 100$) expected from third-generation detectors~\cite{Gamba2021}.
+    These are important points: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.}
 

### 2021-10-13 chapters/introduction.tex
Responses to examiners continued

@@
     The reference value, $f_2=3.3\,\kHz$, was inferred from our model in Chapter~\ref{chapter:DetectionPE} for numerical-relativity simulation of SLy equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2016}.
-    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.\par
+    We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.
+    \redtext{Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}.} \par
 
@@
     Given these possibilities, a four-fold decrease in $e$ together with a reduction in $f_2$ to $2.2\,\kHz$ would result in $t_{\mathrm{GW}} \sim 12\,$s.
-    \redtext{Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.}
+    Although the numerical-relativity simulations do not cover enough time period after the merger, they are vital for extracting the gravitational-wave strain in the early post-merger stage where the amplitude of the strain is at its loudest.
 
@@
     Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
+    \redtext{One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.}
       \par
@@
     Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
-    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests.
+    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests \redtext{(See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness)}.
     This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

### 2021-10-14 chapters/acknowledgements.tex
Response continued

@@
 
-The most important person for me to acknowledge is my wife Andrea. 
+The most important person for me to acknowledge is my wife. 
 It was only with her forbearance and encouragement that I was able to return to University, obtain my Honours in Advanced Science (Research) and continue on towards my doctorate over a total of a seven year period.
-I would also like to thank my daughters: Lauren, Sophie and Phoebe, who have all literally grown up during this period.
+I would also like to thank my daughters who have all literally grown up during this period.
 I look forward to spending more time with you after this submission.

### 2021-10-17 chapters/introduction.tex
Finished

@@
 Extrinsic angles include the angles defining the relative orientation between the orbital plane and each of the detectors.}
-    \redtext{
-    % Equations~\ref{eq:Intro:StationaryPhaseApproxH22}\Hyphdash*\ref{eq:Intro:StationaryPhaseApproxPsi} are valid for $a \gtrsim 6M$ where $a$ is the orbital separation of the binary system and $a\sim 6M$ is an approximation of the innermost stable circular orbit~\cite{Cutler1994}.
     Equations~(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid until close to the merger time.
     The orbital separation, $a$, at the merger time can be approximated by the innermost stable circular orbit, where $a \sim 6M$.
-    Therefore Eqns.(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid for $a \gtrsim 6M$~\cite{Cutler1994}.}
+    Therefore Eqns.(\ref{eq:Intro:StationaryPhaseApproxH22})\Hyphdash*(\ref{eq:Intro:StationaryPhaseApproxPsi}) are valid for $a \gtrsim 6M$~\cite{Cutler1994}.
     Solving Eqs.(\ref{eq:Intro:OrbitalEvolution}\Hyphdash*\ref{eq:Intro:FreqEvolution}) for $a \sim 6M$ gives $f_{max}\approx c^3 ( 6^{3/2}\pi G M )^{-1}\,\mathrm{Hz}$.
@@
     Each of these three requirements make these simulations difficult.
-    \redtext{The equation of state for cold neutron stars is still unknown, and the equation of state for hot post-merger remnants is even more uncertain,} after all, this is the hottest and densest bulk matter in the Universe.\par
+    The equation of state for cold neutron stars is still unknown, and the equation of state for hot post-merger remnants is even more uncertain, after all, this is the hottest and densest bulk matter in the Universe.\par
 
@@
     Codes that have been used for this include \texttt{LORENE}~\cite{Gourgoulhon2001} and \texttt{COCAL}~\cite{Tsokaros2015}.
-    % The \texttt{LORENE} formulation assumes a conformally-flat approximation for  neutron star binaries and uses spectral decomposition to solve for the initial conditions~\cite{Gourgoulhon2001}.
-    % This formulation constructs quasiequilibrium configurations which are approximations of the inspiral dynamics using exactly circular orbits.
-    \redtext{The \texttt{LORENE} code used multi-domain spectral-methods to solve for the initial conditions of neutron star binaries assuming a conformally-flat approximation~\cite{Gourgoulhon2001}.
-    This was achieved by constructing quasiequilibrium configurations, which are approximations of the inspiral dynamics using exactly circular orbits.}
-    % Quasiequilibrium configurations are also generated by \texttt{COCAL}.
-    % This is performed by excising regions around the neutron stars and solving Green's integral formula using multipole expansion and finite-difference methods~\cite{Uryu2012,Tsokaros2015}.
-    \redtext{The \texttt{COCAL} code has also been used to generate quasiequilibrium initial conditions for binary neutron stars mergers.
-    Here, the field equations are expressed in elliptic form and are solved by multipole expansion of Green's integral formula~\cite{Uryu2012,Tsokaros2015}.}
+    The \texttt{LORENE} code used multi-domain spectral-methods to solve for the initial conditions of neutron star binaries assuming a conformally-flat approximation~\cite{Gourgoulhon2001}.
+    This was achieved by constructing quasiequilibrium configurations, which are approximations of the inspiral dynamics using exactly circular orbits.
+    The \texttt{COCAL} code has also been used to generate quasiequilibrium initial conditions for binary neutron stars mergers.
+    Here, the field equations are expressed in elliptic form and are solved by multipole expansion of Green's integral formula~\cite{Uryu2012,Tsokaros2015}.
     Once the initial conditions have been approximated then the space-time can be evolved.
@@
     This is achieved by applying a gauge that consists of a shift vector and lapse function~\cite{Arnowitt2008}.
-    % Unfortunately, ADM numerical simulations were found to be unstable to accumulated errors over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a} and a number of different formulations were developed to improve this.
-    \redtext{Unfortunately, as ADM numerical simulations were weakly hyperbolic in some guages, errors were found to accumulate over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a}.
-    To counter this, a number of different formulations were devised to improve this.}  \par
+    Unfortunately, as ADM numerical simulations are weakly hyperbolic in some guages, errors were found to accumulate over the simulation time~\cite{Brandt2000,Kidder2001,Alcubierre2001a}.
+    To counter this, a number of different formulations were devised to improve this.  \par
 
@@
     This resulted in a propagating Hamiltonian constraint which helps to reduce constraint-violation growth by damping all constraint violations  except for constant modes~\cite{Gundlach2005}.
-    % Constraint violations can be more easily tracked as non-vanishing values of $Z_\mu$ though particular attention needs to be paid to reflection of constraint violations at boundaries~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.
+    The introduction of $Z_\mu$, given an approapriate gauge, enforces hyperbolic solutions of the Einstein field equations by placing additional constraints on the system, thereby improving convergence~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.
 \par
-    \redtext{The introduction of $Z_\mu$, given an approapriate gauge, enforces hyperbolic solutions of the Einstein field equations by placing additional constraints on the system, thereby improving convergence~\cite{Bernuzzi2010,Ruiz2011,Hilditch2013}.} \par
 
@@
     Secondly, numerical-relativity simulations unavoidably include numerical viscosity.
-    \redtext{Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010,Bernuzzi2012,Bernuzzi2012b}.}
+    Early simulations show that numerical viscosity could alter the properties of both the neutron star core and the surface~\cite{Shibata2000bar,Cerda-Duran2010,Bernuzzi2012,Bernuzzi2012b}.
     Numerical viscosity was found to introduce phase errors in the late-stage merger and improvements in numerical methods were subsequently introduced to minimise these errors~\cite{Radice2014}.
@@
     Ref.~\cite{Dietrich2018} state that their post-merger simulations from  the \texttt{CoRe} database are accurate enough to infer the energy and frequency content of the post-merger remnant.
-    \redtext{Additionally, a recent examination of waveform systematics found that existing waveform models, including numerical-relativity simulations, were not accurate enough to inform the equation of state at very loud SNRs ($\gtrsim 100$) expected from third-generation detectors~\cite{Gamba2021}.
-    These are important points: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.}
+    Additionally, a recent examination of waveform systematics found that existing waveform models, including numerical-relativity simulations, were not accurate enough to inform the equation of state at very loud SNRs ($\gtrsim 100$) expected from third-generation detectors~\cite{Gamba2021}.
+    These are important points: the uncertainty in the phase evolution  and collapse time of the post-merger remnant, which is discussed further in Sec.~\ref{sec:Intro:StrainAndSpectra}, has had a significant impact on the design of the models developed in this thesis.
 
@@
     We make rough estimate for the ellipticity,  $e\approx0.2$, from the time evolution of the density profile for numerical-relativity simulation of LS220 equation of state with equal-mass $1.35\,\msun$ neutron stars~\cite{Radice2017}.
-    \redtext{Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}.} \par
+    Numerical-relativity simulations have shown that the gravitational-wave timescale can range from $\mathcal{O}(20)\,\ms$ shortly after merger to $\gtrsim 0.5\,\s$ for longer lived simulations~\cite{Bernuzzi2016,Radice2018}.
+\par
 
@@
     Spectrograms from Refs.~\cite{Rezzolla2016,Dietrich2017,Dietrich2017a} and to a lesser extent from Ref.~\cite{Maione2017}, show some frequency drifts over time in some of the frequency peaks, particularly in $\fpeak$ which lends support to the implementation of a frequency-drift term in a post-merger gravitational-wave strain model.
-    \redtext{One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.}
+    One possible cause of frequency drift would involve the contraction or expansion of the post-merger remnant, which would increase and decrease the frequency, respectively.
       \par
@@
     Phase errors between successive resolutions accumulate over time and it is not unusual to see phase differences of $\sim 1$\,rad by the time of the merger for the finest resolutions~\cite[e.g.,][]{Radice2014b}.
-    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests \redtext{(See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness)}.
+    Reference~\cite{Takami2015} examined the role of the spatial resolution on the instantaneous phase of the gravitational-wave strain and found a phase difference of $0.9$\,rad between the the finest resolution tests (See also Ref.~\cite{Breschi2019} for a discussion on post-merger waveform robustness).
     This results in a loss of $\sim 40\%$ of the signal-to-noise compared to a perfectly matched signal when using a matched-filter detection. \par

### 2021-10-17 papers/paper1.tex
Removed unused comments
Finished

@@
-% !TeX root = ../Thesis.tex
+%!TeX root = ../Thesis.tex
 \documentclass[../Thesis.tex]{subfiles}
-
-% \documentclass[
-% % preprint,
-% % prx,
-% % prl,
-% % twocolumn,
-% %onecolumn,%!TEX encoding = UTF-8 Unicode
-% nofootinbib,
-% % ,secularistic%
-% tightenlines %
-% floatfix,
-% showpacs ,amssymb, aps,% linenumbers
-% superscriptaddress
-% ]{revtex4-1}
-% \documentclass[../main.tex]{subfiles}
-% \usepackage{amsmath,slashed}
-% \usepackage[left]{lineno}
-
-% \usepackage{comment}
-% \usepackage{graphicx}
-% \usepackage{amsmath,amssymb}
-% \usepackage{bm}
-% \usepackage[shortlabels]{enumitem}
-% \usepackage{float,soul}
-% \usepackage{natbib}
-% \usepackage{wasysym}
-% \usepackage{aas_macros}
-% \graphicspath{{images/}}
-% \RequirePackage{lineno}
-% \usepackage[usenames,dvipsnames]{color}
-
-% % \newcommand{\red}[1]{\textcolor{red}{#1}}
-% % \newcommand{\pdl}[1]{\textcolor{red}{[PDL: #1]}}
-% % \newcommand{\pje}[1]{\textcolor{Purple}{[PJE: #1]}}
-% % \newcommand{\blue}[1]{\textcolor{blue}{#1}}
-% \DeclareMathOperator*{\argmin}{arg\,min}
-% \newcommand{\SPA}{School of Physics and Astronomy, Monash University, Vic 3800, Australia}
-% \newcommand{\OzGravMonash}{OzGrav: The ARC Centre of Excellence for Gravitational Wave Discovery, Clayton VIC 3800, Australia}
 
@@
 \chapterauthor{Paul J. Easter \\ Paul D. Lasky\\Andrew R. Casey\\Luciano Rezzolla\\Kentaro Takami}
-% \email{paul.easter@monash.edu}
-% \author{Paul D. Lasky}
-% \email{paul.lasky@monash.edu}
-% \affiliation{\SPA}
-% \affiliation{\OzGravMonash}
-% \author{Andrew R. Casey}
-% \email{andrew.casey@monash.edu}
-% \affiliation{\SPA}
-% \affiliation{\OzGravMonash}
-% \affiliation{Faculty of Information Technology, Monash University, VIC 3800, Australia}
-% \author{Luciano Rezzolla}
-% \affiliation{Institut f\"ur Theoretische Physik, Max-von-Laue-Stra\ss e 1, 60438 Frankfurt, Germany}
-% \author{Kentaro Takami}
-% \affiliation{Kobe City College of Technology, 651-2194 Kobe, Japan}
 
@@
 \end{figure} 
-        It is possible that a phase transition may occur in the remnant during the post-merger phase due to an increase in density~(e.g.~\cite{Most2018,Bauswein2018}).
+        It is possible that a phase transition may occur in the remnant during the post-merger phase due to an increase in density~(e.g.~\cite{Most2018,Bauswein2019}).
 This could be detected by comparing the inferred values of the quadrupolar tidal coupling constant from the inspiral and post-merger phases.
@@
 Parameter estimation of the post-merger spectra could set bounds on the post-merger quadrupolar tidal coupling constant, allowing comparison with the inspiral value.
-This could determine whether a phase change in the equation of state is present~\cite{Bauswein2018,Most2018b}.\par
+This could determine whether a phase change in the equation of state is present~\cite{Bauswein2019,Most2018b}.\par
         To be valuable in search and parameter-estimation studies, our model must be extended to include individual values of mass, spins, compactness and quadrupolar tidal deformabilities for each progenitor.
